### 2017-12-18 eurocg/paper.tex
Dummy file for eurocg

% for section-numbered lemmas etc., use "numberwithinsect"
\documentclass[a4paper,english,numberwithinsect]{eurocg18}

% the recommended bibstyle
\bibliographystyle{plainurl}

%------------------------------------------------------------------- 
%if unwanted, comment out or use option "draft"
\usepackage{todonotes}
\usepackage{microtype}
\usepackage[noend, linesnumbered]{algorithm2e}
\usepackage{amsmath}
\usepackage{color}
\usepackage{cite}
\usepackage{pifont}
\usepackage{enumerate}

% Line numbers are helpful for refereeing
\usepackage[mathlines]{lineno}
\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
\expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
\expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
\renewenvironment{#1}%
     {\linenomath\csname old#1\endcsname}%
     {\csname oldend#1\endcsname\endlinenomath}}%
\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
  \patchAmsMathEnvironmentForLineno{#1}%
  \patchAmsMathEnvironmentForLineno{#1*}}%
\AtBeginDocument{%
\patchBothAmsMathEnvironmentsForLineno{equation}%
\patchBothAmsMathEnvironmentsForLineno{align}%
\patchBothAmsMathEnvironmentsForLineno{flalign}%
\patchBothAmsMathEnvironmentsForLineno{alignat}%
\patchBothAmsMathEnvironmentsForLineno{gather}%
\patchBothAmsMathEnvironmentsForLineno{multline}%
}
\linenumbers

%helpful if your graphic files are in another directory
%\graphicspath{{./graphics/}}

% Author macros::begin %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand\polylog{{\rm polylog}}

% Author metadata::begin %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Maximal Symmetric Partial Edge Drawings With Degree $ > 2$}
%optional, in case that the title is too long; 
%the running title should fit into the top page column
\titlerunning{MaxSPEDs with degree $ > 2$}

%% Please provide for each author the \author and \affil macro, 
%even when authors have the same affiliation, i.e. for each 
%author there needs to be the  \author and \affil macros
\author[1]{Michael H\"oller}
\author[2]{Fabian Klute}
\author[3]{Soeren Nickel}
\author[4]{Martin N\"ollenburg}
\author[5]{Birgit Schreiber}

\affil[1]{TU Vienna \texttt{}}
\affil[2]{Algorithms and Complexity Group TU Vienna \texttt{fklute@ac.tuwien.ac.at}}
\affil[3]{TU Vienna \texttt{}}
\affil[4]{Algorithms and Complexity Group TU Vienna \texttt{noellenburg@ac.tuwien.ac.at}}
\affil[5]{TU Vienna \texttt{}}
%mandatory. First: Use abbreviated first/middle names. 
%Second (only in severe cases): Use first author plus 'et. al.'
\authorrunning{F. Klute and M. N\"ollenburg}

% Author macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\martin}[1]{\todo[inline,color=blue!40]{MN: #1}}
\newcommand{\fabian}[1]{\todo[inline,color=pink!40]{FK: #1}}

\begin{document}

\maketitle

\begin{abstract}

\end{abstract}

\section{Introduction}

\section{Definitions}

\section{Hardness}
\fabian{Use the reduction from Till Bruckdorfers Dissertation \cite{bruckdorferschematics}. The Clauses can be modelled as triangles, the Variables as Circles with every second segment connected to a path. Path should have even length, segments are of length four.}

\section{Trees}
\fabian{Algorithm running in $O(k^2n)$ where $ k $ is the max degree of the tree.}

\section{Bridge? Cactus Graphs}
\fabian{Algorithm running in $ O(kn^2) $?}

\section{Series Parallel Graphs?}

\bibliography{paper}

\end{document}

### 2017-12-19 eurocg/paper.tex
section changes, added commands for todonotes
title change

@@
 % Author metadata::begin %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{Maximal Symmetric Partial Edge Drawings With Degree $ > 2$}
+\title{Maximizing Ink in Symmetric Partial Edge Drawings of $k$-plane Graphs}
+%{Maximal Symmetric Partial Edge Drawings With Degree $ > 2$}
 %optional, in case that the title is too long; 
@@
 \newcommand{\fabian}[1]{\todo[inline,color=pink!40]{FK: #1}}
+\newcommand{\birgit}[1]{\todo[inline,color=red!40]{BS: #1}}
+\newcommand{\michael}[1]{\todo[inline,color=green!40]{MH: #1}}
+\newcommand{\soeren}[1]{\todo[inline,color=orange!40]{SN: #1}}
 
@@
 
-\section{Trees}
-\fabian{Algorithm running in $O(k^2n)$ where $ k $ is the max degree of the tree.}
+\section{Two Polynomial Cases}
+\subsection{Trees}
+\fabian{Algorithm running in $O(kn)$ where $ k $ is the max degree of the tree.}
 
-\section{Bridge? Cactus Graphs}
-\fabian{Algorithm running in $ O(kn^2) $?}
+\subsection{Cacti of Degree Three}
+\fabian{Algorithm running in $ O(kn) $.}
 
-\section{Series Parallel Graphs?}
+\section{Conclusion}
 

### 2017-12-19 eurocg/paper.tex
edits from meeting

@@
 
-\section{Definitions}
+Martin
+
+\begin{itemize}
+	\item motivation
+	\item related work
+	\item contribution
+\end{itemize}
+
+\section{Preliminaries}
+
+Fabian
+
+\begin{itemize}
+	\item basic definitions and notation
+	\item segment intersection graph
+	\item alle planaren Graphen können auftreten (insbes. Bäume und Kakteen)
+\end{itemize}
+
+Input: $k$-plane graph drawing $\Gamma$ of a graph $G$ with edge set $S={s_1, \dots, s_m}$ (set of segments)
+
+Intersection graph $C=(V,E)$ with $V=S$ and edges if two segments intersect
+
+Consider maximum degree of $C$ -- here mostly 3.
 
@@
 
+Soeren?
+
 \section{Two Polynomial Cases}
+
+\begin{itemize}
+	\item dynamic programming
+	\item table entries $T_i(s)$ for $i = 1, \dots, \deg(s)$
+	\item $i$ corresponds to increasing stub lengths
+	\item stub lengths $l_i(s)$ for $i = 1, \dots, \deg(s)$ (not the pair of stubs)
+\end{itemize}
+
 \subsection{Trees}
-\fabian{Algorithm running in $O(kn)$ where $ k $ is the max degree of the tree.}
+\fabian{Algorithm running in $O(kn)$ where $ k $ is the max degree of the tree + intersection graph bauen.}
 
-\subsection{Cacti of Degree Three}
-\fabian{Algorithm running in $ O(kn) $.}
+Birgit
+
+\begin{itemize}
+	\item segment $s$ with $d$ children $u_1, \dots, u_{d}$ and parent $p(s) = p$.
+	\item Assume children are sorted by their intersection points on $s$, i.e., $u_1$ produces stub length $l_1$ etc.
+	\item for stub length relating to parent use offset from next longer stub length
+\end{itemize}
+
+\subsection{Cacti (of Degree Three)}
+\fabian{Algorithm running in $ O(kn) $ + intersection graph bauen.}
+
+\begin{itemize}
+	\item seems to work for all cactus graphs
+\end{itemize}
 

### 2017-12-19 eurocg/paper.tex
start of preliminaries, added macros for ped and maxsped

@@
 
+\newcommand{\ped}{\ensuremath{\textsc{PED}}\xspace}
+\newcommand{\maxsped}{\ensuremath{\textsc{MaxSped}}\xspace}
+
 \begin{document}
@@
 \section{Preliminaries}
+%\begin{itemize}
+%	\item basic definitions and notation
+%	\item segment intersection graph
+%	\item alle planaren Graphen können auftreten (insbes. Bäume und Kakteen)
+%\end{itemize}
+%Input: $k$-plane graph drawing $\Gamma$ of a graph $G$ with edge set $S={s_1, \dots, s_m}$ (set of segments)
+%Intersection graph $C=(V,E)$ with $V=S$ and edges if two segments intersect 
+%Consider maximum degree of $C$ -- here mostly 3.
 
-Fabian
+Let $ G $ be a \emph{simple graph} with edge set $ S = \{s_1,\dots,s_m\}$ and $ \Gamma $ a straight line drawing of $ G $ in the plane.
+We call $ \Gamma $ \emph{$ k $-plane} if every edge $ s_i \in S $ is crossed by at most $ k $ other edges from $ S $ in $ \Gamma $.
+The \emph{intersection graph} $ C = (V,E) $ of $ \Gamma $ is the graph containing one vertex in $ V $ for every $ s_i \in S $ and an edge $ uv \in E $ between vertices $ u,v \in V $ whenever the to $ u $ and $ v $ corresponding edges $ s_i, s_j \in S $ intersect in $ \Gamma $.
+Observe that the intersection graph $ C $ of a $ k $-plane drawing $ \Gamma $ has maximum degree $ k $.
+Its known that every planar graph is also representable as intersection graph of line segments \cite{chalopin2009every}.
 
-\begin{itemize}
-	\item basic definitions and notation
-	\item segment intersection graph
-	\item alle planaren Graphen können auftreten (insbes. Bäume und Kakteen)
-\end{itemize}
-
-Input: $k$-plane graph drawing $\Gamma$ of a graph $G$ with edge set $S={s_1, \dots, s_m}$ (set of segments)
-
-Intersection graph $C=(V,E)$ with $V=S$ and edges if two segments intersect
-
-Consider maximum degree of $C$ -- here mostly 3.
+\fabian{This part has to be adapted once the intro is fixed}
+We define a \emph{partial edge drawing} \ped $ \Gamma' $ of $ G $ as a drawing using the same embedding as $ \Gamma $, but edges are drawn as \emph{stubs}, i.e., we do not depict an edge $ s_i \in S $ by a line segment connecting the two vertices, but remove the middle part of the edges such that no two stubs intersect in $ \Gamma' $.
+We call a \ped $ \Gamma' $ \emph{symmetric} if for one edge $ s_is_j \in S $ its stub have equal lengths in $ \Gamma' $.
+Finally we define \maxsped as a symmetric \ped $ \gamma' $ of $ G $ where the sum of the lengths of the stubs is maximized.
 

### 2018-01-02 eurocg/paper.tex
author formatting
an introduction gearbeitet

@@
 \usepackage{enumerate}
+\usepackage[basic]{complexity}
 
@@
 % Author macros::begin %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newcommand\polylog{{\rm polylog}}
+%\newcommand\polylog{{\rm polylog}}
 
@@
 \author[1]{Michael H\"oller}
-\author[2]{Fabian Klute}
-\author[3]{Soeren Nickel}
-\author[4]{Martin N\"ollenburg}
-\author[5]{Birgit Schreiber}
+\author[1]{Fabian Klute}
+\author[1]{Soeren Nickel}
+\author[1]{Martin~N\"ollenburg}
+\author[1]{Birgit Schreiber}
 
-\affil[1]{TU Vienna \texttt{}}
-\affil[2]{Algorithms and Complexity Group TU Vienna \texttt{fklute@ac.tuwien.ac.at}}
-\affil[3]{TU Vienna \texttt{}}
-\affil[4]{Algorithms and Complexity Group TU Vienna \texttt{noellenburg@ac.tuwien.ac.at}}
-\affil[5]{TU Vienna \texttt{}}
+\affil[1]{Algorithms and Complexity Group, TU Wien, Vienna, Austria \texttt{}}
+%\affil[2]{Algorithms and Complexity Group, TU Wien, Vienna, Austria} %\texttt{fklute@ac.tuwien.ac.at}}
+%\affil[3]{Algorithms and Complexity Group, TU Wien, Vienna, Austria \texttt{}}
+%\affil[4]{Algorithms and Complexity Group, TU Wien, Vienna, Austria} %\texttt{noellenburg@ac.tuwien.ac.at}}
+%\affil[5]{Algorithms and Complexity Group, TU Wien, Vienna, Austria \texttt{}}
 %mandatory. First: Use abbreviated first/middle names. 
@@
 \newcommand{\ped}{\ensuremath{\textsc{PED}}\xspace}
-\newcommand{\maxsped}{\ensuremath{\textsc{MaxSped}}\xspace}
+\newcommand{\maxsped}{\ensuremath{\textsc{MaxSPED}}\xspace}
 
@@
 \begin{abstract}
-
+	TBD
 \end{abstract}
@@
 
-Martin
+Visualizing non-planar graphs as node-link diagrams is challenging due to the visual clutter caused by edge crossings.
+The layout readability deteriorates as the edge density and thus the number of crossings increases.
+Therefore alternative layout styles are necessary for non-planar graphs.
+A radical approach first used in applied work by Becker et al.~\cite{bew-vnd-95} is to start with a traditional straight-line graph drawing and simply drop a large central part of each edge and with it many of the edge crossings.
+This idea relies on the closure and continuation principles in Gestalt theory that based on the remaining edge stubs humans can still see the interrupted line segments by filling in the missing information in our brains.
+User studies have confirmed that such drawings remain readable while reducing clutter significantly~\cite{bvkw-epdldge-12}.
+
+Partial edge drawings have subsequently been defined as a formal notion of this idea~\cite{bk-ecbe-12,bcgkmn-pped-17}. 
+A \emph{partial edge drawing (PED)} is a graph drawing that maps vertices to points and edges to pairs of crossing-free edge stubs of positive length pointing towards each other.
+These edge stubs are obtained by erasing one contiguous central part of the straight-line segment connecting the two endpoints of each edge.
+In other words each straight-line edge is divided into three parts, of which only the two outer ones are drawn.
+More restricted variations of PEDs are \emph{symmetric} PEDs, in which both stubs of an edge must have the same length, and \emph{homogeneous} PEDs, in which the ratio of the stub length to the total edge length is the same for all edges.
+The natural optimization problem in this formal setting is \emph{ink maximization}, i.e., maximizing the total stub length, so that as much information as possible is given in the drawing while all crossings disappear in the negative background space.
+
+In this work we study the ink maximization problem for symmetric partial edge drawings (SPEDs) with a given geometric input drawing.
+This problem is denoted as \maxsped. 
+Bruckdorfer and Kaufmann~\cite{bk-ecbe-12} presented an integer linear program for solving \maxsped. 
+Later, Bruckdorfer et al.~\cite{bcgkmn-pped-17} gave an $O(n \log n)$-time algorithm for \maxsped on the class of 2-plane input drawings (no edge has more than two crossings), where $n$ is the number of vertices, and an efficient 2-approximation algorithm for the dual problem of minimizing the amount of erased ink for arbitrary input drawings.
+Bruckdorfer~\cite{b-sgh-15} further gives an \NP-hardness proof for \maxsped.
 
@@
 \section{Hardness}
-\fabian{Use the reduction from Till Bruckdorfers Dissertation \cite{bruckdorferschematics}.
+\fabian{Use the reduction from Till Bruckdorfers Dissertation \cite{b-sgh-15}.
 The Clauses can be modelled as triangles, the Variables as Circles with every second segment connected to a path.

### 2018-01-02 eurocg/paper.tex
intro soweit fertig

@@
 \usepackage[basic]{complexity}
+\usepackage{subcaption}
+\usepackage{wrapfig}
 
@@
 \begin{abstract}
-	TBD
+	Partial edge drawing (PED) is a drawing style for non-planar graphs, in which edges are drawn only partially as pairs of opposing stubs on the respective end-vertices. 
+	In a PED, by erasing the central parts of edges, all edge crossings and the resulting visual clutter are hidden in the undrawn parts of the edges.
+	We study symmetric partial edge drawings (SPEDs), in which the two stubs of each edge are required to have the same length. 
+	It is known that maximizing the ink (or the total stub length) when transforming a straight-line drawing with crossings into a SPED is generally \NP-hard.
+	We show that the problem remains \NP-hard even for 3-plane input drawings.
+	For $k$-plane input drawings whose edge intersection graph forms a collection of trees or cacti we present polynomial time algorithms for ink maximization.
 \end{abstract}
@@
 Therefore alternative layout styles are necessary for non-planar graphs.
-A radical approach first used in applied work by Becker et al.~\cite{bew-vnd-95} is to start with a traditional straight-line graph drawing and simply drop a large central part of each edge and with it many of the edge crossings.
-This idea relies on the closure and continuation principles in Gestalt theory that based on the remaining edge stubs humans can still see the interrupted line segments by filling in the missing information in our brains.
-User studies have confirmed that such drawings remain readable while reducing clutter significantly~\cite{bvkw-epdldge-12}.
+A radical approach first used in applied network visualization work by Becker et al.~\cite{bew-vnd-95} is to start with a traditional straight-line graph drawing and simply drop a large central part of each edge and with it many of the edge crossings.
+This idea relies on the closure and continuation principles in Gestalt theory, which imply that humans can still see a full line segment based only on the remaining edge stubs by filling in the missing information in our brains.
+User studies have confirmed that such drawings remain readable while reducing clutter significantly~\cite{bvkw-epdldge-12,bkl-us-15}.
 
-Partial edge drawings have subsequently been defined as a formal notion of this idea~\cite{bk-ecbe-12,bcgkmn-pped-17}. 
+\begin{wrapfigure}[11]{r}{0.45\textwidth}
+	\centering
+	\includegraphics{export_full}
+	\hfill
+	\includegraphics{export_alpha_nontouching}
+	\caption{A straight-line graph drawing (a) and a maximum-ink symmetric partial edge drawing (b) of the same graph.}
+	\label{fig:examples}
+\end{wrapfigure}
+
+The idea of drawing edges only partially has been formalized in graph drawing as follows~\cite{bk-ecbe-12}. 
 A \emph{partial edge drawing (PED)} is a graph drawing that maps vertices to points and edges to pairs of crossing-free edge stubs of positive length pointing towards each other.
-These edge stubs are obtained by erasing one contiguous central part of the straight-line segment connecting the two endpoints of each edge.
-In other words each straight-line edge is divided into three parts, of which only the two outer ones are drawn.
-More restricted variations of PEDs are \emph{symmetric} PEDs, in which both stubs of an edge must have the same length, and \emph{homogeneous} PEDs, in which the ratio of the stub length to the total edge length is the same for all edges.
+These edge stubs are obtained by erasing one contiguous central piece of the straight-line segment connecting the two endpoints of each edge.
+In other words each straight-line edge is divided into three parts, of which only the two outer ones are drawn, see Fig.~\ref{fig:examples}.
+More restricted and better readable~\cite{blmt-pedhmitc-16} variations of PEDs are \emph{symmetric} PEDs, in which both stubs of an edge must have the same length (see Fig.~\ref{fig:examples}(b)), and \emph{homogeneous} PEDs, in which the ratio of the stub length to the total edge length is the same for all edges.
 The natural optimization problem in this formal setting is \emph{ink maximization}, i.e., maximizing the total stub length, so that as much information as possible is given in the drawing while all crossings disappear in the negative background space.
 
-In this work we study the ink maximization problem for symmetric partial edge drawings (SPEDs) with a given geometric input drawing.
-This problem is denoted as \maxsped. 
+We study the ink maximization problem for symmetric partial edge drawings (SPEDs) with a given geometric input drawing.
+This problem is known as \maxsped. 
 Bruckdorfer and Kaufmann~\cite{bk-ecbe-12} presented an integer linear program for solving \maxsped. 
@@
 
-\begin{itemize}
-	\item motivation
-	\item related work
-	\item contribution
-\end{itemize}
+\medskip
+
+\noindent\textbf{Contribution.} We extend the results of Bruckdorfer et al.~\cite{bcgkmn-pped-17} on 2-plane geometric graph drawings to $k$-plane graph drawings for $k > 2$. 
+In particular, we show that \maxsped is \NP-hard even for 3-plane input drawings. 
+However, for graph drawings whose edge intersection graphs are trees or cacti of maximum degree $k$, we give polynomial-time algorithms for solving \maxsped.
 
@@
 We call $ \Gamma $ \emph{$ k $-plane} if every edge $ s_i \in S $ is crossed by at most $ k $ other edges from $ S $ in $ \Gamma $.
-The \emph{intersection graph} $ C = (V,E) $ of $ \Gamma $ is the graph containing one vertex in $ V $ for every $ s_i \in S $ and an edge $ uv \in E $ between vertices $ u,v \in V $ whenever the to $ u $ and $ v $ corresponding edges $ s_i, s_j \in S $ intersect in $ \Gamma $.
+The \emph{intersection graph} $ C = (V,E) $ of $ \Gamma $ is the graph containing a vertex $v_i$ in $ V $ for every $ s_i \in S $ and an edge $ v_i v_j \in E $ between vertices $ v_i, v_j \in V $ if the corresponding edges $ s_i, s_j \in S $ intersect in $ \Gamma $.
 Observe that the intersection graph $ C $ of a $ k $-plane drawing $ \Gamma $ has maximum degree $ k $.
-Its known that every planar graph is also representable as intersection graph of line segments \cite{chalopin2009every}.
+It is known that every planar graph is representable as the intersection graph of line segments~\cite{chalopin2009every}.
 
 \fabian{This part has to be adapted once the intro is fixed}
-We define a \emph{partial edge drawing} \ped $ \Gamma' $ of $ G $ as a drawing using the same embedding as $ \Gamma $, but edges are drawn as \emph{stubs}, i.e., we do not depict an edge $ s_i \in S $ by a line segment connecting the two vertices, but remove the middle part of the edges such that no two stubs intersect in $ \Gamma' $.
-We call a \ped $ \Gamma' $ \emph{symmetric} if for one edge $ s_is_j \in S $ its stub have equal lengths in $ \Gamma' $.
-Finally we define \maxsped as a symmetric \ped $ \gamma' $ of $ G $ where the sum of the lengths of the stubs is maximized.
+
+We define a \emph{partial edge drawing} (\ped) $ \Gamma' $ of $ G $ as a drawing using the same embedding as $ \Gamma $, but edges are drawn as \emph{stubs}, i.e., we do not depict an edge $ s_i \in S $ by a line segment connecting the two vertices, but remove the middle part of the edges such that no two stubs intersect in $ \Gamma' $.
+We call a \ped $ \Gamma' $ \emph{symmetric} if for every edge $ s_i s_j \in S $ its two stubs have equal lengths in $ \Gamma' $.
+Finally we define the problem \maxsped as finding a symmetric \ped $ \Gamma' $ of $ G $ where the sum of the lengths of the stubs is maximized.
+This is also known as ink maximization~\cite{bcgkmn-pped-17}.
+
+\martin{Maybe we don't need the above paragraph as PED, SPED, maxSPED are all somewhat defined in the introduction}
 

### 2018-01-02 eurocg/paper.tex
minor edit in header
minor edit

@@
 
-\affil[1]{Algorithms and Complexity Group, TU Wien, Vienna, Austria \texttt{}}
+\affil[1]{Algorithms and Complexity Group, TU Wien, Vienna, Austria}
 %\affil[2]{Algorithms and Complexity Group, TU Wien, Vienna, Austria} %\texttt{fklute@ac.tuwien.ac.at}}
@@
 	We study symmetric partial edge drawings (SPEDs), in which the two stubs of each edge are required to have the same length. 
-	It is known that maximizing the ink (or the total stub length) when transforming a straight-line drawing with crossings into a SPED is generally \NP-hard.
+	It is known that maximizing the ink (or the total stub length) when transforming a straight-line drawing with crossings into a SPED is tractable for 2-plane input drawings, but generally \NP-hard.
 	We show that the problem remains \NP-hard even for 3-plane input drawings.
-	For $k$-plane input drawings whose edge intersection graph forms a collection of trees or cacti we present polynomial time algorithms for ink maximization.
+	Yet, for $k$-plane input drawings whose edge intersection graph forms a collection of trees or cacti we present efficient algorithms for ink maximization.
 \end{abstract}
@@
 In particular, we show that \maxsped is \NP-hard even for 3-plane input drawings. 
-However, for graph drawings whose edge intersection graphs are trees or cacti of maximum degree $k$, we give polynomial-time algorithms for solving \maxsped.
+However, for $k$-plane graph drawings whose edge intersection graphs are collections of trees or cacti (which have maximum degree $k$), we give polynomial-time algorithms for solving \maxsped.
 

### 2018-01-03 eurocg/paper.tex
affiliation stuff
prelims

@@
 
-\affil[1]{Algorithms and Complexity Group, TU Wien, Vienna, Austria}
+\affil[1]{Algorithms and Complexity Group, TU Wien, Vienna, Austria\\ \texttt{[fklute|noellenburg|]@ac.tuwien.ac.at}}
 %\affil[2]{Algorithms and Complexity Group, TU Wien, Vienna, Austria} %\texttt{fklute@ac.tuwien.ac.at}}
@@
 %Second (only in severe cases): Use first author plus 'et. al.'
-\authorrunning{F. Klute and M. N\"ollenburg}
+\authorrunning{M. Höller, F. Klute, S. Nickel, M. N\"ollenburg, and B. Schreiber}
 
@@
 
-Let $ G $ be a \emph{simple graph} with edge set $ S = \{s_1,\dots,s_m\}$ and $ \Gamma $ a straight line drawing of $ G $ in the plane.
+Throughout the paper let $ G $ be a \emph{simple graph} with edge set $ S = \{s_1,\dots,s_m\}$ and $ \Gamma $ a straight line drawing of $ G $ in the plane.
 We call $ \Gamma $ \emph{$ k $-plane} if every edge $ s_i \in S $ is crossed by at most $ k $ other edges from $ S $ in $ \Gamma $.
@@
 Observe that the intersection graph $ C $ of a $ k $-plane drawing $ \Gamma $ has maximum degree $ k $.
-It is known that every planar graph is representable as the intersection graph of line segments~\cite{chalopin2009every}.
+Also it is known that every planar graph is representable as the intersection graph of line segments~\cite{chalopin2009every}.
 
-\fabian{This part has to be adapted once the intro is fixed}
-
-We define a \emph{partial edge drawing} (\ped) $ \Gamma' $ of $ G $ as a drawing using the same embedding as $ \Gamma $, but edges are drawn as \emph{stubs}, i.e., we do not depict an edge $ s_i \in S $ by a line segment connecting the two vertices, but remove the middle part of the edges such that no two stubs intersect in $ \Gamma' $.
+%\fabian{This part has to be adapted once the intro is fixed}
+%
+%We define a \emph{partial edge drawing} (\ped) $ \Gamma' $ of $ G $ as a drawing using the same embedding as $ \Gamma $, but edges are drawn as \emph{stubs}, i.e., we do not depict an edge $ s_i \in S $ by a line segment connecting the two vertices, but remove the middle part of the edges such that no two stubs intersect in $ \Gamma' $.
 We call a \ped $ \Gamma' $ \emph{symmetric} if for every edge $ s_i s_j \in S $ its two stubs have equal lengths in $ \Gamma' $.
@@
 This is also known as ink maximization~\cite{bcgkmn-pped-17}.
-
-\martin{Maybe we don't need the above paragraph as PED, SPED, maxSPED are all somewhat defined in the introduction}
+%
+%\martin{Maybe we don't need the above paragraph as PED, SPED, maxSPED are all somewhat defined in the introduction}
 

### 2018-01-04 eurocg/paper.tex
commit so I can work home

@@
 \newcommand{\maxsped}{\ensuremath{\textsc{MaxSPED}}\xspace}
+\newcommand{\sollong}{\ensuremath{\textit{long}}\xspace}
+\newcommand{\solshort}{\ensuremath{\textit{short}}\xspace}
 
@@
 
-\begin{itemize}
-	\item dynamic programming
-	\item table entries $T_i(s)$ for $i = 1, \dots, \deg(s)$
-	\item $i$ corresponds to increasing stub lengths
-	\item stub lengths $l_i(s)$ for $i = 1, \dots, \deg(s)$ (not the pair of stubs)
-\end{itemize}
+%\begin{itemize}
+%	\item dynamic programming
+%	\item table entries $T_i(s)$ for $i = 1, \dots, \deg(s)$
+%	\item $i$ corresponds to increasing stub lengths
+%	\item stub lengths $l_i(s)$ for $i = 1, \dots, \deg(s)$ (not the pair of stubs)
+%\end{itemize}
+Let $ C = (V,E) $ be the intersection graph of a graph $ G $ with edge set $ S $ and a drawing $ \Gamma $ of $ G $.
+Let $ u \in V $ with degree $ k $ and $ s \in S $ the corresponding edge in $ G $.
+There are $ k $ possible stub lengths $ \ell_1, \dots, \ell_k $.
+We assume $ \ell_1,\dots,\ell_k $ to be sorted from smaller to longer stubs.
+Additionally there is the possibility of drawing $ s $ as a complete segment which we set as $ \ell_0 $
 
 \subsection{Trees}
-\fabian{Algorithm running in $O(kn)$ where $ k $ is the max degree of the tree + intersection graph bauen.}
+Here we assume $ C = (V,E) $ to be a tree of maximum degree $ k $.
+We give a bottom up dynamic program solving the \maxsped problem on $ C $.
+We store the partial solutions for each vertex $ u \in V $ and stub length $ \ell_i $ of $ u $ in $ T_i(u) $.
+If $ u $ is a leaf there are exactly two such values.
+Now let $ u $ be a vertex with degree $ k' \leq k$.
+There are exactly $ k' + 1 $ possibilities for the drawn stubs of $ u $, still of those $ k' + 1 $ possibilities only two are of interest to us.
 
-Birgit
+\begin{figure}
+	\centering
+	\includegraphics{two_solutions_needed}
+	\caption{Consider $ u $, then only the blue and the green solution are relevant.}
+	\label{fig:two_solutions}
+\end{figure}
+Since we know $ C $ is a tree every vertex $ u \in V $ has a parent $ p(u) \in V $.
+For this parent there exists one stub length $ \ell_j(p(u)) $ in the sequence of stub lengths $ \ell_0(p(u)),\dots,\ell_{k'}(p(u)) $ such that drawing $ p(u) $ with these stubs does not intersect a completely drawn $ u $, while drawing $ p(u) $ with $ \ell_{j + 1}(p(u)) $, i.e.
+the next longer stub length, intersects $ \ell_0(u) $.
+We call the solution which would intersect the parent $ \sollong(u) $ and the one not intersecting the parent $ \solshort(u) $.
+More formally let $ j $ be the index such that stubs of length $ \ell_j(p(u)) $ do not intersect $ s(u) $, but $ \ell_{j +1}(p(u)) $ would.
+Then:
+\begin{align*}
+	\solshort(u) = \max\{T_1(u),\dots,T_j(u)\} \\
+	\sollong(u) = \max\{T_0(u),\dots,T_{k'}(u)\}
+\end{align*}
 
-\begin{itemize}
-	\item segment $s$ with $d$ children $u_1, \dots, u_{d}$ and parent $p(s) = p$.
-	\item Assume children are sorted by their intersection points on $s$, i.e., $u_1$ produces stub length $l_1$ etc.
-	\item for stub length relating to parent use offset from next longer stub length
-\end{itemize}
+Assuming the solutions for all children $ c(u) $ of $ u $ are computed we can calculate $ T_i(u) $ as:
+
+\begin{align*}
+	T_i(u) = \ell_i + \sum_{v \in c(u)}
+	\begin{cases}
+		\solshort(v) & \text{if } s(u) \text{ with length } \ell_i \text{ intersects } s(v) \\
+		\max\{\solshort(v),\sollong(v)\} & \text{else}
+	\end{cases}
+\end{align*}
+
+\begin{lemma}
+	\label{lem:compute_vertex}
+	Let $ C = (V,E) $ be a tree with maximum degree $ k $ and vertex weights $ \ell_0,\dots,\ell_{k'} $ for every $ u \in V $ with degree $ k' $.
+The set $ \{T_i(u) \mid 0 \leq i < k' \} $ can be computed in $ O(k') $ time and space, if the values $ \sollong(v) $ and $ \solshort(v) $ are computed for all children $ v \in c(u) $.
+\end{lemma}
+
+\begin{theorem}
+	Let $ G $ be a simple graph and $ \Gamma $ a straight-line drawing of $ G $, if the intersection graph $ C = (V,E) $ of $ G $ is a tree with maximum degree $ k \in \mathbb{N} $ the \maxsped problem can be solved in $ O(kn) $ time and space.
+\end{theorem}
+\begin{proof}
+	The intersection graph $ C = (V,E) $ and the necessary weights $ \ell_i $ can be computed in quadratic time in number of vertices of $ G $.
+
+	Let $ u \in V $ be a leaf, then computing $ T_i(u) $ takes constant time, since there are only two possible stub length, hence computing $ \sollong(u) $ and $ \solshort(u) $ takes also just constant time.
+
+	Let $ u \in V $ be not a leaf, but for all children $ v \in c(u) $ we know the values $ \solshort(v) $ and $ \sollong(v) $ for every stub length $ \ell_i(v) $.
+By Lemma~\ref{lem:compute_vertex} we can compute $ \sollong(u) $ and $ \solshort(u) $ in $ O(k) $.
+Hence if $ r \in V $ is the root of the tree $ C $, computing $ T_i(r) $ for all $ i $ takes $ O(nk) $ time.
+The maximum of
+\end{proof}
+
+%\begin{itemize}
+%	\item segment $s$ with $d$ children $u_1, \dots, u_{d}$ and parent $p(s) = p$.
+%	\item Assume children are sorted by their intersection points on $s$, i.e., $u_1$ produces stub length $l_1$ etc.
+%	\item for stub length relating to parent use offset from next longer stub length
+%\end{itemize}
 

### 2018-01-04 eurocg/paper.tex
minor
...

@@
 %the running title should fit into the top page column
-\titlerunning{MaxSPEDs with degree $ > 2$}
+%\titlerunning{MaxSPEDs with degree $ > 2$}
 
@@
 
-\section{Hardness}
+\section{Hardness of \maxsped for $k\ge3$}
 \fabian{Use the reduction from Till Bruckdorfers Dissertation \cite{b-sgh-15}.
@@
 
-\section{Two Polynomial Cases}
+\section{Two polynomial special cases}
 
@@
 Hence if $ r \in V $ is the root of the tree $ C $, computing $ T_i(r) $ for all $ i $ takes $ O(nk) $ time.
-The maximum of
+
+	It remains to show that the maximum over all $ T_i(r) $, with $ r \in V $ being the root, is the optimal value over all solutions to the \maxsped problem on $ G $.
+Assume $ t $ is the maximum value for some $ T_i(r) $'s and $ r $ the root.
+Let $ t' $ be the value of another solution on $ G $ with drawing $ \Gamma $ and $ t' > t $.
 \end{proof}

### 2018-01-05 eurocg/paper.tex
alpha version section 4.1

@@
 We call $ \Gamma $ \emph{$ k $-plane} if every edge $ s_i \in S $ is crossed by at most $ k $ other edges from $ S $ in $ \Gamma $.
+We use the terms edge in $ S $ and segment in $ \Gamma $ interchangeable, since there will always be a drawing $ \Gamma $ given with the graph $ G $.
+
 The \emph{intersection graph} $ C = (V,E) $ of $ \Gamma $ is the graph containing a vertex $v_i$ in $ V $ for every $ s_i \in S $ and an edge $ v_i v_j \in E $ between vertices $ v_i, v_j \in V $ if the corresponding edges $ s_i, s_j \in S $ intersect in $ \Gamma $.
@@
 Let $ u \in V $ with degree $ k $ and $ s \in S $ the corresponding edge in $ G $.
-There are $ k $ possible stub lengths $ \ell_1, \dots, \ell_k $.
-We assume $ \ell_1,\dots,\ell_k $ to be sorted from smaller to longer stubs.
-Additionally there is the possibility of drawing $ s $ as a complete segment which we set as $ \ell_0 $
+For a vertex $ u \in V $ there are $ k $ possible stub lengths $ \ell_1(u), \dots, \ell_k(u) $.
+We assume $ \ell_1(u),\dots,\ell_k(u) $ to be sorted from smaller to longer stubs.
+Additionally there is the possibility of drawing $ s $ as a complete segment which we set as $ \ell_0(u) $.
+We say an edge $ s $ is drawn with stub length $ \ell_i(u) $ to indicate that we draw only the two stubs of length $ \ell_i(u) $ and delete the middle part.
 
@@
 We give a bottom up dynamic program solving the \maxsped problem on $ C $.
-We store the partial solutions for each vertex $ u \in V $ and stub length $ \ell_i $ of $ u $ in $ T_i(u) $.
+We store the partial solutions for each vertex $ u \in V $ and stub length $ \ell_i(u) $ in $ T_i(u) $.
 If $ u $ is a leaf there are exactly two such values.
@@
 
-\begin{align*}
+\begin{align}
+	\label{rec:tree}
 	T_i(u) = \ell_i + \sum_{v \in c(u)}
@@
 	\end{cases}
-\end{align*}
+\end{align}
 
@@
 \end{lemma}
+\begin{proof}
+	Let $ u \in V $ be a vertex with degree $ k' < k $ for which we have computed $ \sollong $ and $ \solshort $ for every of its children.
+The crucial part is that $ T_i(u) $ can be built iteratively by going through the list of stub lengths for $ u $ in ascending order.
+First we compute $ T_1(u) $, this is simply the solution for $ u $ in which we can draw all its children and the parent with their full segments.
+This means we need to build the sum of $ \ell_1(u) $ and the $ \sollong(v) $ for every $ v \in c(u) $ which runs in $ O(k') $.
+The same can be done for $ T_0(u) $, but taking all the $ \solshort(v) $ values for the children $ v \in c(u) $.
+Assume we computed $ T_j(u) $ for all stub lengths $ \ell_(i) $ with $ i \leq j $.
+Now we want to compute $ T_{j+1}(u) $.
+By drawing $ s(u) $ with stub length $ \ell_{j +1} $ we make it a bit longer since we consider the $ \ell_i(u) $ in ascending order.
+Hence by doing so we now intersect the $ j $-th child $ v_j \in c(u) $.
+Since $ s(v_j) $ is the only segment affected by drawing $ u $ with stub length $ \ell_{j + 1} $ instead of $ \ell_j $ we can compute $ T_{j+1}(u) $ by subtracting simply $ \sollong(v) $ and adding $ \solshort(v) $.
+\end{proof}
 
@@
 Let $ t' $ be the value of another solution on $ G $ with drawing $ \Gamma $ and $ t' > t $.
+This solution has to be the sum of a set of stub lengths for each edge in $ G $ such that no two stubs intersect.
+Further we know the stubs can be categorized into two sets relative to their parent.
+Then if there was a solution $ t' $ with larger value than $ t $ there would necessarily be a third set, since Recurrence~\ref{rec:tree} builds just the maximum over the $ \solshort $ and $ \sollong $ values, such that no two stubs intersect.
 \end{proof}

### 2018-01-05 eurocg/paper.tex
alpha 4.1

@@
 
+Computing the intersection graph requires to identify all the intersections in $ \Gamma $.
+Using a standard sweep line algorithm \cite{}\fabian{add citation} this requires $ O(|S|\log |S| + I) $ where $ I $ is the number of crossings.
+Building the graph then takes $ O(|S|) $ time.
+When computing the intersection we can compute the stub lengths and their order as well.
+
 \subsection{Trees}
@@
 We give a bottom up dynamic program solving the \maxsped problem on $ C $.
-We store the partial solutions for each vertex $ u \in V $ and stub length $ \ell_i(u) $ in $ T_i(u) $.
-If $ u $ is a leaf there are exactly two such values.
-Now let $ u $ be a vertex with degree $ k' \leq k$.
-There are exactly $ k' + 1 $ possibilities for the drawn stubs of $ u $, still of those $ k' + 1 $ possibilities only two are of interest to us.
-
+For a vertex $ u \in V $ we denote with $ p(u) $ the parent of $ u $ in $ C $ and with $ c(u) $ the set of its children.
+The partial solutions for each vertex $ u \in V $ and stub length $ \ell_i(u) $ are stored in $ T_i(u) $.
 \begin{figure}
@@
 
-Assuming the solutions for all children $ c(u) $ of $ u $ are computed we can calculate $ T_i(u) $ as:
+Assuming the solutions for all children $ c(u) $ of $ u $ are computed we can calculate the $ T_i(u) $ as:
 
@@
 \end{align}
+
+\begin{lemma}
+	\label{lem:tree_correctness}	
+	Let $ G $ be a simple graph and $ \Gamma $ a straight-line drawing of $ G $, if the intersection graph $ C = (V,E) $ of $ G $ is a tree with root $ r \in V $ and maximum degree $ k $, then $ \max \{T_i(r) \mid 0 \leq i \leq k \} $ is an optimal solution to the \maxsped problem on $ G $ with drawing $ \Gamma $.
+\end{lemma}
+\begin{proof}
+	Let $ u \in V $ and for all children $ v \in c(u) $ we know the values $ \solshort(v) $ and $ \sollong(v) $ for every stub length $ \ell_i(v) $.
+It remains to show that the maximum over all $ T_i(r) $, with $ r \in V $ being the root, is the optimal value over all solutions to the \maxsped problem on $ G $.
+Assume $ t $ is the maximum value for some $ T_i(r) $'s and $ r $ the root.
+Let $ t' $ be the value of another solution on $ G $ with drawing $ \Gamma $ and $ t' > t $.
+This solution has to be the sum of a set of stub lengths for each edge in $ G $ such that no two stubs intersect.
+Further we know the stubs can be categorized into two sets relative to their parent.
+Then if there was a solution $ t' $ with larger value than $ t $ there would necessarily be a third set, since Recurrence~\ref{rec:tree} builds just the maximum over the $ \solshort $ and $ \sollong $ values, such that no two stubs intersect.
+
+\end{proof}
 
@@
 \begin{theorem}
-	Let $ G $ be a simple graph and $ \Gamma $ a straight-line drawing of $ G $, if the intersection graph $ C = (V,E) $ of $ G $ is a tree with maximum degree $ k \in \mathbb{N} $ the \maxsped problem can be solved in $ O(kn) $ time and space.
+	Let $ G $ be a simple graph and $ \Gamma $ a straight-line drawing of $ G $, if the intersection graph $ C = (V,E) $ of $ G $ is a tree with maximum degree $ k \in \mathbb{N} $ the \maxsped problem can be solved in $ O(n(k + \log n)) $ time and space where $ n = |V| $.
 \end{theorem}
 \begin{proof}
-	The intersection graph $ C = (V,E) $ and the necessary weights $ \ell_i $ can be computed in quadratic time in number of vertices of $ G $.
+	Since $ C $ is known to be a tree, and hence $ |E| = O(n) $, we can compute the intersection graph and the stub lengths in $ O(n\log n) $ time and space.
 
-	Let $ u \in V $ be a leaf, then computing $ T_i(u) $ takes constant time, since there are only two possible stub length, hence computing $ \sollong(u) $ and $ \solshort(u) $ takes also just constant time.
-
-	Let $ u \in V $ be not a leaf, but for all children $ v \in c(u) $ we know the values $ \solshort(v) $ and $ \sollong(v) $ for every stub length $ \ell_i(v) $.
-By Lemma~\ref{lem:compute_vertex} we can compute $ \sollong(u) $ and $ \solshort(u) $ in $ O(k) $.
-Hence if $ r \in V $ is the root of the tree $ C $, computing $ T_i(r) $ for all $ i $ takes $ O(nk) $ time.
-
-	It remains to show that the maximum over all $ T_i(r) $, with $ r \in V $ being the root, is the optimal value over all solutions to the \maxsped problem on $ G $.
-Assume $ t $ is the maximum value for some $ T_i(r) $'s and $ r $ the root.
-Let $ t' $ be the value of another solution on $ G $ with drawing $ \Gamma $ and $ t' > t $.
-This solution has to be the sum of a set of stub lengths for each edge in $ G $ such that no two stubs intersect.
-Further we know the stubs can be categorized into two sets relative to their parent.
-Then if there was a solution $ t' $ with larger value than $ t $ there would necessarily be a third set, since Recurrence~\ref{rec:tree} builds just the maximum over the $ \solshort $ and $ \sollong $ values, such that no two stubs intersect.
+	To solve the \maxsped problem on $ G $ we need to compute $ \max \{T_i(r) \mid 0 \leq i \leq k \} $ for the root $ r \in V $ of $ C $ by Lemma~\ref{lem:tree_correctness}.
+Doing this bottom up requires one pass through $ C $ hence by applying Lemma~\ref{lem:compute_vertex} we need $ O(nk) $ time to compute all $ T_i(u) $ for every $ u \in V $.
+For every vertex we only need to compute $ O(k) $ values and store two of them, namely $ \sollong(u) $ and $ \solshort(u) $, so we do not need more than $ O(nk) $ memory.
+Adding the bounds yields the desired runtime of $ O(n(k + \log n)) $.
 \end{proof}

### 2018-01-05 eurocg/paper.tex
dont know
alpha version section 4.1

@@
 Let $ C = (V,E) $ be the intersection graph of a graph $ G $ with edge set $ S $ and a drawing $ \Gamma $ of $ G $.
-Let $ u \in V $ with degree $ k $ and $ s \in S $ the corresponding edge in $ G $.
-For a vertex $ u \in V $ there are $ k $ possible stub lengths $ \ell_1(u), \dots, \ell_k(u) $.
-We assume $ \ell_1(u),\dots,\ell_k(u) $ to be sorted from smaller to longer stubs.
-Additionally there is the possibility of drawing $ s $ as a complete segment which we set as $ \ell_0(u) $.
-We say an edge $ s $ is drawn with stub length $ \ell_i(u) $ to indicate that we draw only the two stubs of length $ \ell_i(u) $ and delete the middle part.
+Let $ u \in V $ with degree $ k $, then for $ s(u) \in S $ there are $ k + 1 $ possible stubs including drawing the whole edge.
+We save the lengths of the possible stubs as node weights $ \ell_1(u), \dots, \ell_k(u) \in \mathbb{R}_+$.
+Further we assume $ \ell_1(u),\dots,\ell_k(u) $ to be sorted from smaller to longer stubs.
+The possibility of drawing $ s(u) $ as a complete segment we set as $ \ell_0(u) $.
 
@@
 \subsection{Trees}
-Here we assume $ C = (V,E) $ to be a tree of maximum degree $ k $.
+Here we assume $ C = (V,E) $ to be a rooted tree of maximum degree $ k $.
 We give a bottom up dynamic program solving the \maxsped problem on $ C $.
@@
 \end{figure}
-Since we know $ C $ is a tree every vertex $ u \in V $ has a parent $ p(u) \in V $.
+
+Since we know $ C $ is a tree every vertex $ u \in V $ with degree $ k' $ has a parent $ p(u) \in V $.
 For this parent there exists one stub length $ \ell_j(p(u)) $ in the sequence of stub lengths $ \ell_0(p(u)),\dots,\ell_{k'}(p(u)) $ such that drawing $ p(u) $ with these stubs does not intersect a completely drawn $ u $, while drawing $ p(u) $ with $ \ell_{j + 1}(p(u)) $, i.e.
@@
 We call the solution which would intersect the parent $ \sollong(u) $ and the one not intersecting the parent $ \solshort(u) $.
-More formally let $ j $ be the index such that stubs of length $ \ell_j(p(u)) $ do not intersect $ s(u) $, but $ \ell_{j +1}(p(u)) $ would.
-Then:
+See Figure~\ref{fig:two_solutions} for an illustration.
+More formally let $ j $ be the index such that stubs of length $ \ell_j(p(u)) $ do not intersect $ s(u) $, but $ \ell_{j +1}(p(u)) $ would, we set:
 \begin{align*}
@@
 
-Assuming the solutions for all children $ c(u) $ of $ u $ are computed we can calculate the $ T_i(u) $ as:
+It remains to define $ T_i(u) $ for a vertex $ u $ and stub length $ \ell_i(u) $:
 
@@
 	\label{rec:tree}
-	T_i(u) = \ell_i + \sum_{v \in c(u)}
+	T_i(u) = \ell_i(u) + \sum_{v \in c(u)}
 	\begin{cases}
@@
 
+Solving Recurrence~\ref{rec:tree} yields an optimal solution to the \maxsped problem on $ G $ with drawing $ \Gamma $ as the following lemma shows.
+
 \begin{lemma}
@@
 \begin{proof}
-	Let $ u \in V $ and for all children $ v \in c(u) $ we know the values $ \solshort(v) $ and $ \sollong(v) $ for every stub length $ \ell_i(v) $.
-It remains to show that the maximum over all $ T_i(r) $, with $ r \in V $ being the root, is the optimal value over all solutions to the \maxsped problem on $ G $.
-Assume $ t $ is the maximum value for some $ T_i(r) $'s and $ r $ the root.
-Let $ t' $ be the value of another solution on $ G $ with drawing $ \Gamma $ and $ t' > t $.
-This solution has to be the sum of a set of stub lengths for each edge in $ G $ such that no two stubs intersect.
+	Let $ G $, $ \Gamma $ and $ C = (V,E) $ as above and $ t = \max \{T_i(r) \mid 0 \leq i \leq k \} $ is the maximum value for some $ T_i(r) $'s and $ r $ the root.
+Now let $ t' $ be the value of a solution tp the \maxsped problem on $ G $ with drawing $ \Gamma $ and $ t' > t $.
+This solution has to be the sum of a set of stub lengths for each edge in $ G $ such that no two stubs intersect by definition.
 Further we know the stubs can be categorized into two sets relative to their parent.
-Then if there was a solution $ t' $ with larger value than $ t $ there would necessarily be a third set, since Recurrence~\ref{rec:tree} builds just the maximum over the $ \solshort $ and $ \sollong $ values, such that no two stubs intersect.
-
+Recurrence~\ref{rec:tree} builds just the maximum over the $ \solshort(v) $ and $ \sollong(v) $ values for all $ v \in c(u) $, such that no two stubs intersect.
+Hence if $ t $ was not optimal we would need to find at least a third stub length for at least one child $ v \in c(u) $ which was neither equal to $ \sollong(v) $ nor $ \solshort(v) $.
 \end{proof}
 
+\begin{figure}
+	\centering
+	\includegraphics{speedup}
+	\caption{The situation in Lemma~\ref{lem:compute_vertex}.}
+	\label{fig:speedup}
+\end{figure}
+Solving Recurrence~\ref{rec:tree} straight forward would yield an algorithm running in cubic time depending on the number of vertices in $ C $.
+Using the order on the stub lengths though we  can improve this to a quadratic running time by solving $ T_i(u) $ in linear time for one $ u \in V $.
 \begin{lemma}
@@
 The same can be done for $ T_0(u) $, but taking all the $ \solshort(v) $ values for the children $ v \in c(u) $.
-Assume we computed $ T_j(u) $ for all stub lengths $ \ell_(i) $ with $ i \leq j $.
-Now we want to compute $ T_{j+1}(u) $.
+Assume we computed $ T_j(u) $ for all stub lengths $ \ell_i(u) $ with $ i \leq j $.
+Now we want to compute $ T_{j+1}(u) $ (see Figure~\ref{fig:speedup}).
 By drawing $ s(u) $ with stub length $ \ell_{j +1} $ we make it a bit longer since we consider the $ \ell_i(u) $ in ascending order.
-Hence by doing so we now intersect the $ j $-th child $ v_j \in c(u) $.
+Hence by doing so we now intersect the corresponding edge of the $ j $-th child $ v_j \in c(u) $.
 Since $ s(v_j) $ is the only segment affected by drawing $ u $ with stub length $ \ell_{j + 1} $ instead of $ \ell_j $ we can compute $ T_{j+1}(u) $ by subtracting simply $ \sollong(v) $ and adding $ \solshort(v) $.
@@
 Adding the bounds yields the desired runtime of $ O(n(k + \log n)) $.
+
+	Retrieving the stub length for each edge in $ G $ can be done via backtracking the solutions which requires at most $ O(nk) $ time.
 \end{proof}

### 2018-01-05 eurocg/paper.tex
comments and minor edits

@@
 
-\affil[1]{Algorithms and Complexity Group, TU Wien, Vienna, Austria\\ \texttt{[fklute|noellenburg|]@ac.tuwien.ac.at}}
+\affil[1]{Algorithms and Complexity Group, TU Wien, Vienna, Austria\\ \texttt{[fklute|noellenburg]@ac.tuwien.ac.at}}
 %\affil[2]{Algorithms and Complexity Group, TU Wien, Vienna, Austria} %\texttt{fklute@ac.tuwien.ac.at}}
@@
 
-\section{Preliminaries}
+\section{Preliminaries}\label{sec:preliminaries}
 %\begin{itemize}
@@
 
+\martin{why do we need the reference here? Maybe better put it when it's actually used.}
+
+Using a standard sweep-line algorithm, computing the intersection graph $C$ of a set of $m$ line segments takes $O(m \log m + I)$ time~\cite{bcko-cgaa-08}, where $I$ is the number of intersections, i.e., the number of edges of $C$.
+
 %\fabian{This part has to be adapted once the intro is fixed}
@@
 
-\section{Hardness of \maxsped for $k\ge3$}
+\section{Hardness of \maxsped for $k\ge3$}\label{sec:hardness}
 \fabian{Use the reduction from Till Bruckdorfers Dissertation \cite{b-sgh-15}.
@@
 %\end{itemize}
-Let $ C = (V,E) $ be the intersection graph of a graph $ G $ with edge set $ S $ and a drawing $ \Gamma $ of $ G $.
-Let $ u \in V $ with degree $ k $, then for $ s(u) \in S $ there are $ k + 1 $ possible stubs including drawing the whole edge.
-We save the lengths of the possible stubs as node weights $ \ell_1(u), \dots, \ell_k(u) \in \mathbb{R}_+$.
-Further we assume $ \ell_1(u),\dots,\ell_k(u) $ to be sorted from smaller to longer stubs.
-The possibility of drawing $ s(u) $ as a complete segment we set as $ \ell_0(u) $.
+
+Section~\ref{sec:hardness} showed that \maxsped is generally \NP-hard for $k \ge 3$.
+In this section we consider the special case that the intersection graph of the $k$-plane input drawing is a tree or a cactus.
+In both cases we present polynomial-time dynamic programming algorithms.
+
+Let $ C = (V,E) $ be the intersection graph of the edge set $S$ in a given drawing $\Gamma$ of graph $ G $ as defined in Section~\ref{sec:preliminaries}.
+Let $ u \in V $ and $ \deg(u) = k $, then for the corresponding edge $ s(u) \in S $ there are $ k + 1 $ possible stub pairs including drawing the whole edge.
+We store the lengths of the possible stubs as node weights $ \ell_1(u), \dots, \ell_k(u) \in \mathbb{R}_+$ and assume that $ \ell_1(u),\dots,\ell_k(u) $ are sorted from shorter to longer stubs.
+We define  $ \ell_0(u) $ as the length of the whole edge $ s(u) $.
 
@@
 
+\martin{time to build graph should be $O(S+I)$; but maybe this preparation work should go to preliminaries so that we can assume the intersection graph as input}
+
 \subsection{Trees}
-Here we assume $ C = (V,E) $ to be a rooted tree of maximum degree $ k $.
-We give a bottom up dynamic program solving the \maxsped problem on $ C $.
-For a vertex $ u \in V $ we denote with $ p(u) $ the parent of $ u $ in $ C $ and with $ c(u) $ the set of its children.
-The partial solutions for each vertex $ u \in V $ and stub length $ \ell_i(u) $ are stored in $ T_i(u) $.
-\begin{figure}
+Here we assume that the intersection graph $ C = (V,E) $ is a rooted tree of maximum degree $ k $.
+We give a bottom up dynamic programming algorithm for solving \maxsped on $ C $.
+For a vertex $ u \in V $ let $ p(u) $ denote the parent of $ u $ in $ C $ and let $ c(u) $ denote the set of its children.
+The ink values of the partial \maxsped solutions for the subtrees $T_u$ rooted at each vertex $ u \in V $ such that stub length $ \ell_i(u) $ is used for $s(u)$ are computed and stored as $ T_i(u) $.
+
+\begin{figure}[tbp]
 	\centering
@@
 
+\martin{-- maybe $\delta$ instead of $k'$\\
+-- in the else case of $T_i$ long should be enough as always long $\ge$ short}
+
 It remains to define $ T_i(u) $ for a vertex $ u $ and stub length $ \ell_i(u) $:
@@
 	\label{lem:tree_correctness}	
-	Let $ G $ be a simple graph and $ \Gamma $ a straight-line drawing of $ G $, if the intersection graph $ C = (V,E) $ of $ G $ is a tree with root $ r \in V $ and maximum degree $ k $, then $ \max \{T_i(r) \mid 0 \leq i \leq k \} $ is an optimal solution to the \maxsped problem on $ G $ with drawing $ \Gamma $.
+	Let $ G $ be a simple graph and $ \Gamma $ a straight-line drawing of $ G $.
+If the intersection graph $ C = (V,E) $ of $ G $ is a tree with root $ r \in V $ and maximum degree $ k $, then $ \max \{T_i(r) \mid 0 \leq i \leq k \} $ is an optimal solution to the \maxsped problem on $ G $ with drawing $ \Gamma $.
 \end{lemma}
+
+\martin{could also use long(r) in the lemma}
+
 \begin{proof}
 	Let $ G $, $ \Gamma $ and $ C = (V,E) $ as above and $ t = \max \{T_i(r) \mid 0 \leq i \leq k \} $ is the maximum value for some $ T_i(r) $'s and $ r $ the root.
-Now let $ t' $ be the value of a solution tp the \maxsped problem on $ G $ with drawing $ \Gamma $ and $ t' > t $.
+Now let $ t' $ be the value of a solution to the \maxsped problem on $ G $ with drawing $ \Gamma $ and $ t' > t $.
 This solution has to be the sum of a set of stub lengths for each edge in $ G $ such that no two stubs intersect by definition.
@@
 \end{proof}
+
+\martin{could also use induction on the tree structure. not sure what's better.}
 
@@
 Using the order on the stub lengths though we  can improve this to a quadratic running time by solving $ T_i(u) $ in linear time for one $ u \in V $.
+
+\martin{the naïve solution would take $O(m\cdot k^2)$ and the improvement $O(mk)$, right?}
+
 \begin{lemma}
@@
 \end{lemma}
+
+\martin{The lemma should state the time to compute not only the $T_i(u)$ values but also long and short}
+
 \begin{proof}
@@
 First we compute $ T_1(u) $, this is simply the solution for $ u $ in which we can draw all its children and the parent with their full segments.
+	\martin{we don't draw the parent yet; it's not included in $T_i(u)$. Moreover the children don't need to be all full segments, but they are unrestricted.}	
 	This means we need to build the sum of $ \ell_1(u) $ and the $ \sollong(v) $ for every $ v \in c(u) $ which runs in $ O(k') $.
@@
 Hence by doing so we now intersect the corresponding edge of the $ j $-th child $ v_j \in c(u) $.
-Since $ s(v_j) $ is the only segment affected by drawing $ u $ with stub length $ \ell_{j + 1} $ instead of $ \ell_j $ we can compute $ T_{j+1}(u) $ by subtracting simply $ \sollong(v) $ and adding $ \solshort(v) $.
+Since $ s(v_j) $ is the only segment affected by drawing $ u $ with stub length $ \ell_{j + 1} $ instead of $ \ell_j $ we can compute $ T_{j+1}(u) $ in $O(1)$ time by simply subtracting $ \sollong(v) $ and adding $ \solshort(v) $.
 \end{proof}
@@
 \begin{theorem}
-	Let $ G $ be a simple graph and $ \Gamma $ a straight-line drawing of $ G $, if the intersection graph $ C = (V,E) $ of $ G $ is a tree with maximum degree $ k \in \mathbb{N} $ the \maxsped problem can be solved in $ O(n(k + \log n)) $ time and space where $ n = |V| $.
+	Let $ G $ be a simple graph with $m$ edges and $ \Gamma $ a straight-line drawing of $ G $.
+If the intersection graph $ C = (V,E) $ of $ G $ is a tree with maximum degree $ k \in \mathbb{N} $ the \maxsped problem can be solved in $ O(mk + m \log m)) $ time and space.
 \end{theorem}
 \begin{proof}
-	Since $ C $ is known to be a tree, and hence $ |E| = O(n) $, we can compute the intersection graph and the stub lengths in $ O(n\log n) $ time and space.
+	Since $ C $ is known to be a tree, and hence $ |E| = O(m) $, we can compute the intersection graph and the stub lengths in $ O(m\log m) $ time and space.
 
 	To solve the \maxsped problem on $ G $ we need to compute $ \max \{T_i(r) \mid 0 \leq i \leq k \} $ for the root $ r \in V $ of $ C $ by Lemma~\ref{lem:tree_correctness}.
-Doing this bottom up requires one pass through $ C $ hence by applying Lemma~\ref{lem:compute_vertex} we need $ O(nk) $ time to compute all $ T_i(u) $ for every $ u \in V $.
-For every vertex we only need to compute $ O(k) $ values and store two of them, namely $ \sollong(u) $ and $ \solshort(u) $, so we do not need more than $ O(nk) $ memory.
-Adding the bounds yields the desired runtime of $ O(n(k + \log n)) $.
+Doing this bottom up requires one pass through $ C $ hence by applying Lemma~\ref{lem:compute_vertex} we need $ O(mk) $ time to compute all $ T_i(u) $ for all $ u \in V $.
+For every vertex we only need to compute $ O(k) $ values and store two of them, namely $ \sollong(u) $ and $ \solshort(u) $, so we do not need more than $ O(mk) $ memory.
+Adding the bounds yields the desired runtime of $ O(mk + m \log m)) $.
 
-	Retrieving the stub length for each edge in $ G $ can be done via backtracking the solutions which requires at most $ O(nk) $ time.
+	Retrieving the actual stub length for each edge in $ G $ needed to produce the SPED can be done by backtracking the subsolutions which requires at most $ O(mk) $ time.
 \end{proof}
+
+\martin{the figures can still be improved in style and size; maybe merge into one, use nicer colors, draw just one segment and not multiple shifted ones etc}
 

### 2018-01-05 eurocg/paper.tex
beta 4.2

@@
 Observe that the intersection graph $ C $ of a $ k $-plane drawing $ \Gamma $ has maximum degree $ k $.
-Also it is known that every planar graph is representable as the intersection graph of line segments~\cite{chalopin2009every}.
-
-\martin{why do we need the reference here? Maybe better put it when it's actually used.}
 
@@
 In this section we consider the special case that the intersection graph of the $k$-plane input drawing is a tree or a cactus.
+Since every planar graph is representable as intersection graph of line segments~\cite{chalopin2009every}.
+we can also encounter every tree or cactus.
 In both cases we present polynomial-time dynamic programming algorithms.
@@
 
-Computing the intersection graph requires to identify all the intersections in $ \Gamma $.
+%Computing the intersection graph requires to identify all the intersections in $ \Gamma $.
 Using a standard sweep line algorithm \cite{}\fabian{add citation} this requires $ O(|S|\log |S| + I) $ where $ I $ is the number of crossings.
@@
 When computing the intersection we can compute the stub lengths and their order as well.
-
-\martin{time to build graph should be $O(S+I)$; but maybe this preparation work should go to preliminaries so that we can assume the intersection graph as input}
+%
+%\martin{time to build graph should be $O(S+I)$; but maybe this preparation work should go to preliminaries so that we can assume the intersection graph as input}
+%\fabian{add to prelims}
 
@@
 
-Since we know $ C $ is a tree every vertex $ u \in V $ with degree $ k' $ has a parent $ p(u) \in V $.
-For this parent there exists one stub length $ \ell_j(p(u)) $ in the sequence of stub lengths $ \ell_0(p(u)),\dots,\ell_{k'}(p(u)) $ such that drawing $ p(u) $ with these stubs does not intersect a completely drawn $ u $, while drawing $ p(u) $ with $ \ell_{j + 1}(p(u)) $, i.e.
+Since we know $ C $ is a tree every vertex $ u \in V $ with degree $ \delta $ has a parent $ p(u) \in V $.
+For this parent there exists one stub length $ \ell_j(p(u)) $ in the sequence of stub lengths $ \ell_0(p(u)),\dots,\ell_{\delta}(p(u)) $ such that drawing $ p(u) $ with these stubs does not intersect a completely drawn $ u $, while drawing $ p(u) $ with $ \ell_{j + 1}(p(u)) $, i.e.
 the next longer stub length, intersects $ \ell_0(u) $.
@@
 	\solshort(u) = \max\{T_1(u),\dots,T_j(u)\} \\
-	\sollong(u) = \max\{T_0(u),\dots,T_{k'}(u)\}
+	\sollong(u) = \max\{T_0(u),\dots,T_{\delta}(u)\}
 \end{align*}
-
-\martin{-- maybe $\delta$ instead of $k'$\\
--- in the else case of $T_i$ long should be enough as always long $\ge$ short}
-
 It remains to define $ T_i(u) $ for a vertex $ u $ and stub length $ \ell_i(u) $:
-
 \begin{align}
@@
 		\solshort(v) & \text{if } s(u) \text{ with length } \ell_i \text{ intersects } s(v) \\
-		\max\{\solshort(v),\sollong(v)\} & \text{else}
+		\sollong(v) & \text{else}
 	\end{cases}
 \end{align}
-
 Solving Recurrence~\ref{rec:tree} yields an optimal solution to the \maxsped problem on $ G $ with drawing $ \Gamma $ as the following lemma shows.
@@
 	Let $ G $ be a simple graph and $ \Gamma $ a straight-line drawing of $ G $.
-If the intersection graph $ C = (V,E) $ of $ G $ is a tree with root $ r \in V $ and maximum degree $ k $, then $ \max \{T_i(r) \mid 0 \leq i \leq k \} $ is an optimal solution to the \maxsped problem on $ G $ with drawing $ \Gamma $.
+If the intersection graph $ C = (V,E) $ of $ G $ is a tree with root $ r \in V $ and maximum degree $ k $, then $ \sollong(r) = \max \{T_i(r) \mid 0 \leq i \leq k \} $ is an optimal solution to the \maxsped problem on $ G $ with drawing $ \Gamma $.
 \end{lemma}
 
-\martin{could also use long(r) in the lemma}
+\begin{proof}
+	Let $ G $, $ \Gamma $ and $ C = (V,E) $ as above and $ u \in V $ a leaf of $ C $.
+Since $ c(u) $ is empty we simply get $ \solshort(u) = T_0(u) = \ell_1(u) $ and $ \sollong(u) = T_1(u) = \max\{\ell_0(u), \ell_1(u)\} $.
+Now let $ u \in V $ with degree $ \delta $ be an inner vertex.
+The values $ \solshort(v) $ and $ \sollong(v) $ are computed for all $ v \in c(u) $.
+$ T_i(u) $ for one $ 0 \leq i \leq \delta $ is then the sum of the maximum stubs we can draw among the children plus the stub length $ \ell_i(u) $ itself, which are just the values we are interested in.
+Setting then $ \sollong(u) $ and $ \solshort(u) $ as above gives the two necessary values for $ p(u) $.
 
-\begin{proof}
-	Let $ G $, $ \Gamma $ and $ C = (V,E) $ as above and $ t = \max \{T_i(r) \mid 0 \leq i \leq k \} $ is the maximum value for some $ T_i(r) $'s and $ r $ the root.
+%	and $ t = \sollong(r) $ is the maximum value for some $ T_i(r) $'s and $ r $ the root.
 Now let $ t' $ be the value of a solution to the \maxsped problem on $ G $ with drawing $ \Gamma $ and $ t' > t $.
@@
 
-\martin{could also use induction on the tree structure. not sure what's better.}
-
 \begin{figure}
@@
 \end{figure}
-Solving Recurrence~\ref{rec:tree} straight forward would yield an algorithm running in cubic time depending on the number of vertices in $ C $.
-Using the order on the stub lengths though we  can improve this to a quadratic running time by solving $ T_i(u) $ in linear time for one $ u \in V $.
-
-\martin{the naïve solution would take $O(m\cdot k^2)$ and the improvement $O(mk)$, right?}
+Solving Recurrence~\ref{rec:tree} straight forward would yield an algorithm running in $ O(mk^2) $.
+Using the order on the stub lengths though we can improve this to $ O(mk) $ by solving $ T_i(u) $ in linear time for one $ u \in V $.
 
@@
 	\label{lem:compute_vertex}
-	Let $ C = (V,E) $ be a tree with maximum degree $ k $ and vertex weights $ \ell_0,\dots,\ell_{k'} $ for every $ u \in V $ with degree $ k' $.
-The set $ \{T_i(u) \mid 0 \leq i < k' \} $ can be computed in $ O(k') $ time and space, if the values $ \sollong(v) $ and $ \solshort(v) $ are computed for all children $ v \in c(u) $.
+	Let $ C = (V,E) $ be a tree with maximum degree $ k $ and vertex weights $ \ell_0,\dots,\ell_{\delta} $ for every $ u \in V $ with degree $ \delta $.
+We can compute $ \solshort(u) $ and $ \sollong(u) $ in $ O(\delta) $ time and space, if the values $ \sollong(v) $ and $ \solshort(v) $ are computed for all children $ v \in c(u) $.
 \end{lemma}
-
-\martin{The lemma should state the time to compute not only the $T_i(u)$ values but also long and short}
-
 \begin{proof}
-	Let $ u \in V $ be a vertex with degree $ k' < k $ for which we have computed $ \sollong $ and $ \solshort $ for every of its children.
+	Let $ u \in V $ be a vertex with degree $ \delta < k $ for which we have computed $ \sollong $ and $ \solshort $ for every of its children.
 The crucial part is that $ T_i(u) $ can be built iteratively by going through the list of stub lengths for $ u $ in ascending order.
-First we compute $ T_1(u) $, this is simply the solution for $ u $ in which we can draw all its children and the parent with their full segments.
-	\martin{we don't draw the parent yet; it's not included in $T_i(u)$. Moreover the children don't need to be all full segments, but they are unrestricted.}	
-	This means we need to build the sum of $ \ell_1(u) $ and the $ \sollong(v) $ for every $ v \in c(u) $ which runs in $ O(k') $.
+First we compute $ T_1(u) $, this is simply the solution for $ u $ in which we can draw all its children with their full segments.
+	\fabian{make clear that we do not draw them, but could potentially do so if one wished for it}	
+	This means we need to build the sum of $ \ell_1(u) $ and the $ \sollong(v) $ for every $ v \in c(u) $ which runs in $ O(\delta) $.
 The same can be done for $ T_0(u) $, but taking all the $ \solshort(v) $ values for the children $ v \in c(u) $.
@@
 Since $ s(v_j) $ is the only segment affected by drawing $ u $ with stub length $ \ell_{j + 1} $ instead of $ \ell_j $ we can compute $ T_{j+1}(u) $ in $O(1)$ time by simply subtracting $ \sollong(v) $ and adding $ \solshort(v) $.
+
+	Finally we need to compute $ \solshort(u) $ and $ \sollong(u) $, but this are just two maximizations over at most $ \delta + 1 $ values at most.
+This can be done in $ O(\delta) $ by simply sweeping over the values.
 \end{proof}

### 2018-01-05 eurocg/paper.tex
beta 4.2

@@
 The crucial part is that $ T_i(u) $ can be built iteratively by going through the list of stub lengths for $ u $ in ascending order.
-First we compute $ T_1(u) $, this is simply the solution for $ u $ in which we can draw all its children with their full segments.
-	\fabian{make clear that we do not draw them, but could potentially do so if one wished for it}	
+First we compute $ T_1(u) $, this is the solution for $ u $ in which we are able to draw the children with any of their stub lengths.
 This means we need to build the sum of $ \ell_1(u) $ and the $ \sollong(v) $ for every $ v \in c(u) $ which runs in $ O(\delta) $.
@@
 By drawing $ s(u) $ with stub length $ \ell_{j +1} $ we make it a bit longer since we consider the $ \ell_i(u) $ in ascending order.
-Hence by doing so we now intersect the corresponding edge of the $ j $-th child $ v_j \in c(u) $.
+Hence by doing so we now potentially intersect the corresponding edge of the $ j $-th child $ v_j \in c(u) $.
 Since $ s(v_j) $ is the only segment affected by drawing $ u $ with stub length $ \ell_{j + 1} $ instead of $ \ell_j $ we can compute $ T_{j+1}(u) $ in $O(1)$ time by simply subtracting $ \sollong(v) $ and adding $ \solshort(v) $.
@@
 \subsection{Cacti (of Degree Three)}
-\fabian{Algorithm running in $ O(kn) $ + intersection graph bauen.}
 

### 2018-01-07 eurocg/paper.tex
change in Paper.tex (started on reduction)

--- 2018-01-05 eurocg/paper.tex
@@
 
-\affil[1]{Algorithms and Complexity Group, TU Wien, Vienna, Austria\\ \texttt{[fklute|noellenburg]@ac.tuwien.ac.at}}
+\affil[1]{Algorithms and Complexity Group, TU Wien, Vienna, Austria\\ \texttt{[fklute|noellenburg|]@ac.tuwien.ac.at}}
 %\affil[2]{Algorithms and Complexity Group, TU Wien, Vienna, Austria} %\texttt{fklute@ac.tuwien.ac.at}}
@@
 
-\section{Preliminaries}\label{sec:preliminaries}
+\section{Preliminaries}
 %\begin{itemize}
@@
 We call $ \Gamma $ \emph{$ k $-plane} if every edge $ s_i \in S $ is crossed by at most $ k $ other edges from $ S $ in $ \Gamma $.
-We use the terms edge in $ S $ and segment in $ \Gamma $ interchangeable, since there will always be a drawing $ \Gamma $ given with the graph $ G $.
-
 The \emph{intersection graph} $ C = (V,E) $ of $ \Gamma $ is the graph containing a vertex $v_i$ in $ V $ for every $ s_i \in S $ and an edge $ v_i v_j \in E $ between vertices $ v_i, v_j \in V $ if the corresponding edges $ s_i, s_j \in S $ intersect in $ \Gamma $.
@@
 Also it is known that every planar graph is representable as the intersection graph of line segments~\cite{chalopin2009every}.
-
-\martin{why do we need the reference here? Maybe better put it when it's actually used.}
-
-Using a standard sweep-line algorithm, computing the intersection graph $C$ of a set of $m$ line segments takes $O(m \log m + I)$ time~\cite{bcko-cgaa-08}, where $I$ is the number of intersections, i.e., the number of edges of $C$.
 
@@
 
-\section{Hardness of \maxsped for $k\ge3$}\label{sec:hardness}
+\section{Hardness of \maxsped for $k\ge3$}
 \fabian{Use the reduction from Till Bruckdorfers Dissertation \cite{b-sgh-15}.
@@
 
-Soeren?
+\soeren{comilation didn't work yet}
+
+To that \maxsped is \NP-hard even for 3-plane input drawings, we will provide a reduction from the \NP-hard positive planar 1-in-3SAT to \maxsped, similar to Till Bruckdorfers Dissertation \cite{b-sgh-15}.
+An incremental explanation of positive planar 1-in-3SAT would be: Given a formula $\phi$ in 3-CNF, in which every single occurring literal is not negated (positive).
+Consider the graph $P(\phi)$ that contains a vertex for each clause (clause vertex) and a vertex for each variable (variable vertex) and every variable vertex is connected via an edge with a clause vertex, if and only if the variable occurs in the corresponding clause.
+If $P(\phi)$ is planar, $\phi$ and a planar embedding $\Pi$ of $P(\phi)$ is an instance of positive planar 1-in-3SAT.
+The problem is to decide whether $\phi$ has a satisfying truth assignment, such that only one (and since it is a satisfying assignment, exactly one) variable in each clause is set to true.
+
+We create an instance of \maxsped out of an instance of positive planar 1-in-3SAT by creating:
+\begin{itemize}
+	\item a triangle of overlapping lines for every clause vertex
+	\item a circle of overlapping lines of even length for every variable vertex
+	\item a path of overlapping lines of even length for every connection between a variable vertex and a clause vertex in $\Pi$
+\end{itemize}
+\soeren{all these points need to be illustrated with suitable pictures}
+\soeren{TODO: explain construction in more detail, argue equivalence (both direction)}
+
+We construct our instance of positive planar 1-in-3SAT in the following fashion:
+\begin{itemize}
+	\item 
+\end{itemize}
 
@@
 %\end{itemize}
-
-Section~\ref{sec:hardness} showed that \maxsped is generally \NP-hard for $k \ge 3$.
-In this section we consider the special case that the intersection graph of the $k$-plane input drawing is a tree or a cactus.
-In both cases we present polynomial-time dynamic programming algorithms.
-
-Let $ C = (V,E) $ be the intersection graph of the edge set $S$ in a given drawing $\Gamma$ of graph $ G $ as defined in Section~\ref{sec:preliminaries}.
-Let $ u \in V $ and $ \deg(u) = k $, then for the corresponding edge $ s(u) \in S $ there are $ k + 1 $ possible stub pairs including drawing the whole edge.
-We store the lengths of the possible stubs as node weights $ \ell_1(u), \dots, \ell_k(u) \in \mathbb{R}_+$ and assume that $ \ell_1(u),\dots,\ell_k(u) $ are sorted from shorter to longer stubs.
-We define  $ \ell_0(u) $ as the length of the whole edge $ s(u) $.
-
-Computing the intersection graph requires to identify all the intersections in $ \Gamma $.
-Using a standard sweep line algorithm \cite{}\fabian{add citation} this requires $ O(|S|\log |S| + I) $ where $ I $ is the number of crossings.
-Building the graph then takes $ O(|S|) $ time.
-When computing the intersection we can compute the stub lengths and their order as well.
-
-\martin{time to build graph should be $O(S+I)$; but maybe this preparation work should go to preliminaries so that we can assume the intersection graph as input}
+Let $ C = (V,E) $ be the intersection graph of a graph $ G $ with edge set $ S $ and a drawing $ \Gamma $ of $ G $.
+Let $ u \in V $ with degree $ k $ and $ s \in S $ the corresponding edge in $ G $.
+There are $ k $ possible stub lengths $ \ell_1, \dots, \ell_k $.
+We assume $ \ell_1,\dots,\ell_k $ to be sorted from smaller to longer stubs.
+Additionally there is the possibility of drawing $ s $ as a complete segment which we set as $ \ell_0 $
 
 \subsection{Trees}
-Here we assume that the intersection graph $ C = (V,E) $ is a rooted tree of maximum degree $ k $.
-We give a bottom up dynamic programming algorithm for solving \maxsped on $ C $.
-For a vertex $ u \in V $ let $ p(u) $ denote the parent of $ u $ in $ C $ and let $ c(u) $ denote the set of its children.
-The ink values of the partial \maxsped solutions for the subtrees $T_u$ rooted at each vertex $ u \in V $ such that stub length $ \ell_i(u) $ is used for $s(u)$ are computed and stored as $ T_i(u) $.
+Here we assume $ C = (V,E) $ to be a tree of maximum degree $ k $.
+We give a bottom up dynamic program solving the \maxsped problem on $ C $.
+We store the partial solutions for each vertex $ u \in V $ and stub length $ \ell_i $ of $ u $ in $ T_i(u) $.
+If $ u $ is a leaf there are exactly two such values.
+Now let $ u $ be a vertex with degree $ k' \leq k$.
+There are exactly $ k' + 1 $ possibilities for the drawn stubs of $ u $, still of those $ k' + 1 $ possibilities only two are of interest to us.
 
-\begin{figure}[tbp]
+\begin{figure}
 	\centering
@@
 \end{figure}
-
-Since we know $ C $ is a tree every vertex $ u \in V $ with degree $ k' $ has a parent $ p(u) \in V $.
+Since we know $ C $ is a tree every vertex $ u \in V $ has a parent $ p(u) \in V $.
 For this parent there exists one stub length $ \ell_j(p(u)) $ in the sequence of stub lengths $ \ell_0(p(u)),\dots,\ell_{k'}(p(u)) $ such that drawing $ p(u) $ with these stubs does not intersect a completely drawn $ u $, while drawing $ p(u) $ with $ \ell_{j + 1}(p(u)) $, i.e.
@@
 We call the solution which would intersect the parent $ \sollong(u) $ and the one not intersecting the parent $ \solshort(u) $.
-See Figure~\ref{fig:two_solutions} for an illustration.
-More formally let $ j $ be the index such that stubs of length $ \ell_j(p(u)) $ do not intersect $ s(u) $, but $ \ell_{j +1}(p(u)) $ would, we set:
+More formally let $ j $ be the index such that stubs of length $ \ell_j(p(u)) $ do not intersect $ s(u) $, but $ \ell_{j +1}(p(u)) $ would.
+Then:
 \begin{align*}
@@
 
-\martin{-- maybe $\delta$ instead of $k'$\\
--- in the else case of $T_i$ long should be enough as always long $\ge$ short}
+Assuming the solutions for all children $ c(u) $ of $ u $ are computed we can calculate $ T_i(u) $ as:
 
-It remains to define $ T_i(u) $ for a vertex $ u $ and stub length $ \ell_i(u) $:
-
-\begin{align}
-	\label{rec:tree}
-	T_i(u) = \ell_i(u) + \sum_{v \in c(u)}
+\begin{align*}
+	T_i(u) = \ell_i + \sum_{v \in c(u)}
 	\begin{cases}
@@
 	\end{cases}
-\end{align}
-
-Solving Recurrence~\ref{rec:tree} yields an optimal solution to the \maxsped problem on $ G $ with drawing $ \Gamma $ as the following lemma shows.
-
-\begin{lemma}
-	\label{lem:tree_correctness}	
-	Let $ G $ be a simple graph and $ \Gamma $ a straight-line drawing of $ G $.
-If the intersection graph $ C = (V,E) $ of $ G $ is a tree with root $ r \in V $ and maximum degree $ k $, then $ \max \{T_i(r) \mid 0 \leq i \leq k \} $ is an optimal solution to the \maxsped problem on $ G $ with drawing $ \Gamma $.
-\end{lemma}
-
-\martin{could also use long(r) in the lemma}
-
-\begin{proof}
-	Let $ G $, $ \Gamma $ and $ C = (V,E) $ as above and $ t = \max \{T_i(r) \mid 0 \leq i \leq k \} $ is the maximum value for some $ T_i(r) $'s and $ r $ the root.
-Now let $ t' $ be the value of a solution to the \maxsped problem on $ G $ with drawing $ \Gamma $ and $ t' > t $.
-This solution has to be the sum of a set of stub lengths for each edge in $ G $ such that no two stubs intersect by definition.
-Further we know the stubs can be categorized into two sets relative to their parent.
-Recurrence~\ref{rec:tree} builds just the maximum over the $ \solshort(v) $ and $ \sollong(v) $ values for all $ v \in c(u) $, such that no two stubs intersect.
-Hence if $ t $ was not optimal we would need to find at least a third stub length for at least one child $ v \in c(u) $ which was neither equal to $ \sollong(v) $ nor $ \solshort(v) $.
-\end{proof}
-
-\martin{could also use induction on the tree structure. not sure what's better.}
-
-\begin{figure}
-	\centering
-	\includegraphics{speedup}
-	\caption{The situation in Lemma~\ref{lem:compute_vertex}.}
-	\label{fig:speedup}
-\end{figure}
-Solving Recurrence~\ref{rec:tree} straight forward would yield an algorithm running in cubic time depending on the number of vertices in $ C $.
-Using the order on the stub lengths though we  can improve this to a quadratic running time by solving $ T_i(u) $ in linear time for one $ u \in V $.
-
-\martin{the naïve solution would take $O(m\cdot k^2)$ and the improvement $O(mk)$, right?}
+\end{align*}
 
@@
 
-\martin{The lemma should state the time to compute not only the $T_i(u)$ values but also long and short}
-
-\begin{proof}
-	Let $ u \in V $ be a vertex with degree $ k' < k $ for which we have computed $ \sollong $ and $ \solshort $ for every of its children.
-The crucial part is that $ T_i(u) $ can be built iteratively by going through the list of stub lengths for $ u $ in ascending order.
-First we compute $ T_1(u) $, this is simply the solution for $ u $ in which we can draw all its children and the parent with their full segments.
-	\martin{we don't draw the parent yet; it's not included in $T_i(u)$. Moreover the children don't need to be all full segments, but they are unrestricted.}	
-	This means we need to build the sum of $ \ell_1(u) $ and the $ \sollong(v) $ for every $ v \in c(u) $ which runs in $ O(k') $.
-The same can be done for $ T_0(u) $, but taking all the $ \solshort(v) $ values for the children $ v \in c(u) $.
-Assume we computed $ T_j(u) $ for all stub lengths $ \ell_i(u) $ with $ i \leq j $.
-Now we want to compute $ T_{j+1}(u) $ (see Figure~\ref{fig:speedup}).
-By drawing $ s(u) $ with stub length $ \ell_{j +1} $ we make it a bit longer since we consider the $ \ell_i(u) $ in ascending order.
-Hence by doing so we now intersect the corresponding edge of the $ j $-th child $ v_j \in c(u) $.
-Since $ s(v_j) $ is the only segment affected by drawing $ u $ with stub length $ \ell_{j + 1} $ instead of $ \ell_j $ we can compute $ T_{j+1}(u) $ in $O(1)$ time by simply subtracting $ \sollong(v) $ and adding $ \solshort(v) $.
-\end{proof}
-
 \begin{theorem}
-	Let $ G $ be a simple graph with $m$ edges and $ \Gamma $ a straight-line drawing of $ G $.
-If the intersection graph $ C = (V,E) $ of $ G $ is a tree with maximum degree $ k \in \mathbb{N} $ the \maxsped problem can be solved in $ O(mk + m \log m)) $ time and space.
+	Let $ G $ be a simple graph and $ \Gamma $ a straight-line drawing of $ G $, if the intersection graph $ C = (V,E) $ of $ G $ is a tree with maximum degree $ k \in \mathbb{N} $ the \maxsped problem can be solved in $ O(kn) $ time and space.
 \end{theorem}
 \begin{proof}
-	Since $ C $ is known to be a tree, and hence $ |E| = O(m) $, we can compute the intersection graph and the stub lengths in $ O(m\log m) $ time and space.
+	The intersection graph $ C = (V,E) $ and the necessary weights $ \ell_i $ can be computed in quadratic time in number of vertices of $ G $.
 
-	To solve the \maxsped problem on $ G $ we need to compute $ \max \{T_i(r) \mid 0 \leq i \leq k \} $ for the root $ r \in V $ of $ C $ by Lemma~\ref{lem:tree_correctness}.
-Doing this bottom up requires one pass through $ C $ hence by applying Lemma~\ref{lem:compute_vertex} we need $ O(mk) $ time to compute all $ T_i(u) $ for all $ u \in V $.
-For every vertex we only need to compute $ O(k) $ values and store two of them, namely $ \sollong(u) $ and $ \solshort(u) $, so we do not need more than $ O(mk) $ memory.
-Adding the bounds yields the desired runtime of $ O(mk + m \log m)) $.
+	Let $ u \in V $ be a leaf, then computing $ T_i(u) $ takes constant time, since there are only two possible stub length, hence computing $ \sollong(u) $ and $ \solshort(u) $ takes also just constant time.
 
-	Retrieving the actual stub length for each edge in $ G $ needed to produce the SPED can be done by backtracking the subsolutions which requires at most $ O(mk) $ time.
+	Let $ u \in V $ be not a leaf, but for all children $ v \in c(u) $ we know the values $ \solshort(v) $ and $ \sollong(v) $ for every stub length $ \ell_i(v) $.
+By Lemma~\ref{lem:compute_vertex} we can compute $ \sollong(u) $ and $ \solshort(u) $ in $ O(k) $.
+Hence if $ r \in V $ is the root of the tree $ C $, computing $ T_i(r) $ for all $ i $ takes $ O(nk) $ time.
+
+	It remains to show that the maximum over all $ T_i(r) $, with $ r \in V $ being the root, is the optimal value over all solutions to the \maxsped problem on $ G $.
+Assume $ t $ is the maximum value for some $ T_i(r) $'s and $ r $ the root.
+Let $ t' $ be the value of another solution on $ G $ with drawing $ \Gamma $ and $ t' > t $.
 \end{proof}
-
-\martin{the figures can still be improved in style and size; maybe merge into one, use nicer colors, draw just one segment and not multiple shifted ones etc}
 

### 2018-01-07 eurocg/paper.tex
beta 4.2

--- 2018-01-07 eurocg/paper.tex
@@
 
-\soeren{comilation didn't work yet}
-
-To that \maxsped is \NP-hard even for 3-plane input drawings, we will provide a reduction from the \NP-hard positive planar 1-in-3SAT to \maxsped, similar to Till Bruckdorfers Dissertation \cite{b-sgh-15}.
-An incremental explanation of positive planar 1-in-3SAT would be: Given a formula $\phi$ in 3-CNF, in which every single occurring literal is not negated (positive).
-Consider the graph $P(\phi)$ that contains a vertex for each clause (clause vertex) and a vertex for each variable (variable vertex) and every variable vertex is connected via an edge with a clause vertex, if and only if the variable occurs in the corresponding clause.
-If $P(\phi)$ is planar, $\phi$ and a planar embedding $\Pi$ of $P(\phi)$ is an instance of positive planar 1-in-3SAT.
-The problem is to decide whether $\phi$ has a satisfying truth assignment, such that only one (and since it is a satisfying assignment, exactly one) variable in each clause is set to true.
-
-We create an instance of \maxsped out of an instance of positive planar 1-in-3SAT by creating:
-\begin{itemize}
-	\item a triangle of overlapping lines for every clause vertex
-	\item a circle of overlapping lines of even length for every variable vertex
-	\item a path of overlapping lines of even length for every connection between a variable vertex and a clause vertex in $\Pi$
-\end{itemize}
-\soeren{all these points need to be illustrated with suitable pictures}
-\soeren{TODO: explain construction in more detail, argue equivalence (both direction)}
-
-We construct our instance of positive planar 1-in-3SAT in the following fashion:
-\begin{itemize}
-	\item 
-\end{itemize}
+Soeren?
 
@@
 \end{lemma}
+\begin{proof}
+	\fabian{todo}
+\end{proof}
 

### 2018-01-07 eurocg/paper.tex
Hardness section is almost done but needs to be reviewed. I'm not completely sure about multiple parts.

--- 2018-01-07 eurocg/paper.tex
@@
 \usepackage{wrapfig}
+\usepackage{graphicx}
 
@@
 \newcommand{\maxsped}{\ensuremath{\textsc{MaxSPED}}\xspace}
+\newcommand{\ppsat}{\ensuremath{\textsc{Positive Planar 1-in-3 SAT}}\xspace}
 \newcommand{\sollong}{\ensuremath{\textit{long}}\xspace}
@@
 
-\soeren{comilation didn't work yet}
+\soeren{Should the graph in this section too be called G and its drawing $\Gamma$?
+Also,~\cite{b-sgh-15} claims, paths need to have not only even but equal length, which seems plausible and helps my argument.
+I'm however not sure if it is actually necessary.}
 
-To that \maxsped is \NP-hard even for 3-plane input drawings, we will provide a reduction from the \NP-hard positive planar 1-in-3SAT to \maxsped, similar to Till Bruckdorfers Dissertation \cite{b-sgh-15}.
-An incremental explanation of positive planar 1-in-3SAT would be: Given a formula $\phi$ in 3-CNF, in which every single occurring literal is not negated (positive).
+To show that \maxsped is \NP-hard even for 3-plane input drawings, we will provide a reduction from the \NP-hard \ppsat to \maxsped, similar to Till Bruckdorfers Dissertation \cite{b-sgh-15}.
+An incremental explanation of \ppsat would be: Given a formula $\phi$ in 3-CNF, in which every single occurring literal is not negated (positive).
 Consider the graph $P(\phi)$ that contains a vertex for each clause (clause vertex) and a vertex for each variable (variable vertex) and every variable vertex is connected via an edge with a clause vertex, if and only if the variable occurs in the corresponding clause.
-If $P(\phi)$ is planar, $\phi$ and a planar embedding $\Pi$ of $P(\phi)$ is an instance of positive planar 1-in-3SAT.
+If $P(\phi)$ is planar, $\phi$ and a planar embedding $\Pi$ of $P(\phi)$ is an instance of \ppsat.
 The problem is to decide whether $\phi$ has a satisfying truth assignment, such that only one (and since it is a satisfying assignment, exactly one) variable in each clause is set to true.
 
-We create an instance of \maxsped out of an instance of positive planar 1-in-3SAT by creating:
-\begin{itemize}
-	\item a triangle of overlapping lines for every clause vertex
-	\item a circle of overlapping lines of even length for every variable vertex
-	\item a path of overlapping lines of even length for every connection between a variable vertex and a clause vertex in $\Pi$
-\end{itemize}
-\soeren{all these points need to be illustrated with suitable pictures}
-\soeren{TODO: explain construction in more detail, argue equivalence (both direction)}
+\begin{figure}
+	\centering
+	\hfill
+	\includegraphics[width=.2\linewidth]{clause_to_triangle}
+	\hfill
+	\includegraphics[width=.5\linewidth]{variable_to_circle}
+	\caption{A triangle representing a clause, with unit lengths (a) and circles of segments representing variables, coloured in red and green (b).}
+	\label{fig:gadgets}
+\end{figure}
 
-We construct our instance of positive planar 1-in-3SAT in the following fashion:
-\begin{itemize}
-	\item 
-\end{itemize}
+We create a graph G which is an instance of \maxsped out of an instance of \ppsat by creating the following components.
+
+%\begin{itemize}
+%	\item a triangle of overlapping lines for every clause vertex (Fig.~\ref{fig:gadgets} (a))
+%	\item a circle of overlapping lines of even length for every variable vertex (Fig.~\ref{fig:gadgets} (b))
+%	\item a path of overlapping lines of even length for every connection between a variable vertex and a clause vertex in $\Pi$ (Fig.~\ref{fig:gadget_connection})
+%\end{itemize}
+
+A triangle (gadget clause) of overlapping lines $l_{x_a}$, $l_{x_b}$, and $l_{x_c}$ is constructed for each clause $(x_a \lor x_b \lor x_c)$ such that each edge of the trinagle is split into two stubs of length 1 and a middle part of length 2 (see Fig~\ref{fig:gadgets}(a)).
+Each edge of this clause gadget will be connected to exactly one variable gadget since every clause in $\phi$ consists of exactly three literals.
+Observe that only one edge at a time can be drawn completely since it intersects both other edges, that the optimal solution in terms of maximized ink is drawing one edge completely (4 units) and the stubs of the other two edges (2 units respectively), so 8 units in total and that all three solutions are equivalent.
+
+\begin{wrapfigure}[11]{r}{0.45\textwidth}
+	\centering
+	\includegraphics[width=.45\linewidth]{circle_true}
+	\hfill
+	\includegraphics[width=.45\linewidth]{circle_false}
+	\caption{Variable gadget representing a true variable (a) and a false variable (b)}
+	\label{fig:optimal_circles}
+\end{wrapfigure}
+
+Circles of segments (variable gadgets) are constructed in a very similar fashion.
+Let $m$ be the maximum number of clauses in which a single variable appears.
+Every variable $x_n$ will be represented by $m$ pairs of segments one red and one green which are connected as seen in Fig.~\ref{fig:gadgets} (b) and form a circle.
+It can be observed that every segment $s_n$ intersects exactly 2 other segments $s_{n-1}$ and $s_{n+1}$ ($s_{2m}$ intersects with $s_1$ to create the circle).
+These intersections split every segmet in three parts, which again have length 1 for the stubs and 2 for the middle part.
+If $x$ is true, the green edges will be drawn completely, if it is false the red edges will be drawn completely.
+Observe that fixing a green edge to be drawn completely fixes all green edges to be drawn completely, like in Fig.~\ref{fig:optimal_circles} (a) (in order to have an optimal solution for this circle) and conversely for the red edges, like in Fig.~\ref{fig:optimal_circles} (b).
+We will associate the first version, with subdivided green edges, with the variable being true and the second one with the variable being false.
+Since all edges are cut in the same fashion, the two solutions are equivalent in terms of maximized ink and amount to $4m + 2m = 6m$.
+
+The variable clauses are arranged on a spine and the clause gadgets are placed above or below the spine, according to the planar embedding of the \ppsat.
+Then the clause gadgets are connected to the variable gadgets corresponding to the occuring variables via segment chains (see Fig.\ref{fig:gadget_connection}, segment chains are for the sake of visualization not all equally long).
+These chains are composed of $t$ (where $t$ is an sufficiently large even integer) equal length segments $s_1$, \dots, $s_t$, which are again coloured alternatingly red and green.
+Every segment $s_i$ crosses with $s_{i-1}$ (except for $s_1$) and $s_{i+1}$ (except for $s_t$).
+All segments are divided again into 3 parts with the stubs having length 1, and the middle part having length 2.
+$s_1$ is also split in this ratio by one of the green stubs of a variable gadget, however the stub is intersected such that the remaining stub has length $1-\varepsilon$.
+The same holds for $s_t$ and a stub of an edge in the clause gadget.
+Here again the optimal solution in terms of maximizing ink is by drawing either all green edges completely and all red ones as stubs, or vice versa.
+Both solutions have a value of $4\frac{t}{2} + 2\frac{t}{2} = 3t$.
+It should be noted that completely drawing the red edges cuts of a small piece (of length $\varepsilon$) of the stubs of both the green edge of the variable gadget and the edge of the clause gadget, to which the chain is connected.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=.8\linewidth]{variable_clause_connection}
+	\caption{Connection between the clause and variable gadgets, created by chains of an even number of segments.}
+	\label{fig:gadget_connection}
+\end{figure}
+
+Let $q$ be the number of clauses and $p$ be the number of distinct variables in $\phi$.
+
+\begin{lemma}
+	\label{lem:optimal value}	 
+	The construcetd Graph G has an optimal  value in terms of maximizing ink of $8q + 6mp + 9qt - (4q + 2p)\varepsilon$ if $\varepsilon$ is chosen sufficiently small.
+\end{lemma}
+
+\begin{proof}
+	\soeren{i have an idea for this but i don't know if it would be out of scope, or if i might not even need this lemma.
+Also is our \maxsped defined as "Does this graph have a drawing with at least x amount of ink?" since thats what this bound is aiming at?"}
+\end{proof}
+\soeren{it also still needs to be shown that this bound ensures that the two criterias from~\cite{b-sgh-15}, namely that at least one edge in the triangle is drawn and that the segment chains are alternatingly drawn and not drawn.}
+
+\begin{theorem}
+	Computing a \maxsped for 3-planar graphs is \NP-hard.
+\end{theorem}
+
+\begin{proof}
+	Let $\phi$ be a boolean formula and $\Pi$ be a planar embedding corresponding to that formula sucht that ($\phi$, $\Pi$) is an instance of \ppsat.
+Let $\Gamma$ be a drawing constructed from that instance in the previously described way.
+We will show that there is a valid truth assignement for $\phi$ if and only if $\Gamma$ admits a \maxsped with a value of at least $8q + 6mp + 3q (3t - 4\varepsilon)$ in terms of maximizing ink.
+	~
+	\begin{enumerate}
+		\item[$\Rightarrow$:] Let $x$ be a variable and $c$ be a clause containing $x$.
+If $x$ is true draw all red edges in the associated variable gadget completely, which in turn means all green edges are drawn partially and vice versa for when $x$ is false.
+By drawing the green edges partially we enable the red edges in the segment chains to be drawn completely (if we shorten the stubs for one green edge in the variable gadget by $\varepsilon$).
+this also means we can draw the edges of the clause gadgets to which this segment chain is connected completely.
+Since $\phi$ has a satisfying variable assignement where exactly one variable per clause is true we have exactly one of these segment chains which allows exactly one edge per clause graph to be drawn completely.
+This results in every clause graph having the value $8 - 4\varepsilon$, every variable gadget having the value of $6m - 2\varepsilon$ and every segment chain having the value of $3t$.
+Let $q$ be the number of clauses and $p$ be the number of distinct variables in $\phi$.
+This results in
+
+		\soeren{is there a nice way to format this?}
+		$(8 - 4\varepsilon)q + (6m - 2\varepsilon)p + 3t*3q = $
+		$8q + 6mp + 9qt - (4q + 2p)\varepsilon$
+
+		\item[$\Leftarrow$:]  Assume $\phi$ does not have a stisfying variable assignement.
+Then there must always exist a clause in which all three contained variables $x_1$, $x_2$ and $x_3$ are false.
+That however means that, by construction, there is a clause gadget where all the corresponding variable gadgets have all their green edges completely drawn.
+This adds at most the value of $2\varepsilon$.
+In order to get the maximal drawing, this propagates through the segment chains up to the clause gadget where it causes all three edges of that gadget to be not drawn.
+This in turn means that this specific gadget has only the value $6 - 6\varepsilon$.
+The total value of this drawing in terms of maximizing ink is therefore at least $2$ smaller than normal.
+If we would try to supplement the missing drawn edge of the clause gadget, we would need to switch up the alternation of drawing segments in the chains, which causes 2 consecutive segments to not be drawn.
+Since the last edge in the chain was dran before, it now can't be drawn and it therefore again results in a difference of at least $2$ in the complete drawing.
+	\end{enumerate}
+\end{proof}
+
+\soeren{TODO: argue equivalence (both direction)}
 

### 2018-01-07 eurocg/paper.tex
Merge, If I'm right the proof was deleted and replaced by a todo (fabian?) but I'm not sure so i just commented it out

@@
 \begin{proof}
-	Let $ u \in V $ be a vertex with degree $ \delta < k $ for which we have computed $ \sollong $ and $ \solshort $ for every of its children.
+%<<<<<<< HEAD
+	\fabian{todo}
+%=======
+	% Let $ u \in V $ be a vertex with degree $ \delta < k $ for which we have computed $ \sollong $ and $ \solshort $ for every of its children.
 The crucial part is that $ T_i(u) $ can be built iteratively by going through the list of stub lengths for $ u $ in ascending order.
@@
 Since $ s(v_j) $ is the only segment affected by drawing $ u $ with stub length $ \ell_{j + 1} $ instead of $ \ell_j $ we can compute $ T_{j+1}(u) $ in $O(1)$ time by simply subtracting $ \sollong(v) $ and adding $ \solshort(v) $.
-
-	Finally we need to compute $ \solshort(u) $ and $ \sollong(u) $, but this are just two maximizations over at most $ \delta + 1 $ values at most.
+%	
+%	Finally we need to compute $ \solshort(u) $ and $ \sollong(u) $, but this are just two maximizations over at most $ \delta + 1 $ values at most.
 This can be done in $ O(\delta) $ by simply sweeping over the values.
+%>>>>>>> 579c12b7895efe3946892997d4f93b4ff0a1bff7
 \end{proof}

### 2018-01-07 eurocg/paper.tex

@@
 \newcommand{\sollong}{\ensuremath{\textit{long}}\xspace}
+\newcommand{\solmid}{\ensuremath{\textit{midt}}\xspace}
 \newcommand{\solshort}{\ensuremath{\textit{short}}\xspace}
@@
 
-\subsection{Cacti (of Degree Three)}
+\subsection{Cacti}
+In this section we restrict $ C = (V,E) $ to be a cactus graph, i.e.
+a simple graph in which no edge $ e \in E $ is part of more than one cycle.
+For a cactus graph $ C $ the \emph{block-cut tree} is the tree $ \mathcal T = (\mathcal V, \mathcal E) $where $ \mathcal V $ contains a vertex for every cycle and \emph{articulation}\fabian{define articulation in prelims?} vertex in $ C $ and an edge between a vertex in $ \mathcal V $ representing a cycle and one representing an articulation point if that articulation vertex is part of the cycle.
+\fabian{improve description, add reference}
 
-\begin{itemize}
-	\item seems to work for all cactus graphs
-\end{itemize}
+\fabian{We can solve \maxsped on cactus intersection graph iff every vertex in the block-cut tree representing a cycle carries the three optimal values depending on how the articulation point is drawn.}
+Let $ C = (V,E) $ be a cactus graph and moreover the intersection graph of a graph $ G $ and straight-line drawing $ \Gamma $.
+Now let $ \mathcal T = (\mathcal V, \mathcal E) $ be the block-cut tree of $ C $.
+Similar to Section~\ref{sec:trees} we define three values for every vertex in $ \mathcal V $.
+\begin{align*}
+	\solshort(u) = \\
+	\solmid(u) = \\
+	\sollong(u) = \\
+\end{align*}
 
+\fabian{For one cycle we use the same approach as in Tills work.
+Cut up the cycle at some edge, then the cylce edges lie all on one path.
+From this path there are edges, but using the BC-tree this cut open graph is a tree.
+Other way to argue this?}
 \section{Conclusion}

### 2018-01-08 eurocg/paper.tex
hardness proof polishing

@@
 \newcommand{\maxsped}{\ensuremath{\textsc{MaxSPED}}\xspace}
-\newcommand{\ppsat}{\ensuremath{\textsc{Positive Planar 1-in-3 SAT}}\xspace}
+\newcommand{\ppsat}{\ensuremath{\textsc{Planar 3-Sat}}\xspace}
 \newcommand{\sollong}{\ensuremath{\textit{long}}\xspace}
@@
 We call $ \Gamma $ \emph{$ k $-plane} if every edge $ s_i \in S $ is crossed by at most $ k $ other edges from $ S $ in $ \Gamma $.
-We use the terms edge in $ S $ and segment in $ \Gamma $ interchangeable, since there will always be a drawing $ \Gamma $ given with the graph $ G $.
+We use the terms edge in $ S $ and segment in $ \Gamma $ interchangeably, since our input is always a graph $G$ with a given drawing $ \Gamma $.
+Hence we can also interpret $S$ as a set of line segments.
 
@@
 \section{Hardness of \maxsped for $k\ge3$}\label{sec:hardness}
-\fabian{Use the reduction from Till Bruckdorfers Dissertation \cite{b-sgh-15}.
-The Clauses can be modelled as triangles, the Variables as Circles with every second segment connected to a path.
-Path should have even length, segments are of length four.}
-
 \soeren{Should the graph in this section too be called G and its drawing $\Gamma$?
@@
 
-To show that \maxsped is \NP-hard even for 3-plane input drawings, we will provide a reduction from the \NP-hard \ppsat to \maxsped, similar to Till Bruckdorfers Dissertation \cite{b-sgh-15}.
+In this section we close the gap between the known hardness of \maxsped~\cite{b-sgh-15} and the polynomial-time algorithm for 2-plane drawings~\cite{bcgkmn-pped-17}.
+
+\begin{theorem}\label{thm:hard}
+	The problem \maxsped is \NP-hard even for 3-plane graph drawings.
+\end{theorem}
+
+\begin{proof}
+	We reduce from the \NP-hard problem \ppsat~\cite{l-pftu-82} using similar ideas as Bruckdorfer's general proof of the hardness of \maxsped~\cite{b-sgh-15}. 
+	Let $\phi$ be a planar 3-Sat formula with $n$ variables $\{x_1, \dots, x_n\}$ and $m$ clauses $\{c_1, \dots, c_m\}$, each consisting of three literals.
+	We can assume that $\phi$ comes with a planar drawing of its variable-clause graph $H_\phi$ that has a vertex for each variable $x_i$ placed on a horizontal line and a vertex for each clause $c_j$ connected to the three variables appearing in $c_j$.
+	In our gadget-based reduction we create a 3-plane drawing $\Gamma_\phi$ as a set of line segments of uniform length and a value $L$ such that $\Gamma$ has a SPED of total ink at least $L$ if and only if $\phi$ is satisfiable.
+
+	All segments
+
+% To show that \maxsped is \NP-hard even for 3-plane input drawings, we will provide a reduction from the \NP-hard \ppsat to \maxsped, similar to Till Bruckdorfers Dissertation \cite{b-sgh-15}.
 An incremental explanation of \ppsat would be: Given a formula $\phi$ in 3-CNF, in which every single occurring literal is not negated (positive).
@@
 
-We create a graph G which is an instance of \maxsped out of an instance of \ppsat by creating the following components.
+% We create a graph G which is an instance of \maxsped out of an instance of \ppsat by creating the following components.
 
@@
 
-Let $q$ be the number of clauses and $p$ be the number of distinct variables in $\phi$.
-
-\begin{lemma}
-	\label{lem:optimal value}	 
-	The construcetd Graph G has an optimal  value in terms of maximizing ink of $8q + 6mp + 9qt - (4q + 2p)\varepsilon$ if $\varepsilon$ is chosen sufficiently small.
-\end{lemma}
-
-\begin{proof}
-	\soeren{i have an idea for this but i don't know if it would be out of scope, or if i might not even need this lemma.
+% Let $q$ be the number of clauses and $p$ be the number of distinct variables in $\phi$.
+%
+% \begin{lemma}
+% 	\label{lem:optimal value}
+% 	The construcetd Graph G has an optimal  value in terms of maximizing ink of $8q + 6mp + 9qt - (4q + 2p)\varepsilon$ if $\varepsilon$ is chosen sufficiently small.
+% \end{lemma}
+%
+% \begin{proof}
+% 	\soeren{i have an idea for this but i don't know if it would be out of scope, or if i might not even need this lemma.
 Also is our \maxsped defined as "Does this graph have a drawing with at least x amount of ink?" since thats what this bound is aiming at?"}
-\end{proof}
+% \end{proof}
 \soeren{it also still needs to be shown that this bound ensures that the two criterias from~\cite{b-sgh-15}, namely that at least one edge in the triangle is drawn and that the segment chains are alternatingly drawn and not drawn.}
 
-\begin{theorem}
-	Computing a \maxsped for 3-planar graphs is \NP-hard.
-\end{theorem}
-
-\begin{proof}
 	Let $\phi$ be a boolean formula and $\Pi$ be a planar embedding corresponding to that formula sucht that ($\phi$, $\Pi$) is an instance of \ppsat.
@@
 	\end{enumerate}
-\end{proof}
 
 \soeren{TODO: argue equivalence (both direction)}
+
+\end{proof}
 

### 2018-01-08 eurocg/paper.tex
sec 4.2 alpha

--- 2018-01-07 eurocg/paper.tex
@@
 \newcommand{\sollong}{\ensuremath{\textit{long}}\xspace}
-\newcommand{\solmid}{\ensuremath{\textit{midt}}\xspace}
+\newcommand{\solmid}{\ensuremath{\textit{mid}}\xspace}
 \newcommand{\solshort}{\ensuremath{\textit{short}}\xspace}
@@
 \subsection{Trees}
+\label{sec:tree}
 Here we assume that the intersection graph $ C = (V,E) $ is a rooted tree of maximum degree $ k $.
@@
 \subsection{Cacti}
+\label{sec:cactus}
 In this section we restrict $ C = (V,E) $ to be a cactus graph, i.e.
 a simple graph in which no edge $ e \in E $ is part of more than one cycle.
-For a cactus graph $ C $ the \emph{block-cut tree} is the tree $ \mathcal T = (\mathcal V, \mathcal E) $where $ \mathcal V $ contains a vertex for every cycle and \emph{articulation}\fabian{define articulation in prelims?} vertex in $ C $ and an edge between a vertex in $ \mathcal V $ representing a cycle and one representing an articulation point if that articulation vertex is part of the cycle.
-\fabian{improve description, add reference}
+For a cactus graph $ C $ the \emph{block-cut tree} is the tree $ \mathcal T = (\mathcal V, \mathcal E) $where $ \mathcal V $ contains a vertex for every cycle and \emph{cut vertex} in $ C $.
+We call a vertex $ u \in \mathcal V $ a \emph{block vertex} if it represents a cycle and \emph{articulation vertex} else.
+For a block vertex $ u \in \mathcal V $ representing an induced cycle with vertices $ V_c \subseteq V $ we define $ \textit{cycle}(u) = V_c$.
+For an articulation vertex $ u \in \mathcal V $ representing the cut vertex $ v \in V $ we define $ A(u) = v $.
+There is an edge $ uv \in \mathcal E $ between a block vertex $ u $ and articulation vertex $ v $ whenever $ v \in \textit{cycle}(u) $.
 
-\fabian{We can solve \maxsped on cactus intersection graph iff every vertex in the block-cut tree representing a cycle carries the three optimal values depending on how the articulation point is drawn.}
 Let $ C = (V,E) $ be a cactus graph and moreover the intersection graph of a graph $ G $ and straight-line drawing $ \Gamma $.
-Now let $ \mathcal T = (\mathcal V, \mathcal E) $ be the block-cut tree of $ C $.
-Similar to Section~\ref{sec:trees} we define three values for every vertex in $ \mathcal V $.
+Additionally let $ \mathcal T = (\mathcal V, \mathcal E) $ be the block-cut tree of $ C $ we define for each vertex $ u \in \mathcal V $ two values $ a,b \in \mathbb{R}_+ $, w.l.o.g $ a \leq b $.
+If $ u $ is an articulation vertex this are exactly the positions where the cycle represented by $ p(u) $ intersects $ s(A(u)) $ in $ \Gamma $.
+If $ u $ is a block vertex $ a,b $ are the points where the by $ \textit{cycle}(u) $ represented edges intersect $ s(A(p(u))) $.
+Similar to Section~\ref{sec:tree} we define three values for every vertex $ u \in \mathcal V $ with degree $ \delta $:
 \begin{align*}
-	\solshort(u) = \\
-	\solmid(u) = \\
-	\sollong(u) = \\
+	\solshort(u) &= \max\{T_0(u),\dots,T_\delta(u)\}\\
+	\solmid(u) &= \max\{T_1(u),\dots,T_{a}(u)\}\\
+	\sollong(u) &= \max\{T_1(u),\dots,T_{b}(u)\}\\
 \end{align*}
 
-\fabian{For one cycle we use the same approach as in Tills work.
-Cut up the cycle at some edge, then the cylce edges lie all on one path.
-From this path there are edges, but using the BC-tree this cut open graph is a tree.
-Other way to argue this?}
+For an articulation vertex $ u \in \mathcal V $:
+\begin{align}
+\label{rec:cactus_block}
+T_i(u) = \ell_i(A(u)) + \sum_{v \in c(u)}
+\begin{cases}
+\solshort(v) & \text{if } \ell_i(A(u)) \leq a \\
+\solmid(v) & \text{if } a < \ell_i(A(u)) \leq b \\
+\sollong(v) & \text{else}
+\end{cases}
+\end{align}
+For a block vertex $ u \in \mathcal V $ we have to cut open the cycle it represents (compare \cite{bcgkmn-pped-17}).
+We do this at $ p = A(p(u)) $.
+The resulting graph is a path over the vertices in $ \textit{cycle}(u) $ with attached articulation vertices.
+This is especially a tree and we use the aglorithm from Section~\ref{sec:tree}, maximizing over three values of the articulation vertices though,  to compute the $ T_i(u) $.
+
+\begin{lemma}
+	\label{lem:compute_vertex_cactus}
+	Let $ \mathcal T = (\mathcal V,\mathcal E) $ be the block-cut tree of a cactus graph $ C = (V,E) $ of maximum degree $ k $ and vertex weights $ \ell_0,\dots,\ell_{\delta} $ for every $ u \in V $ with degree $ \delta $.
+We can compute $ \solshort(u) $, $ \solmid(u) $ and $ \sollong(u) $ in $ O(\delta^2) $ time and space, if the values $ \solshort(v) $, $ \solmid(v) $ and $ \sollong(v) $ are computed for all children $ v \in c(u) $.
+\end{lemma}
+\begin{proof}
+	\fabian{todo}
+\end{proof}
+
+\begin{theorem}
+	Let $ G $ be a simple graph with $m$ edges and $ \Gamma $ a straight-line drawing of $ G $.
+If the intersection graph $ C = (V,E) $ of $ G $ is a cactus with maximum degree $ k \in \mathbb{N} $ the \maxsped problem can be solved in $ O(mk^2 + m \log m)) $ time and space.
+\end{theorem}
+\begin{proof}
+	\fabian{todo}
+\end{proof}
+
 \section{Conclusion}

### 2018-01-08 eurocg/paper.tex
sec 4.2 alpha

@@
 \end{lemma}
-
 \begin{proof}
@@
 \end{align}
-For a block vertex $ u \in \mathcal V $ we have to cut open the cycle it represents (compare \cite{bcgkmn-pped-17}).
+
+\begin{lemma}
+	\label{lem:cactus_correctness}	
+	Let $ G $ be a simple graph and $ \Gamma $ a straight-line drawing of $ G $.
+If the intersection graph $ C = (V,E) $ of $ G $ is a cactus with root $ r \in V $ and maximum degree $ k $, then $ \sollong(r) = \max \{T_i(r) \mid 0 \leq i \leq k \} $ is an optimal solution to the \maxsped problem on $ G $ with drawing $ \Gamma $.
+\end{lemma}
+\begin{proof}	
+	Let $ G $, $ \Gamma $,  $ C = (V,E) $ and $ \mathcal T = (\mathcal V,\mathcal E) $ as above and $ u \in \mathcal V $ a leaf of $ \mathcal T$.
+Then either $ u $ is an articulation vertex and to be treated as in Lemma~\ref{lem:compute_vertex} or a cycle.
+In the latter case we find the optimal solution to the \maxsped problem on $ C[\textit{cycle}(u)] $ with the algorithm from Bruckdorfer et al.
+\cite{bcgkmn-pped-17}.
+Using the above described method we derive $ \solshort(u) $, $ \solmid(u) $ and $ \sollong(u) $, this are exactly the optimal values depending on how the cycle interacts with the articulation vertex $ p(u) $.
+
+	Now let $ u \in V $ with degree $ \delta $ be an inner vertex of $ \mathcal T $.
+The values $ \solshort(v) $, $ \solmid(v) $ and $ \sollong(v) $ are computed for all $ v \in c(u) $.
+Again first let $ u $ be an articulation vertex, then the maximum for stub $ \ell_i(A(u)) $ is said value plus the sum of the compatible children, which is just Recurrence~\ref{rec:cactus_block}.
+Now let $ u $ be a block vertex.
+We have to cut open the cycle it represents (compare \cite{bcgkmn-pped-17}).
 We do this at $ p = A(p(u)) $.
@@
 This is especially a tree and we use the aglorithm from Section~\ref{sec:tree}, maximizing over three values of the articulation vertices though,  to compute the $ T_i(u) $.
+\end{proof}
 
 \begin{lemma}
-	\label{lem:compute_vertex_cactus}
-	Let $ \mathcal T = (\mathcal V,\mathcal E) $ be the block-cut tree of a cactus graph $ C = (V,E) $ of maximum degree $ k $ and vertex weights $ \ell_0,\dots,\ell_{\delta} $ for every $ u \in V $ with degree $ \delta $.
-We can compute $ \solshort(u) $, $ \solmid(u) $ and $ \sollong(u) $ in $ O(\delta^2) $ time and space, if the values $ \solshort(v) $, $ \solmid(v) $ and $ \sollong(v) $ are computed for all children $ v \in c(u) $.
+	\label{lem:compute_cactus}
+	Let $ \mathcal T = (\mathcal V,\mathcal E) $ be a cactus with maximum degree $ k $ and vertex weights $ \ell_0,\dots,\ell_{\delta} $ for every articulation vertex $ u \in \mathcal V $ with degree $ \delta $.
+We can compute $ \solshort(u) $, $ \solmid(u) $ and $ \sollong(u) $ in $ O(\delta) $ time and space, if the values $ \sollong(v) $, $ \solmid(v) $ and $ \solshort(v) $ are computed for all children $ v \in c(u) $.
 \end{lemma}

### 2018-01-08 eurocg/paper.tex
sec 4.2 alpha

@@
 \begin{theorem}
+	\label{thm:tree}
 	Let $ G $ be a simple graph with $m$ edges and $ \Gamma $ a straight-line drawing of $ G $.
@@
 For an articulation vertex $ u \in \mathcal V $ representing the cut vertex $ v \in V $ we define $ A(u) = v $.
-There is an edge $ uv \in \mathcal E $ between a block vertex $ u $ and articulation vertex $ v $ whenever $ v \in \textit{cycle}(u) $.
+There is an edge $ uv \in \mathcal E $ between a block vertex $ u $ and articulation vertex $ v $ whenever $ v \in \textit{cycle}(u) $ and an edge $ uv \in \mathcal E $ between two articulation vertices whenever $ A(u)A(v) $ was an edge in $ E $.
 
@@
 \begin{align}
-\label{rec:cactus_block}
+\label{rec:cactus_articulation}
 T_i(u) = \ell_i(A(u)) + \sum_{v \in c(u)}
@@
 \begin{lemma}
-	\label{lem:compute_cactus}
+	\label{lem:compute_articulation}
 	Let $ \mathcal T = (\mathcal V,\mathcal E) $ be a cactus with maximum degree $ k $ and vertex weights $ \ell_0,\dots,\ell_{\delta} $ for every articulation vertex $ u \in \mathcal V $ with degree $ \delta $.
@@
 \begin{proof}
-	\fabian{todo}
+	Follows from Lemma~\ref{lem:compute_vertex}, since an articulation vertex has at most $ \delta $ many children in $ \mathcal{T} $, one for each cycle it is part of in $ C $ and one for every other articulation vertex it was already connected to in $ C $.
+\end{proof}
+
+\begin{lemma}
+	\label{lem:compute_block}
+	Let $ \mathcal T = (\mathcal V,\mathcal E) $ be a cactus with maximum degree $ k $ and vertex weights $ \ell_0,\dots,\ell_{\delta} $ for every block vertex $ u \in \mathcal V $ with degree $ \delta $.
+We can compute $ \solshort(u) $, $ \solmid(u) $ and $ \sollong(u) $ in $ O(n\delta) $ time and space, if the values $ \sollong(v) $, $ \solmid(v) $ and $ \solshort(v) $ are computed for all children $ v \in c(u) $.
+\end{lemma}
+\begin{proof}
+	Let $ G $, $ \Gamma $,  $ C = (V,E) $ and $ \mathcal T = (\mathcal V,\mathcal E) $ as above and $ u \in \mathcal V $ a block vertex with degree $ \delta $.
+Computing the value for the cycle itself $ \textit{cycle}(u) $ takes $ O(n) $ time by Bruckdorfer et al.
+Now at every vertex $ v \in \textit{cycle}(u) $ there might be a set of up to $ \delta $ many children, which are all articulation points.
+Their values are already known and maximization over them can hence be done in $ O(\delta) $ time by the same argument as in Lemma~\ref{lem:compute_vertex}.
+Following we need $ O(n\delta) $ to compute the three values $ \solshort(u) $, $ \solmid(u) $ and $ \sollong(u) $.
 \end{proof}
@@
 \begin{proof}
-	\fabian{todo}
+	Follows as in Theorem~\ref{thm:tree} by maximizing over the tree using Lemma~\ref{lem:cactus_correctness} and Recurrence~\ref{rec:cactus_articulation}.
+Then the running time follows from constructing the intersection graph and the block-cut tree as well as Lemma~\ref{lem:compute_articulation} and~\ref{lem:compute_block}.
 \end{proof}

### 2018-01-08 eurocg/paper.tex
zwischenstand

--- 2018-01-08 eurocg/paper.tex
@@
 \newcommand{\sollong}{\ensuremath{\textit{long}}\xspace}
-\newcommand{\solmid}{\ensuremath{\textit{mid}}\xspace}
+\newcommand{\solmid}{\ensuremath{\textit{midt}}\xspace}
 \newcommand{\solshort}{\ensuremath{\textit{short}}\xspace}
@@
 
-In this section we close the gap between the known hardness of \maxsped~\cite{b-sgh-15} and the polynomial-time algorithm for 2-plane drawings~\cite{bcgkmn-pped-17}.
+In this section we close the gap between the known hardness of \maxsped~\cite{b-sgh-15} and the polynomial-time algorithm for 2-plane drawings~\cite{bcgkmn-pped-17} as summarized in the following theorem.
 
@@
 \begin{proof}
-	We reduce from the \NP-hard problem \ppsat~\cite{l-pftu-82} using similar ideas as Bruckdorfer's general proof of the hardness of \maxsped~\cite{b-sgh-15}. 
+	We reduce from the \NP-hard problem \ppsat~\cite{l-pftu-82} using similar ideas as Bruckdorfer's~\cite{b-sgh-15} general proof of the hardness of \maxsped.
+%, which reduces from the more restricted problem \textsc{Planar Positive 1-in-3 Sat}.
 	Let $\phi$ be a planar 3-Sat formula with $n$ variables $\{x_1, \dots, x_n\}$ and $m$ clauses $\{c_1, \dots, c_m\}$, each consisting of three literals.
-	We can assume that $\phi$ comes with a planar drawing of its variable-clause graph $H_\phi$ that has a vertex for each variable $x_i$ placed on a horizontal line and a vertex for each clause $c_j$ connected to the three variables appearing in $c_j$.
+	We can assume that $\phi$ comes with a planar drawing of its variable-clause graph $H_\phi$ that has a vertex for each variable $x_i$ placed on a horizontal line and a vertex for each clause $c_j$ connected on one side of the horizontal line to the three variables appearing in $c_j$.
 	In our gadget-based reduction we create a 3-plane drawing $\Gamma_\phi$ as a set of line segments of uniform length and a value $L$ such that $\Gamma$ has a SPED of total ink at least $L$ if and only if $\phi$ is satisfiable.
 
-	All segments
+	All segments are of length 5.
+	We construct our gadgets using pairs of intersecting segments, one colored green, the other one red.
+	Their mutual intersection point is at distance 1 from an endpoint, which has the effect that the maximum amount of ink contributed by such a pair is 7 (one full segment and the other one with two stubs of length 1 each).
 
@@
 \subsection{Trees}
-\label{sec:tree}
 Here we assume that the intersection graph $ C = (V,E) $ is a rooted tree of maximum degree $ k $.
@@
 \subsection{Cacti}
-\label{sec:cactus}
 In this section we restrict $ C = (V,E) $ to be a cactus graph, i.e.
 a simple graph in which no edge $ e \in E $ is part of more than one cycle.
-For a cactus graph $ C $ the \emph{block-cut tree} is the tree $ \mathcal T = (\mathcal V, \mathcal E) $where $ \mathcal V $ contains a vertex for every cycle and \emph{cut vertex} in $ C $.
-We call a vertex $ u \in \mathcal V $ a \emph{block vertex} if it represents a cycle and \emph{articulation vertex} else.
-For a block vertex $ u \in \mathcal V $ representing an induced cycle with vertices $ V_c \subseteq V $ we define $ \textit{cycle}(u) = V_c$.
-For an articulation vertex $ u \in \mathcal V $ representing the cut vertex $ v \in V $ we define $ A(u) = v $.
-There is an edge $ uv \in \mathcal E $ between a block vertex $ u $ and articulation vertex $ v $ whenever $ v \in \textit{cycle}(u) $.
+For a cactus graph $ C $ the \emph{block-cut tree} is the tree $ \mathcal T = (\mathcal V, \mathcal E) $where $ \mathcal V $ contains a vertex for every cycle and \emph{articulation}\fabian{define articulation in prelims?} vertex in $ C $ and an edge between a vertex in $ \mathcal V $ representing a cycle and one representing an articulation point if that articulation vertex is part of the cycle.
+\fabian{improve description, add reference}
 
+\fabian{We can solve \maxsped on cactus intersection graph iff every vertex in the block-cut tree representing a cycle carries the three optimal values depending on how the articulation point is drawn.}
 Let $ C = (V,E) $ be a cactus graph and moreover the intersection graph of a graph $ G $ and straight-line drawing $ \Gamma $.
-Additionally let $ \mathcal T = (\mathcal V, \mathcal E) $ be the block-cut tree of $ C $ we define for each vertex $ u \in \mathcal V $ two values $ a,b \in \mathbb{R}_+ $, w.l.o.g $ a \leq b $.
-If $ u $ is an articulation vertex this are exactly the positions where the cycle represented by $ p(u) $ intersects $ s(A(u)) $ in $ \Gamma $.
-If $ u $ is a block vertex $ a,b $ are the points where the by $ \textit{cycle}(u) $ represented edges intersect $ s(A(p(u))) $.
-Similar to Section~\ref{sec:tree} we define three values for every vertex $ u \in \mathcal V $ with degree $ \delta $:
+Now let $ \mathcal T = (\mathcal V, \mathcal E) $ be the block-cut tree of $ C $.
+Similar to Section~\ref{sec:trees} we define three values for every vertex in $ \mathcal V $.
 \begin{align*}
-	\solshort(u) &= \max\{T_0(u),\dots,T_\delta(u)\}\\
-	\solmid(u) &= \max\{T_1(u),\dots,T_{a}(u)\}\\
-	\sollong(u) &= \max\{T_1(u),\dots,T_{b}(u)\}\\
+	\solshort(u) = \\
+	\solmid(u) = \\
+	\sollong(u) = \\
 \end{align*}
 
-For an articulation vertex $ u \in \mathcal V $:
-\begin{align}
-\label{rec:cactus_block}
-T_i(u) = \ell_i(A(u)) + \sum_{v \in c(u)}
-\begin{cases}
-\solshort(v) & \text{if } \ell_i(A(u)) \leq a \\
-\solmid(v) & \text{if } a < \ell_i(A(u)) \leq b \\
-\sollong(v) & \text{else}
-\end{cases}
-\end{align}
-For a block vertex $ u \in \mathcal V $ we have to cut open the cycle it represents (compare \cite{bcgkmn-pped-17}).
-We do this at $ p = A(p(u)) $.
-The resulting graph is a path over the vertices in $ \textit{cycle}(u) $ with attached articulation vertices.
-This is especially a tree and we use the aglorithm from Section~\ref{sec:tree}, maximizing over three values of the articulation vertices though,  to compute the $ T_i(u) $.
-
-\begin{lemma}
-	\label{lem:compute_vertex_cactus}
-	Let $ \mathcal T = (\mathcal V,\mathcal E) $ be the block-cut tree of a cactus graph $ C = (V,E) $ of maximum degree $ k $ and vertex weights $ \ell_0,\dots,\ell_{\delta} $ for every $ u \in V $ with degree $ \delta $.
-We can compute $ \solshort(u) $, $ \solmid(u) $ and $ \sollong(u) $ in $ O(\delta^2) $ time and space, if the values $ \solshort(v) $, $ \solmid(v) $ and $ \sollong(v) $ are computed for all children $ v \in c(u) $.
-\end{lemma}
-\begin{proof}
-	\fabian{todo}
-\end{proof}
-
-\begin{theorem}
-	Let $ G $ be a simple graph with $m$ edges and $ \Gamma $ a straight-line drawing of $ G $.
-If the intersection graph $ C = (V,E) $ of $ G $ is a cactus with maximum degree $ k \in \mathbb{N} $ the \maxsped problem can be solved in $ O(mk^2 + m \log m)) $ time and space.
-\end{theorem}
-\begin{proof}
-	\fabian{todo}
-\end{proof}
-
+\fabian{For one cycle we use the same approach as in Tills work.
+Cut up the cycle at some edge, then the cylce edges lie all on one path.
+From this path there are edges, but using the BC-tree this cut open graph is a tree.
+Other way to argue this?}
 \section{Conclusion}

### 2018-01-08 eurocg/paper.tex
NP Beweis gestrafft

--- 2018-01-08 eurocg/paper.tex
@@
 Observe that the intersection graph $ C $ of a $ k $-plane drawing $ \Gamma $ has maximum degree $ k $.
+%
+Using a standard sweep-line algorithm, computing the intersection graph $C$ of a set of $m$ line segments takes $O(m \log m + |E|)$ time~\cite{bcko-cgaa-08}, where $|E|$ is  the number of intersections.
 
-Using a standard sweep-line algorithm, computing the intersection graph $C$ of a set of $m$ line segments takes $O(m \log m + I)$ time~\cite{bcko-cgaa-08}, where $I$ is the number of intersections, i.e., the number of edges of $C$.
+A \emph{symmetric partial edge drawing} (SPED) $D$ of $\Gamma$ draws a fraction $0 < f_s \le 1$ of each edge $s = uv \in S$ by drawing two symmetric edge stubs at $u$ and $v$ of length $f_s \cdot |s|/2$ each.
+No two stubs in $D$ may intersect.
+The \emph{ink} or \emph{ink value}  $I(D)$ of a SPED $D$ is the total stub length $I(D) = \sum_{s \in S} f_s |s|$.
+In the problem \maxsped the task is to find for a given drawing $\Gamma$ a SPED $D^*$ such that its ink $I(D^*)$ is maximum over all SPEDs.
 
@@
 \section{Hardness of \maxsped for $k\ge3$}\label{sec:hardness}
-\soeren{Should the graph in this section too be called G and its drawing $\Gamma$?
-Also,~\cite{b-sgh-15} claims, paths need to have not only even but equal length, which seems plausible and helps my argument.
-I'm however not sure if it is actually necessary.}
 
-In this section we close the gap between the known hardness of \maxsped~\cite{b-sgh-15} and the polynomial-time algorithm for 2-plane drawings~\cite{bcgkmn-pped-17}.
+In this section we close the gap between the known hardness of \maxsped~\cite{b-sgh-15} and the polynomial-time algorithm for 2-plane drawings~\cite{bcgkmn-pped-17} as stated in the following theorem.
 
@@
 \begin{proof}
-	We reduce from the \NP-hard problem \ppsat~\cite{l-pftu-82} using similar ideas as Bruckdorfer's general proof of the hardness of \maxsped~\cite{b-sgh-15}. 
+	We reduce from the \NP-hard problem \ppsat~\cite{l-pftu-82} using similar ideas as in Bruckdorfer's general proof of the hardness of \maxsped~\cite{b-sgh-15}.
+%, which reduces from the more restricted problem \textsc{Planar Positive 1-in-3 Sat}.
 	Let $\phi$ be a planar 3-Sat formula with $n$ variables $\{x_1, \dots, x_n\}$ and $m$ clauses $\{c_1, \dots, c_m\}$, each consisting of three literals.
-	We can assume that $\phi$ comes with a planar drawing of its variable-clause graph $H_\phi$ that has a vertex for each variable $x_i$ placed on a horizontal line and a vertex for each clause $c_j$ connected to the three variables appearing in $c_j$.
-	In our gadget-based reduction we create a 3-plane drawing $\Gamma_\phi$ as a set of line segments of uniform length and a value $L$ such that $\Gamma$ has a SPED of total ink at least $L$ if and only if $\phi$ is satisfiable.
-
-	All segments
-
-% To show that \maxsped is \NP-hard even for 3-plane input drawings, we will provide a reduction from the \NP-hard \ppsat to \maxsped, similar to Till Bruckdorfers Dissertation \cite{b-sgh-15}.
-An incremental explanation of \ppsat would be: Given a formula $\phi$ in 3-CNF, in which every single occurring literal is not negated (positive).
-Consider the graph $P(\phi)$ that contains a vertex for each clause (clause vertex) and a vertex for each variable (variable vertex) and every variable vertex is connected via an edge with a clause vertex, if and only if the variable occurs in the corresponding clause.
-If $P(\phi)$ is planar, $\phi$ and a planar embedding $\Pi$ of $P(\phi)$ is an instance of \ppsat.
-The problem is to decide whether $\phi$ has a satisfying truth assignment, such that only one (and since it is a satisfying assignment, exactly one) variable in each clause is set to true.
+	We can assume that $\phi$ comes with a planar drawing of its variable-clause graph $H_\phi$ that has a vertex for each variable $x_i$ placed on a horizontal line and a vertex for each clause $c_j$ either above or below the horizontal line and connected on that side of the line to the three variables appearing in $c_j$.
+	In our gadget-based reduction we create a 3-plane drawing $\Gamma_\phi$ as a set of line segments of uniform length and a value $L$ such that $\Gamma$ has a SPED with ink at least $L$ if and only if $\phi$ is satisfiable.
+	Figure~\ref{fig:gadget_connection} shows a sketch of the gadgets.
 
@@
 		\centering
-	\hfill
-	\includegraphics[width=.2\linewidth]{clause_to_triangle}
-	\hfill
-	\includegraphics[width=.5\linewidth]{variable_to_circle}
-	\caption{A triangle representing a clause, with unit lengths (a) and circles of segments representing variables, coloured in red and green (b).}
-	\label{fig:gadgets}
-\end{figure}
-
-% We create a graph G which is an instance of \maxsped out of an instance of \ppsat by creating the following components.
-
-%\begin{itemize}
-%	\item a triangle of overlapping lines for every clause vertex (Fig.~\ref{fig:gadgets} (a))
-%	\item a circle of overlapping lines of even length for every variable vertex (Fig.~\ref{fig:gadgets} (b))
-%	\item a path of overlapping lines of even length for every connection between a variable vertex and a clause vertex in $\Pi$ (Fig.~\ref{fig:gadget_connection})
-%\end{itemize}
-
-A triangle (gadget clause) of overlapping lines $l_{x_a}$, $l_{x_b}$, and $l_{x_c}$ is constructed for each clause $(x_a \lor x_b \lor x_c)$ such that each edge of the trinagle is split into two stubs of length 1 and a middle part of length 2 (see Fig~\ref{fig:gadgets}(a)).
-Each edge of this clause gadget will be connected to exactly one variable gadget since every clause in $\phi$ consists of exactly three literals.
-Observe that only one edge at a time can be drawn completely since it intersects both other edges, that the optimal solution in terms of maximized ink is drawing one edge completely (4 units) and the stubs of the other two edges (2 units respectively), so 8 units in total and that all three solutions are equivalent.
-
-\begin{wrapfigure}[11]{r}{0.45\textwidth}
-	\centering
-	\includegraphics[width=.45\linewidth]{circle_true}
-	\hfill
-	\includegraphics[width=.45\linewidth]{circle_false}
-	\caption{Variable gadget representing a true variable (a) and a false variable (b)}
-	\label{fig:optimal_circles}
-\end{wrapfigure}
-
-Circles of segments (variable gadgets) are constructed in a very similar fashion.
-Let $m$ be the maximum number of clauses in which a single variable appears.
-Every variable $x_n$ will be represented by $m$ pairs of segments one red and one green which are connected as seen in Fig.~\ref{fig:gadgets} (b) and form a circle.
-It can be observed that every segment $s_n$ intersects exactly 2 other segments $s_{n-1}$ and $s_{n+1}$ ($s_{2m}$ intersects with $s_1$ to create the circle).
-These intersections split every segmet in three parts, which again have length 1 for the stubs and 2 for the middle part.
-If $x$ is true, the green edges will be drawn completely, if it is false the red edges will be drawn completely.
-Observe that fixing a green edge to be drawn completely fixes all green edges to be drawn completely, like in Fig.~\ref{fig:optimal_circles} (a) (in order to have an optimal solution for this circle) and conversely for the red edges, like in Fig.~\ref{fig:optimal_circles} (b).
-We will associate the first version, with subdivided green edges, with the variable being true and the second one with the variable being false.
-Since all edges are cut in the same fashion, the two solutions are equivalent in terms of maximized ink and amount to $4m + 2m = 6m$.
-
-The variable clauses are arranged on a spine and the clause gadgets are placed above or below the spine, according to the planar embedding of the \ppsat.
-Then the clause gadgets are connected to the variable gadgets corresponding to the occuring variables via segment chains (see Fig.\ref{fig:gadget_connection}, segment chains are for the sake of visualization not all equally long).
-These chains are composed of $t$ (where $t$ is an sufficiently large even integer) equal length segments $s_1$, \dots, $s_t$, which are again coloured alternatingly red and green.
-Every segment $s_i$ crosses with $s_{i-1}$ (except for $s_1$) and $s_{i+1}$ (except for $s_t$).
-All segments are divided again into 3 parts with the stubs having length 1, and the middle part having length 2.
-$s_1$ is also split in this ratio by one of the green stubs of a variable gadget, however the stub is intersected such that the remaining stub has length $1-\varepsilon$.
-The same holds for $s_t$ and a stub of an edge in the clause gadget.
-Here again the optimal solution in terms of maximizing ink is by drawing either all green edges completely and all red ones as stubs, or vice versa.
-Both solutions have a value of $4\frac{t}{2} + 2\frac{t}{2} = 3t$.
-It should be noted that completely drawing the red edges cuts of a small piece (of length $\varepsilon$) of the stubs of both the green edge of the variable gadget and the edge of the clause gadget, to which the chain is connected.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.8\linewidth]{variable_clause_connection}
+		\includegraphics[width=.5\linewidth]{variable_clause_connection}
 		\caption{Connection between the clause and variable gadgets, created by chains of an even number of segments.}
@@
 
-% Let $q$ be the number of clauses and $p$ be the number of distinct variables in $\phi$.
-%
-% \begin{lemma}
-% 	\label{lem:optimal value}
-% 	The construcetd Graph G has an optimal  value in terms of maximizing ink of $8q + 6mp + 9qt - (4q + 2p)\varepsilon$ if $\varepsilon$ is chosen sufficiently small.
-% \end{lemma}
-%
-% \begin{proof}
-% 	\soeren{i have an idea for this but i don't know if it would be out of scope, or if i might not even need this lemma.
-Also is our \maxsped defined as "Does this graph have a drawing with at least x amount of ink?" since thats what this bound is aiming at?"}
-% \end{proof}
-\soeren{it also still needs to be shown that this bound ensures that the two criterias from~\cite{b-sgh-15}, namely that at least one edge in the triangle is drawn and that the segment chains are alternatingly drawn and not drawn.}
+	All segments in the gadgets are of length 5.
+	We use pairs of intersecting segments, alternatingly colored red and green.
+	The intersection point of each red-green segment pair is at distance 1 from an endpoint.
+	Thus the maximum amount of ink contributed by such a pair is 7 (one full segment of length 5 and the other one with two stubs of length 1 each).
 
-	Let $\phi$ be a boolean formula and $\Pi$ be a planar embedding corresponding to that formula sucht that ($\phi$, $\Pi$) is an instance of \ppsat.
-Let $\Gamma$ be a drawing constructed from that instance in the previously described way.
-We will show that there is a valid truth assignement for $\phi$ if and only if $\Gamma$ admits a \maxsped with a value of at least $8q + 6mp + 3q (3t - 4\varepsilon)$ in terms of maximizing ink.
-	~
-	\begin{enumerate}
-		\item[$\Rightarrow$:] Let $x$ be a variable and $c$ be a clause containing $x$.
-If $x$ is true draw all red edges in the associated variable gadget completely, which in turn means all green edges are drawn partially and vice versa for when $x$ is false.
-By drawing the green edges partially we enable the red edges in the segment chains to be drawn completely (if we shorten the stubs for one green edge in the variable gadget by $\varepsilon$).
-this also means we can draw the edges of the clause gadgets to which this segment chain is connected completely.
-Since $\phi$ has a satisfying variable assignement where exactly one variable per clause is true we have exactly one of these segment chains which allows exactly one edge per clause graph to be drawn completely.
-This results in every clause graph having the value $8 - 4\varepsilon$, every variable gadget having the value of $6m - 2\varepsilon$ and every segment chain having the value of $3t$.
-Let $q$ be the number of clauses and $p$ be the number of distinct variables in $\phi$.
-This results in
+	Each variable gadget is a cycle of segment pairs, with at least one pair for each occurence of the variable in $\phi$, see the three variable gadgets in the bottom of Fig.~\ref{fig:gadget_connection}.
+	Observe that this cycle has exactly two ink-maximal SPEDs: either all red edges are full segments and all green edges are length-1 stubs or vice versa.
+	We associate the configuration with full red segments and green stubs with the value \emph{true} and the configuration with full green segments and red stubs with the value \emph{false}.
 
-		\soeren{is there a nice way to format this?}
-		$(8 - 4\varepsilon)q + (6m - 2\varepsilon)p + 3t*3q = $
-		$8q + 6mp + 9qt - (4q + 2p)\varepsilon$
+	For each clause we construct a triple of mutually intersecting segments, see the black triangle at the top of Fig.~\ref{fig:gadget_connection}.
+	Again, their intersection points are at distance 1 from the endpoints.
+	It is clear that in such a clause triangle at most one of the three segments can be fully drawn, while the stubs of the other two can have length at most 1.
+	Hence the maximum amount of ink in a SPED contributed by a clause gadget is 9.
 
-		\item[$\Leftarrow$:]  Assume $\phi$ does not have a stisfying variable assignement.
-Then there must always exist a clause in which all three contained variables $x_1$, $x_2$ and $x_3$ are false.
-That however means that, by construction, there is a clause gadget where all the corresponding variable gadgets have all their green edges completely drawn.
-This adds at most the value of $2\varepsilon$.
-In order to get the maximal drawing, this propagates through the segment chains up to the clause gadget where it causes all three edges of that gadget to be not drawn.
-This in turn means that this specific gadget has only the value $6 - 6\varepsilon$.
-The total value of this drawing in terms of maximizing ink is therefore at least $2$ smaller than normal.
-If we would try to supplement the missing drawn edge of the clause gadget, we would need to switch up the alternation of drawing segments in the chains, which causes 2 consecutive segments to not be drawn.
-Since the last edge in the chain was dran before, it now can't be drawn and it therefore again results in a difference of at least $2$ in the complete drawing.
-	\end{enumerate}
+	Finally, we connect variable and clause gadgets in such a way that a clause gadget can contribute its maximum ink value of 9 if and only if the clause is satisfied by the selected truth assignment to the variables.
+	For a positive (negative) literal, we create a path of even length between a green (red) edge of the variable gadget and one of the three edges of the clause gadget as shown in Fig.~\ref{fig:gadget_connection}.
+	The first edge $s$ of this path intersects the corresponding variable edge $s'$ such that $s'$ is split into a piece of length 2 and a piece of length 3, whereas $s$ is split into a piece of length 1 and a piece of length 4.
+	The last edge of the path intersects the corresponding clause edge with the same length ratios.
+	The path itself consists of a chain of red-green segment pairs, where, as before, each pair can contribute an ink value of at most~7.
 
-\soeren{TODO: argue equivalence (both direction)}
+	Since the drawing of $H_\phi$ has polynomial size, the number of segments created in the reduction is polynomial. 
+	Further, no segment intersects more than three other segments, so the constructed drawing is 3-plane.
 
+	For the correctness of the reduction, let $L$ be the ink value obtained by counting 7 for each red-green segment pair and 9 for each clause gadget. 
+	First assume that $\phi$ has a satisfying truth assignment and put each variable gadget in its corresponding state.
+	For each clause, select exactly one literal with value \emph{true} in the satisfying truth assignment. 
+	We draw the clause segment that connects to the selected literal as a full segment and the other two as length-1 stubs.
+	Since the last segment of the selected literal path must be drawn as length-1 stubs, the only way of having a maximum contribution of that path is by alternating stubs and full segments.
+	Hence the first segment of the path must be a full segment.
+	But because the variable is in the state that sets the literal to \emph{true}, the intersecting variable segment is drawn as two stubs and the path configuration is valid.
+	For the two non-selected literals, we can draw the last segments of their paths as full segments, as well as every segment at an even position, while the segments at odd positions are drawn as stubs.
+	This is compatible with any of the two variable configurations and proves that we can indeed achieve ink value $L$.
+
+	Conversely, assume that we have a SPED with ink $L$. 
+	By construction, every red-green segment pair and all clause gadgets must contribute their respective maximum ink value.
+	In particular, each variable gadget is either in state \emph{true} or \emph{false}.
+	We claim that this truth assignment satisfies $\phi$.
+	For each clause, one segment must have full length and as argued before the path connecting that segment to a variable gadget must have all segments at even positions as stubs and all segments at odd positions as full segments.
+	In particular the first segment of the literal path is fully drawn and therefore in a SPED with ink $L$ the intersecting segment of the variable cannot be fully drawn.
+	In fact, it must be drawn as length-1 stubs for otherwise the variable gadget would not contribute its maximum ink value.
+	But then by construction the selected variable value sets our literal to \emph{true}.
 \end{proof}
@@
 \end{document}
+
+%%%% OLD TEXT PARTS %%%%%
+
+\begin{proof}
+
+% To show that \maxsped is \NP-hard even for 3-plane input drawings, we will provide a reduction from the \NP-hard \ppsat to \maxsped, similar to Till Bruckdorfers Dissertation \cite{b-sgh-15}.
+An incremental explanation of \ppsat would be: Given a formula $\phi$ in 3-CNF, in which every single occurring literal is not negated (positive).
+Consider the graph $P(\phi)$ that contains a vertex for each clause (clause vertex) and a vertex for each variable (variable vertex) and every variable vertex is connected via an edge with a clause vertex, if and only if the variable occurs in the corresponding clause.
+If $P(\phi)$ is planar, $\phi$ and a planar embedding $\Pi$ of $P(\phi)$ is an instance of \ppsat.
+The problem is to decide whether $\phi$ has a satisfying truth assignment, such that only one (and since it is a satisfying assignment, exactly one) variable in each clause is set to true.
+
+\begin{figure}
+	\centering
+	\hfill
+	\includegraphics[width=.2\linewidth]{clause_to_triangle}
+	\hfill
+	\includegraphics[width=.5\linewidth]{variable_to_circle}
+	\caption{A triangle representing a clause, with unit lengths (a) and circles of segments representing variables, coloured in red and green (b).}
+	\label{fig:gadgets}
+\end{figure}
+
+% We create a graph G which is an instance of \maxsped out of an instance of \ppsat by creating the following components.
+
+%\begin{itemize}
+%	\item a triangle of overlapping lines for every clause vertex (Fig.~\ref{fig:gadgets} (a))
+%	\item a circle of overlapping lines of even length for every variable vertex (Fig.~\ref{fig:gadgets} (b))
+%	\item a path of overlapping lines of even length for every connection between a variable vertex and a clause vertex in $\Pi$ (Fig.~\ref{fig:gadget_connection})
+%\end{itemize}
+
+A triangle (gadget clause) of overlapping lines $l_{x_a}$, $l_{x_b}$, and $l_{x_c}$ is constructed for each clause $(x_a \lor x_b \lor x_c)$ such that each edge of the trinagle is split into two stubs of length 1 and a middle part of length 2 (see Fig~\ref{fig:gadgets}(a)).
+Each edge of this clause gadget will be connected to exactly one variable gadget since every clause in $\phi$ consists of exactly three literals.
+Observe that only one edge at a time can be drawn completely since it intersects both other edges, that the optimal solution in terms of maximized ink is drawing one edge completely (4 units) and the stubs of the other two edges (2 units respectively), so 8 units in total and that all three solutions are equivalent.
+
+\begin{wrapfigure}[11]{r}{0.45\textwidth}
+	\centering
+	\includegraphics[width=.45\linewidth]{circle_true}
+	\hfill
+	\includegraphics[width=.45\linewidth]{circle_false}
+	\caption{Variable gadget representing a true variable (a) and a false variable (b)}
+	\label{fig:optimal_circles}
+\end{wrapfigure}
+
+Circles of segments (variable gadgets) are constructed in a very similar fashion.
+Let $m$ be the maximum number of clauses in which a single variable appears.
+Every variable $x_n$ will be represented by $m$ pairs of segments one red and one green which are connected as seen in Fig.~\ref{fig:gadgets} (b) and form a circle.
+It can be observed that every segment $s_n$ intersects exactly 2 other segments $s_{n-1}$ and $s_{n+1}$ ($s_{2m}$ intersects with $s_1$ to create the circle).
+These intersections split every segmet in three parts, which again have length 1 for the stubs and 2 for the middle part.
+If $x$ is true, the green edges will be drawn completely, if it is false the red edges will be drawn completely.
+Observe that fixing a green edge to be drawn completely fixes all green edges to be drawn completely, like in Fig.~\ref{fig:optimal_circles} (a) (in order to have an optimal solution for this circle) and conversely for the red edges, like in Fig.~\ref{fig:optimal_circles} (b).
+We will associate the first version, with subdivided green edges, with the variable being true and the second one with the variable being false.
+Since all edges are cut in the same fashion, the two solutions are equivalent in terms of maximized ink and amount to $4m + 2m = 6m$.
+
+The variable clauses are arranged on a spine and the clause gadgets are placed above or below the spine, according to the planar embedding of the \ppsat.
+Then the clause gadgets are connected to the variable gadgets corresponding to the occuring variables via segment chains (see Fig.\ref{fig:gadget_connection}, segment chains are for the sake of visualization not all equally long).
+These chains are composed of $t$ (where $t$ is an sufficiently large even integer) equal length segments $s_1$, \dots, $s_t$, which are again coloured alternatingly red and green.
+Every segment $s_i$ crosses with $s_{i-1}$ (except for $s_1$) and $s_{i+1}$ (except for $s_t$).
+All segments are divided again into 3 parts with the stubs having length 1, and the middle part having length 2.
+$s_1$ is also split in this ratio by one of the green stubs of a variable gadget, however the stub is intersected such that the remaining stub has length $1-\varepsilon$.
+The same holds for $s_t$ and a stub of an edge in the clause gadget.
+Here again the optimal solution in terms of maximizing ink is by drawing either all green edges completely and all red ones as stubs, or vice versa.
+Both solutions have a value of $4\frac{t}{2} + 2\frac{t}{2} = 3t$.
+It should be noted that completely drawing the red edges cuts of a small piece (of length $\varepsilon$) of the stubs of both the green edge of the variable gadget and the edge of the clause gadget, to which the chain is connected.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=.8\linewidth]{variable_clause_connection}
+	\caption{Connection between the clause and variable gadgets, created by chains of an even number of segments.}
+	\label{fig:gadget_connection}
+\end{figure}
+
+% Let $q$ be the number of clauses and $p$ be the number of distinct variables in $\phi$.
+%
+% \begin{lemma}
+% 	\label{lem:optimal value}
+% 	The construcetd Graph G has an optimal  value in terms of maximizing ink of $8q + 6mp + 9qt - (4q + 2p)\varepsilon$ if $\varepsilon$ is chosen sufficiently small.
+% \end{lemma}
+%
+% \begin{proof}
+% 	\soeren{i have an idea for this but i don't know if it would be out of scope, or if i might not even need this lemma.
+Also is our \maxsped defined as "Does this graph have a drawing with at least x amount of ink?" since thats what this bound is aiming at?"}
+% \end{proof}
+\soeren{it also still needs to be shown that this bound ensures that the two criterias from~\cite{b-sgh-15}, namely that at least one edge in the triangle is drawn and that the segment chains are alternatingly drawn and not drawn.}
+
+	Let $\phi$ be a boolean formula and $\Pi$ be a planar embedding corresponding to that formula sucht that ($\phi$, $\Pi$) is an instance of \ppsat.
+Let $\Gamma$ be a drawing constructed from that instance in the previously described way.
+We will show that there is a valid truth assignement for $\phi$ if and only if $\Gamma$ admits a \maxsped with a value of at least $8q + 6mp + 3q (3t - 4\varepsilon)$ in terms of maximizing ink.
+	~
+	\begin{enumerate}
+		\item[$\Rightarrow$:] Let $x$ be a variable and $c$ be a clause containing $x$.
+If $x$ is true draw all red edges in the associated variable gadget completely, which in turn means all green edges are drawn partially and vice versa for when $x$ is false.
+By drawing the green edges partially we enable the red edges in the segment chains to be drawn completely (if we shorten the stubs for one green edge in the variable gadget by $\varepsilon$).
+this also means we can draw the edges of the clause gadgets to which this segment chain is connected completely.
+Since $\phi$ has a satisfying variable assignement where exactly one variable per clause is true we have exactly one of these segment chains which allows exactly one edge per clause graph to be drawn completely.
+This results in every clause graph having the value $8 - 4\varepsilon$, every variable gadget having the value of $6m - 2\varepsilon$ and every segment chain having the value of $3t$.
+Let $q$ be the number of clauses and $p$ be the number of distinct variables in $\phi$.
+This results in
+
+		\soeren{is there a nice way to format this?}
+		$(8 - 4\varepsilon)q + (6m - 2\varepsilon)p + 3t*3q = $
+		$8q + 6mp + 9qt - (4q + 2p)\varepsilon$
+
+		\item[$\Leftarrow$:]  Assume $\phi$ does not have a stisfying variable assignement.
+Then there must always exist a clause in which all three contained variables $x_1$, $x_2$ and $x_3$ are false.
+That however means that, by construction, there is a clause gadget where all the corresponding variable gadgets have all their green edges completely drawn.
+This adds at most the value of $2\varepsilon$.
+In order to get the maximal drawing, this propagates through the segment chains up to the clause gadget where it causes all three edges of that gadget to be not drawn.
+This in turn means that this specific gadget has only the value $6 - 6\varepsilon$.
+The total value of this drawing in terms of maximizing ink is therefore at least $2$ smaller than normal.
+If we would try to supplement the missing drawn edge of the clause gadget, we would need to switch up the alternation of drawing segments in the chains, which causes 2 consecutive segments to not be drawn.
+Since the last edge in the chain was dran before, it now can't be drawn and it therefore again results in a difference of at least $2$ in the complete drawing.
+	\end{enumerate}
+
+\soeren{TODO: argue equivalence (both direction)}
+
+\end{proof}

### 2018-01-08 eurocg/paper.tex
sec 4.2 alpha

--- 2018-01-08 eurocg/paper.tex
@@
 
-In this section we close the gap between the known hardness of \maxsped~\cite{b-sgh-15} and the polynomial-time algorithm for 2-plane drawings~\cite{bcgkmn-pped-17} as summarized in the following theorem.
+In this section we close the gap between the known hardness of \maxsped~\cite{b-sgh-15} and the polynomial-time algorithm for 2-plane drawings~\cite{bcgkmn-pped-17}.
 
@@
 \begin{proof}
-	We reduce from the \NP-hard problem \ppsat~\cite{l-pftu-82} using similar ideas as Bruckdorfer's~\cite{b-sgh-15} general proof of the hardness of \maxsped.
-%, which reduces from the more restricted problem \textsc{Planar Positive 1-in-3 Sat}.
+	We reduce from the \NP-hard problem \ppsat~\cite{l-pftu-82} using similar ideas as Bruckdorfer's general proof of the hardness of \maxsped~\cite{b-sgh-15}. 
 	Let $\phi$ be a planar 3-Sat formula with $n$ variables $\{x_1, \dots, x_n\}$ and $m$ clauses $\{c_1, \dots, c_m\}$, each consisting of three literals.
-	We can assume that $\phi$ comes with a planar drawing of its variable-clause graph $H_\phi$ that has a vertex for each variable $x_i$ placed on a horizontal line and a vertex for each clause $c_j$ connected on one side of the horizontal line to the three variables appearing in $c_j$.
+	We can assume that $\phi$ comes with a planar drawing of its variable-clause graph $H_\phi$ that has a vertex for each variable $x_i$ placed on a horizontal line and a vertex for each clause $c_j$ connected to the three variables appearing in $c_j$.
 	In our gadget-based reduction we create a 3-plane drawing $\Gamma_\phi$ as a set of line segments of uniform length and a value $L$ such that $\Gamma$ has a SPED of total ink at least $L$ if and only if $\phi$ is satisfiable.
 
-	All segments are of length 5.
-	We construct our gadgets using pairs of intersecting segments, one colored green, the other one red.
-	Their mutual intersection point is at distance 1 from an endpoint, which has the effect that the maximum amount of ink contributed by such a pair is 7 (one full segment and the other one with two stubs of length 1 each).
+	All segments
 
@@
 \end{proof}
-
+\newpage %delete later just for editing purposes
 \section{Two polynomial special cases}
@@
 In this section we consider the special case that the intersection graph of the $k$-plane input drawing is a tree or a cactus.
-Since every planar graph is representable as intersection graph of line segments~\cite{chalopin2009every}.
-we can also encounter every tree or cactus.
 In both cases we present polynomial-time dynamic programming algorithms.
-
-Let $ C = (V,E) $ be the intersection graph of the edge set $S$ in a given drawing $\Gamma$ of graph $ G $ as defined in Section~\ref{sec:preliminaries}.
-Let $ u \in V $ and $ \deg(u) = k $, then for the corresponding edge $ s(u) \in S $ there are $ k + 1 $ possible stub pairs including drawing the whole edge.
-We store the lengths of the possible stubs as node weights $ \ell_1(u), \dots, \ell_k(u) \in \mathbb{R}_+$ and assume that $ \ell_1(u),\dots,\ell_k(u) $ are sorted from shorter to longer stubs.
+Let $ C = (V,E) $ be the intersection graph of a given drawing $\Gamma$ of graph $ G $ as defined in Section~\ref{sec:preliminaries}.
+Let $ u \in V $ and $ \deg(u) = \delta $, then for the corresponding edge $ s(u) \in S $ there are $ k + 1 $ possible stub pairs including drawing the whole edge.
+We store the lengths of the possible stubs as node weights $ \ell_1(u), \dots, \ell_\delta(u) \in \mathbb{R}_+$ and assume that $ \ell_1(u),\dots,\ell_\delta(u) $ are sorted from shorter to longer stubs.
 We define  $ \ell_0(u) $ as the length of the whole edge $ s(u) $.
@@
 \label{sec:tree}
+\begin{figure}[tbp]
+	\centering
+	\includegraphics{tree_illustration}
+	\caption{One edge $ u $ with intersecting segments. The shaded areas are the stubs, $ p(u) $ is the parent of $ u $. The stubs in the boxed area contribute to $ \solshort(u) $.}
+	\label{fig:tree}
+\end{figure}
 Here we assume that the intersection graph $ C = (V,E) $ is a rooted tree of maximum degree $ k $.
@@
 For a vertex $ u \in V $ let $ p(u) $ denote the parent of $ u $ in $ C $ and let $ c(u) $ denote the set of its children.
-The ink values of the partial \maxsped solutions for the subtrees $T_u$ rooted at each vertex $ u \in V $ such that stub length $ \ell_i(u) $ is used for $s(u)$ are computed and stored as $ T_i(u) $.
-
-\begin{figure}[tbp]
-	\centering
-	\includegraphics{two_solutions_needed}
-	\caption{Consider $ u $, then only the blue and the green solution are relevant.}
-	\label{fig:two_solutions}
-\end{figure}
-
-Since we know $ C $ is a tree every vertex $ u \in V $ with degree $ \delta $ has a parent $ p(u) \in V $.
+For a vertex $ u \in V $ with parent $ p(u) \in V $ let $ a \in \mathbb{R}_+ $ be the shorter length between one of $ s(u) $'s endpoint and the intersection point between $ s(u) $ and $ s(p(u)) $.
+Now we can define the following two values, allowing us to categorize the stub-lengths into those not affecting the parents choice of stub and those which don't:
+%Since we know $ C $ is a tree every vertex $ u \in V $ with degree $ \delta $ has a parent $ p(u) \in V $.
 For this parent there exists one stub length $ \ell_j(p(u)) $ in the sequence of stub lengths $ \ell_0(p(u)),\dots,\ell_{\delta}(p(u)) $ such that drawing $ p(u) $ with these stubs does not intersect a completely drawn $ u $, while drawing $ p(u) $ with $ \ell_{j + 1}(p(u)) $, i.e.
@@
 We call the solution which would intersect the parent $ \sollong(u) $ and the one not intersecting the parent $ \solshort(u) $.
-See Figure~\ref{fig:two_solutions} for an illustration.
+See Figure~\ref{fig:tree} for an illustration.
 More formally let $ j $ be the index such that stubs of length $ \ell_j(p(u)) $ do not intersect $ s(u) $, but $ \ell_{j +1}(p(u)) $ would, we set:
 \begin{align*}
-	\solshort(u) = \max\{T_1(u),\dots,T_j(u)\} \\
+	\solshort(u) = \max\{T_1(u),\dots,T_a(u)\} &&
 	\sollong(u) = \max\{T_0(u),\dots,T_{\delta}(u)\}
@@
 Setting then $ \sollong(u) $ and $ \solshort(u) $ as above gives the two necessary values for $ p(u) $.
-
 %	and $ t = \sollong(r) $ is the maximum value for some $ T_i(r) $'s and $ r $ the root.
@@
 
-\begin{figure}
-	\centering
-	\includegraphics{speedup}
-	\caption{The situation in Lemma~\ref{lem:compute_vertex}.}
-	\label{fig:speedup}
-\end{figure}
 Solving Recurrence~\ref{rec:tree} straight forward would yield an algorithm running in $ O(mk^2) $.
@@
 Assume we computed $ T_j(u) $ for all stub lengths $ \ell_i(u) $ with $ i \leq j $.
-Now we want to compute $ T_{j+1}(u) $ (see Figure~\ref{fig:speedup}).
+Now we want to compute $ T_{j+1}(u) $ (see Figure~\ref{fig:tree}).
 By drawing $ s(u) $ with stub length $ \ell_{j +1} $ we make it a bit longer since we consider the $ \ell_i(u) $ in ascending order.
@@
 \end{proof}
-
-\martin{the figures can still be improved in style and size; maybe merge into one, use nicer colors, draw just one segment and not multiple shifted ones etc}
 
@@
 The values $ \solshort(v) $, $ \solmid(v) $ and $ \sollong(v) $ are computed for all $ v \in c(u) $.
-Again first let $ u $ be an articulation vertex, then the maximum for stub $ \ell_i(A(u)) $ is said value plus the sum of the compatible children, which is just Recurrence~\ref{rec:cactus_block}.
+Again first let $ u $ be an articulation vertex, then the maximum for stub $ \ell_i(A(u)) $ is said value plus the sum of the compatible children, which is just Recurrence~\ref{rec:cactus_articulation}.
 Now let $ u $ be a block vertex.
@@
 \begin{proof}
-	Let $ G $, $ \Gamma $,  $ C = (V,E) $ and $ \mathcal T = (\mathcal V,\mathcal E) $ as above and $ u \in \mathcal V $ a block vertex with degree $ \delta $.
-Computing the value for the cycle itself $ \textit{cycle}(u) $ takes $ O(n) $ time by Bruckdorfer et al.
-Now at every vertex $ v \in \textit{cycle}(u) $ there might be a set of up to $ \delta $ many children, which are all articulation points.
-Their values are already known and maximization over them can hence be done in $ O(\delta) $ time by the same argument as in Lemma~\ref{lem:compute_vertex}.
-Following we need $ O(n\delta) $ to compute the three values $ \solshort(u) $, $ \solmid(u) $ and $ \sollong(u) $.
+	Let $ G $, $ \Gamma $,  $ C = (V,E) $ with maximum degree $ k $ and $ \mathcal T = (\mathcal V,\mathcal E) $ as above and $ u \in \mathcal V $ a block vertex.
+Computing the value for the cycle itself $ \textit{cycle}(u) $ takes $ O(n_c) $ time by Bruckdorfer et al.
+with $ n_c $ being the number of vertices in the cycle.
+Now at every vertex $ v \in \textit{cycle}(u) $ there might be a set of up to $ k $ many children, which are all articulation points.
+Their values are already known and maximization over them can hence be done in $ O(k) $ time by the same argument as in Lemma~\ref{lem:compute_vertex}.
+Following we need $ O(n_ck) $ to compute the three values $ \solshort(u) $, $ \solmid(u) $ and $ \sollong(u) $.
 \end{proof}
@@
 	Let $ G $ be a simple graph with $m$ edges and $ \Gamma $ a straight-line drawing of $ G $.
-If the intersection graph $ C = (V,E) $ of $ G $ is a cactus with maximum degree $ k \in \mathbb{N} $ the \maxsped problem can be solved in $ O(mk^2 + m \log m)) $ time and space.
+If the intersection graph $ C = (V,E) $ of $ G $ is a cactus with maximum degree $ k \in \mathbb{N} $ the \maxsped problem can be solved in $ O(mk + m \log m)) $ time and space.
 \end{theorem}

### 2018-01-08 eurocg/paper.tex

--- 2018-01-08 eurocg/paper.tex
@@
 
+\newpage %delete later
 \section{Two polynomial special cases}

### 2018-01-08 eurocg/paper.tex
some polishing

@@
 	The last edge of the path intersects the corresponding clause edge with the same length ratios.
-	The path itself consists of a chain of red-green segment pairs, where, as before, each pair can contribute an ink value of at most~7.
+	The path itself consists of a chain of red-green segment pairs, so each pair contributes an ink value of at most~7.
 
@@
 	For the two non-selected literals, we can draw the last segments of their paths as full segments, as well as every segment at an even position, while the segments at odd positions are drawn as stubs.
-	This is compatible with any of the two variable configurations and proves that we can indeed achieve ink value $L$.
+	This is compatible with any of the two variable configurations and proves that we can indeed achieve ink value~$L$.
 
-	Conversely, assume that we have a SPED with ink $L$. 
-	By construction, every red-green segment pair and all clause gadgets must contribute their respective maximum ink value.
+	Conversely, assume that we have a SPED with ink value $L$. 
+	By construction, every red-green segment pair and every clause gadget must contribute its respective maximum ink value.
 	In particular, each variable gadget is either in state \emph{true} or \emph{false}.
 	We claim that this truth assignment satisfies $\phi$.
-	For each clause, one segment must have full length and as argued before the path connecting that segment to a variable gadget must have all segments at even positions as stubs and all segments at odd positions as full segments.
+	For each clause, one segment must have full length and as argued before the path connecting that segment to a variable gadget must have all segments at even positions as length-1 stubs and all segments at odd positions as full segments.
 	In particular the first segment of the literal path is fully drawn and therefore in a SPED with ink $L$ the intersecting segment of the variable cannot be fully drawn.

### 2018-01-08 eurocg/paper.tex
section 4.1 beta, 4.2 alpha

--- 2018-01-08 eurocg/paper.tex
@@
 	\centering
-	\includegraphics{tree_illustration}
+%	\includegraphics{tree_illustration}
 	\caption{One edge $ u $ with intersecting segments. The shaded areas are the stubs, $ p(u) $ is the parent of $ u $. The stubs in the boxed area contribute to $ \solshort(u) $.}
@@
 \end{align}
-Solving Recurrence~\ref{rec:tree} yields an optimal solution to the \maxsped problem on $ G $ with drawing $ \Gamma $ as the following lemma shows.
-
-\begin{lemma}
-	\label{lem:tree_correctness}	
-	Let $ G $ be a simple graph and $ \Gamma $ a straight-line drawing of $ G $.
+By solving Recurrence~\ref{rec:tree} we find an optimal solution to the \maxsped problem on $ G $ with drawing $ \Gamma $.	Let $ C = (V,E) $ as above and $ u \in V $ a leaf of $ C $.
+Since $ c(u) $ is empty we simply get $ \solshort(u) = T_0(u) = \ell_1(u) $ and $ \sollong(u) = T_1(u) = \max\{\ell_0(u), \ell_1(u)\} $.
+Now let $ u \in V $ with degree $ \delta $ be an inner vertex.
+The values $ \solshort(v) $ and $ \sollong(v) $ are computed for all $ v \in c(u) $.
+$ T_i(u) $ for one $ 0 \leq i \leq \delta $ is then the sum of the maximum stubs we can draw among the children plus the stub length $ \ell_i(u) $ itself, which are just the values we are interested in.
+Setting then $ \sollong(u) $ and $ \solshort(u) $ as above gives the two necessary values for $ p(u) $.
+%\begin{lemma}
+%	\label{lem:tree_correctness}	
+%	Let $ G $ be a simple graph and $ \Gamma $ a straight-line drawing of $ G $.
 If the intersection graph $ C = (V,E) $ of $ G $ is a tree with root $ r \in V $ and maximum degree $ k $, then $ \sollong(r) = \max \{T_i(r) \mid 0 \leq i \leq k \} $ is an optimal solution to the \maxsped problem on $ G $ with drawing $ \Gamma $.
-\end{lemma}
-\begin{proof}
-	Let $ G $, $ \Gamma $ and $ C = (V,E) $ as above and $ u \in V $ a leaf of $ C $.
+%\end{lemma}
+%\begin{proof}
+%	Let $ G $, $ \Gamma $ and $ C = (V,E) $ as above and $ u \in V $ a leaf of $ C $.
 Since $ c(u) $ is empty we simply get $ \solshort(u) = T_0(u) = \ell_1(u) $ and $ \sollong(u) = T_1(u) = \max\{\ell_0(u), \ell_1(u)\} $.
@@
 Hence if $ t $ was not optimal we would need to find at least a third stub length for at least one child $ v \in c(u) $ which was neither equal to $ \sollong(v) $ nor $ \solshort(v) $.
-\end{proof}
+%\end{proof}
 
@@
 	 Let $ u \in V $ be a vertex with degree $ \delta < k $ for which we have computed $ \sollong $ and $ \solshort $ for every of its children. 
-The crucial part is that $ T_i(u) $ can be built iteratively by going through the list of stub lengths for $ u $ in ascending order.
-First we compute $ T_1(u) $, this is the solution for $ u $ in which we are able to draw the children with any of their stub lengths.
-This means we need to build the sum of $ \ell_1(u) $ and the $ \sollong(v) $ for every $ v \in c(u) $ which runs in $ O(\delta) $.
-The same can be done for $ T_0(u) $, but taking all the $ \solshort(v) $ values for the children $ v \in c(u) $.
-Assume we computed $ T_j(u) $ for all stub lengths $ \ell_i(u) $ with $ i \leq j $.
-Now we want to compute $ T_{j+1}(u) $ (see Figure~\ref{fig:tree}).
-By drawing $ s(u) $ with stub length $ \ell_{j +1} $ we make it a bit longer since we consider the $ \ell_i(u) $ in ascending order.
-Hence by doing so we now potentially intersect the corresponding edge of the $ j $-th child $ v_j \in c(u) $.
-Since $ s(v_j) $ is the only segment affected by drawing $ u $ with stub length $ \ell_{j + 1} $ instead of $ \ell_j $ we can compute $ T_{j+1}(u) $ in $O(1)$ time by simply subtracting $ \sollong(v) $ and adding $ \solshort(v) $.
+%	 The crucial part is that $ T_i(u) $ can be built iteratively by going through the list of stub lengths for $ u $ in ascending order. 
+	First we compute $ T_0(u) $ and $ T_1(u) $, this can be done since they represent the longest (whole edge) and shortest stub lengths. 	 
+	 Now assume we computed $ T_j(u) $ for all stub lengths $ \ell_i(u) $ with $ i \leq j $, hence we want to compute $ T_{j+1}(u) $ next.
+Doing so only requires to consider the next stub-length $ \ell_{j + 1}(u) $, since all other affected by drawing $ s(u) $ with $ \ell_{j+1}(u) $ where already considered.
+Hence we can compute $ T_{j+1}(u) $ in $O(1)$ time by simply subtracting $ \sollong(v) $ and adding $ \solshort(v) $.
 
@@
 	Since $ C $ is known to be a tree, and hence $ |E| = O(m) $, we can compute the intersection graph and the stub lengths in $ O(m\log m) $ time and space.
-
-	To solve the \maxsped problem on $ G $ we need to compute $ \max \{T_i(r) \mid 0 \leq i \leq k \} $ for the root $ r \in V $ of $ C $ by Lemma~\ref{lem:tree_correctness}.
+	With Lemma~\ref{lem:compute_vertex} and Recurrence~\ref{rec:tree} we need one pass through the tree to find the roots maximum in $ O(mk) $ time.
+Adding the bounds yields the desired runtime of $ O(mk + m \log m)) $.
+%	To solve the \maxsped problem on $ G $ we need to compute $ \max \{T_i(r) \mid 0 \leq i \leq k \} $ for the root $ r \in V $ of $ C $ as seen above.
 Doing this bottom up requires one pass through $ C $ hence by applying Lemma~\ref{lem:compute_vertex} we need $ O(mk) $ time to compute all $ T_i(u) $ for all $ u \in V $.
@@
 Adding the bounds yields the desired runtime of $ O(mk + m \log m)) $.
-
+%	
 	Retrieving the actual stub length for each edge in $ G $ needed to produce the SPED can be done by backtracking the subsolutions which requires at most $ O(mk) $ time.

### 2018-01-09 eurocg/paper.tex
smaller tree illustration
minor edits in hardness
removed some lemmas

@@
 	\centering
-%	\includegraphics{tree_illustration}
+	\includegraphics{tree_illustration}
 	\caption{One edge $ u $ with intersecting segments. The shaded areas are the stubs, $ p(u) $ is the parent of $ u $. The stubs in the boxed area contribute to $ \solshort(u) $.}
@@
 \end{align}
-
-\begin{lemma}
-	\label{lem:cactus_correctness}	
-	Let $ G $ be a simple graph and $ \Gamma $ a straight-line drawing of $ G $.
-If the intersection graph $ C = (V,E) $ of $ G $ is a cactus with root $ r \in V $ and maximum degree $ k $, then $ \sollong(r) = \max \{T_i(r) \mid 0 \leq i \leq k \} $ is an optimal solution to the \maxsped problem on $ G $ with drawing $ \Gamma $.
-\end{lemma}
-\begin{proof}	
 	Let $ G $, $ \Gamma $,  $ C = (V,E) $ and $ \mathcal T = (\mathcal V,\mathcal E) $ as above and $ u \in \mathcal V $ a leaf of $ \mathcal T$.
@@
 This is especially a tree and we use the aglorithm from Section~\ref{sec:tree}, maximizing over three values of the articulation vertices though,  to compute the $ T_i(u) $.
-\end{proof}
-
-\begin{lemma}
-	\label{lem:compute_articulation}
-	Let $ \mathcal T = (\mathcal V,\mathcal E) $ be a cactus with maximum degree $ k $ and vertex weights $ \ell_0,\dots,\ell_{\delta} $ for every articulation vertex $ u \in \mathcal V $ with degree $ \delta $.
-We can compute $ \solshort(u) $, $ \solmid(u) $ and $ \sollong(u) $ in $ O(\delta) $ time and space, if the values $ \sollong(v) $, $ \solmid(v) $ and $ \solshort(v) $ are computed for all children $ v \in c(u) $.
-\end{lemma}
-\begin{proof}
-	Follows from Lemma~\ref{lem:compute_vertex}, since an articulation vertex has at most $ \delta $ many children in $ \mathcal{T} $, one for each cycle it is part of in $ C $ and one for every other articulation vertex it was already connected to in $ C $.
-\end{proof}
-
+%\begin{lemma}
+%	\label{lem:cactus_correctness}	
+%	Let $ G $ be a simple graph and $ \Gamma $ a straight-line drawing of $ G $.
+If the intersection graph $ C = (V,E) $ of $ G $ is a cactus with root $ r \in V $ and maximum degree $ k $, then $ \sollong(r) = \max \{T_i(r) \mid 0 \leq i \leq k \} $ is an optimal solution to the \maxsped problem on $ G $ with drawing $ \Gamma $.
+%\end{lemma}
+%\begin{proof}	
+%	Let $ G $, $ \Gamma $,  $ C = (V,E) $ and $ \mathcal T = (\mathcal V,\mathcal E) $ as above and $ u \in \mathcal V $ a leaf of $ \mathcal T$.
+Then either $ u $ is an articulation vertex and to be treated as in Lemma~\ref{lem:compute_vertex} or a cycle.
+In the latter case we find the optimal solution to the \maxsped problem on $ C[\textit{cycle}(u)] $ with the algorithm from Bruckdorfer et al.
+\cite{bcgkmn-pped-17}.
+Using the above described method we derive $ \solshort(u) $, $ \solmid(u) $ and $ \sollong(u) $, this are exactly the optimal values depending on how the cycle interacts with the articulation vertex $ p(u) $.
+%
+%	Now let $ u \in V $ with degree $ \delta $ be an inner vertex of $ \mathcal T $.
+The values $ \solshort(v) $, $ \solmid(v) $ and $ \sollong(v) $ are computed for all $ v \in c(u) $.
+Again first let $ u $ be an articulation vertex, then the maximum for stub $ \ell_i(A(u)) $ is said value plus the sum of the compatible children, which is just Recurrence~\ref{rec:cactus_articulation}.
+Now let $ u $ be a block vertex.
+We have to cut open the cycle it represents (compare \cite{bcgkmn-pped-17}).
+We do this at $ p = A(p(u)) $.
+The resulting graph is a path over the vertices in $ \textit{cycle}(u) $ with attached articulation vertices.
+This is especially a tree and we use the aglorithm from Section~\ref{sec:tree}, maximizing over three values of the articulation vertices though,  to compute the $ T_i(u) $.
+%\end{proof}
 \begin{lemma}
 	\label{lem:compute_block}
-	Let $ \mathcal T = (\mathcal V,\mathcal E) $ be a cactus with maximum degree $ k $ and vertex weights $ \ell_0,\dots,\ell_{\delta} $ for every block vertex $ u \in \mathcal V $ with degree $ \delta $.
+	Let $ \mathcal T = (\mathcal V,\mathcal E) $ be a cactus with maximum degree $ k $ and vertex weights $ \ell_0,\dots,\ell_{\delta} $ for every vertex $ u \in \mathcal V $ with degree $ \delta $.
 We can compute $ \solshort(u) $, $ \solmid(u) $ and $ \sollong(u) $ in $ O(n\delta) $ time and space, if the values $ \sollong(v) $, $ \solmid(v) $ and $ \solshort(v) $ are computed for all children $ v \in c(u) $.
@@
 \begin{proof}
-	Let $ G $, $ \Gamma $,  $ C = (V,E) $ with maximum degree $ k $ and $ \mathcal T = (\mathcal V,\mathcal E) $ as above and $ u \in \mathcal V $ a block vertex.
+	Let $ G $, $ \Gamma $,  $ C = (V,E) $ with maximum degree $ k $ and $ \mathcal T = (\mathcal V,\mathcal E) $ as above.
+First let $ u \in \mathcal V $ be an articulation vertex, then the result follows from Lemma~\ref{lem:compute_vertex}, since an articulation vertex has at most $ \delta $ many children in $ \mathcal{T} $, one for each cycle it is part of in $ C $ and one for every other articulation vertex it was already connected to in $ C $.
+
+	Now let $ u \in \mathcal V $ be a block vertex.
 Computing the value for the cycle itself $ \textit{cycle}(u) $ takes $ O(n_c) $ time by Bruckdorfer et al.
@@
 \begin{proof}
-	Follows as in Theorem~\ref{thm:tree} by maximizing over the tree using Lemma~\ref{lem:cactus_correctness} and Recurrence~\ref{rec:cactus_articulation}.
-Then the running time follows from constructing the intersection graph and the block-cut tree as well as Lemma~\ref{lem:compute_articulation} and~\ref{lem:compute_block}.
+	Follows as in Theorem~\ref{thm:tree} by maximizing over the tree using Recurrence~\ref{rec:cactus_articulation}.
+Then the running time follows from constructing the intersection graph and the block-cut tree as well as Lemma~\ref{lem:compute_block}.
 \end{proof}

### 2018-01-09 eurocg/paper.tex
removed some lemmas

@@
 
-\newpage %delete later
+%\newpage %delete later
 \section{Two polynomial special cases}
@@
 If $ u $ is an articulation vertex this are exactly the positions where the cycle represented by $ p(u) $ intersects $ s(A(u)) $ in $ \Gamma $.
-If $ u $ is a block vertex $ a,b $ are the points where the by $ \textit{cycle}(u) $ represented edges intersect $ s(A(p(u))) $.
+If $ u $ is a block vertex $ a,b $ are the points where edges $ s(v), v \in \textit{cycle}(u) $ intersect $ s(A(p(u))) $.
 Similar to Section~\ref{sec:tree} we define three values for every vertex $ u \in \mathcal V $ with degree $ \delta $:
@@
 	\solmid(u) &= \max\{T_1(u),\dots,T_{a}(u)\}\\
-	\sollong(u) &= \max\{T_1(u),\dots,T_{b}(u)\}\\
+	\sollong(u) &= \max\{T_1(u),\dots,T_{b}(u)\}
 \end{align*}
-
 For an articulation vertex $ u \in \mathcal V $:
@@
 	\label{lem:compute_block}
-	Let $ \mathcal T = (\mathcal V,\mathcal E) $ be a cactus with maximum degree $ k $ and vertex weights $ \ell_0,\dots,\ell_{\delta} $ for every vertex $ u \in \mathcal V $ with degree $ \delta $.
-We can compute $ \solshort(u) $, $ \solmid(u) $ and $ \sollong(u) $ in $ O(n\delta) $ time and space, if the values $ \sollong(v) $, $ \solmid(v) $ and $ \solshort(v) $ are computed for all children $ v \in c(u) $.
+	Let $ \mathcal T = (\mathcal V,\mathcal E) $ be a cactus with maximum degree $ k $ and vertex weights $ \ell_0,\dots,\ell_{\delta} $ for every block vertex $ u \in \mathcal V $ with degree $ \delta $.
+We can compute $ \solshort(u) $, $ \solmid(u) $ and $ \sollong(u) $ in $ O(n_c\delta) $ time and space, if the values $ \sollong(v) $, $ \solmid(v) $ and $ \solshort(v) $ are computed for all children $ v \in c(u) $.
 \end{lemma}
@@
 	Let $ G $, $ \Gamma $,  $ C = (V,E) $ with maximum degree $ k $ and $ \mathcal T = (\mathcal V,\mathcal E) $ as above.
-First let $ u \in \mathcal V $ be an articulation vertex, then the result follows from Lemma~\ref{lem:compute_vertex}, since an articulation vertex has at most $ \delta $ many children in $ \mathcal{T} $, one for each cycle it is part of in $ C $ and one for every other articulation vertex it was already connected to in $ C $.
-
-	Now let $ u \in \mathcal V $ be a block vertex.
+Let $ u \in \mathcal V $ be a block vertex.
 Computing the value for the cycle itself $ \textit{cycle}(u) $ takes $ O(n_c) $ time by Bruckdorfer et al.
@@
 	Follows as in Theorem~\ref{thm:tree} by maximizing over the tree using Recurrence~\ref{rec:cactus_articulation}.
-Then the running time follows from constructing the intersection graph and the block-cut tree as well as Lemma~\ref{lem:compute_block}.
+Then the running time follows from constructing the intersection graph and the block-cut tree as well as Lemma~\ref{lem:compute_block} for the block vertices and applying Lemma~\ref{lem:compute_vertex} for the articulation vertices.
 \end{proof}

### 2018-01-09 eurocg/paper.tex
minor
preliminaries polishing

@@
 We call $ \Gamma $ \emph{$ k $-plane} if every edge $ s_i \in S $ is crossed by at most $ k $ other edges from $ S $ in $ \Gamma $.
-We use the terms edge in $ S $ and segment in $ \Gamma $ interchangeably, since our input is always a graph $G$ with a given drawing $ \Gamma $.
+We use the terms edge in $ S $ and segment in $ \Gamma $ interchangeably.
+%, since our input is always a graph $G$ with a given drawing $ \Gamma $.
 Hence we can also interpret $S$ as a set of line segments.
@@
 The \emph{intersection graph} $ C = (V,E) $ of $ \Gamma $ is the graph containing a vertex $v_i$ in $ V $ for every $ s_i \in S $ and an edge $ v_i v_j \in E $ between vertices $ v_i, v_j \in V $ if the corresponding edges $ s_i, s_j \in S $ intersect in $ \Gamma $.
+We also denote the segment in $S$ corresponding to a vertex $v \in V$ by $s(v)$.
 Observe that the intersection graph $ C $ of a $ k $-plane drawing $ \Gamma $ has maximum degree $ k $.  
@@
 The \emph{ink} or \emph{ink value}  $I(D)$ of a SPED $D$ is the total stub length $I(D) = \sum_{s \in S} f_s |s|$.
-In the problem \maxsped the task is to find for a given drawing $\Gamma$ a SPED $D^*$ such that its ink $I(D^*)$ is maximum over all SPEDs.
+In the problem \maxsped the task is to find for a given drawing~$\Gamma$ a SPED $D^*$ such that its ink $I(D^*)$ is maximum over all SPEDs.
 
@@
 \begin{theorem}\label{thm:hard}
-	The problem \maxsped is \NP-hard even for 3-plane graph drawings.
+	\maxsped is \NP-hard even for 3-plane graph drawings.
 \end{theorem}

### 2018-01-09 eurocg/paper.tex
minor edit in hardness

@@
 	We can assume that $\phi$ comes with a planar drawing of its variable-clause graph $H_\phi$ that has a vertex for each variable $x_i$ placed on a horizontal line and a vertex for each clause $c_j$ either above or below the horizontal line and connected on that side of the line to the three variables appearing in $c_j$.
-	In our gadget-based reduction we create a 3-plane drawing $\Gamma_\phi$ as a set of line segments of uniform length and a value $L$ such that $\Gamma$ has a SPED with ink at least $L$ if and only if $\phi$ is satisfiable.
+	In our gadget-based reduction we mimic the drawing of $H_\phi$ by creating a 3-plane drawing $\Gamma_\phi$ as a set of line segments of uniform length and a value $L$ such that $\Gamma$ has a SPED with ink at least $L$ if and only if $\phi$ is satisfiable.
 	Figure~\ref{fig:gadget_connection} shows a sketch of the gadgets.

### 2018-01-09 eurocg/paper.tex
zwischenstand

@@
 Let $ C = (V,E) $ be the intersection graph of a given drawing $\Gamma$ of graph $ G $ as defined in Section~\ref{sec:preliminaries}.
-Let $ u \in V $ and $ \deg(u) = \delta $, then for the corresponding edge $ s(u) \in S $ there are $ k + 1 $ possible stub pairs including drawing the whole edge.
+Let $ u \in V $ and $ \deg(u) = \delta $.
+Then for the corresponding segment $ s(u) \in S $ there are $ \delta + 1 $ possible stub pairs including drawing the whole segment.
 We store the lengths of the possible stubs as node weights $ \ell_1(u), \dots, \ell_\delta(u) \in \mathbb{R}_+$ and assume that $ \ell_1(u),\dots,\ell_\delta(u) $ are sorted from shorter to longer stubs.
-We define  $ \ell_0(u) $ as the length of the whole edge $ s(u) $.
+We define  $ \ell_0(u) $ as the length of the whole segment $ s(u) $.
 

### 2018-01-09 eurocg/paper.tex
zwischenstand

@@
 Then for the corresponding segment $ s(u) \in S $ there are $ \delta + 1 $ possible stub pairs including drawing the whole segment.
-We store the lengths of the possible stubs as node weights $ \ell_1(u), \dots, \ell_\delta(u) \in \mathbb{R}_+$ and assume that $ \ell_1(u),\dots,\ell_\delta(u) $ are sorted from shorter to longer stubs.
+We store the lengths of the possible stub pairs as node weights $ \ell_1(u), \dots, \ell_\delta(u) \in \mathbb{R}_+$ and assume that $ \ell_1(u),\dots,\ell_\delta(u) $ are sorted from shorter to longer stubs.
 We define  $ \ell_0(u) $ as the length of the whole segment $ s(u) $.
@@
 \end{figure}
-Here we assume that the intersection graph $ C = (V,E) $ is a rooted tree of maximum degree $ k $.
+Here we assume that $ C =(V,E) $ is a rooted tree of maximum degree $ k $. 
 We give a bottom up dynamic programming algorithm for solving \maxsped on $ C $. 
-For a vertex $ u \in V $ let $ p(u) $ denote the parent of $ u $ in $ C $ and let $ c(u) $ denote the set of its children.
-For a vertex $ u \in V $ with parent $ p(u) \in V $ let $ a \in \mathbb{R}_+ $ be the shorter length between one of $ s(u) $'s endpoint and the intersection point between $ s(u) $ and $ s(p(u)) $.
-Now we can define the following two values, allowing us to categorize the stub-lengths into those not affecting the parents choice of stub and those which don't:
+For each vertex $u \in V$ we compute and store the maximum ink values $T_i(u)$ for $i = 0, \dots, \delta$ with $\delta = \deg(u)$ for the subtree rooted at $u$ such that $s(u)$ is drawn as a pair of stubs of length $\ell_i(u)$.
+For $ u \in V $ let $ p(u) $ denote the parent of $ u $ in $ C $ and let $ c(u) $ denote the set of its children. 
+For $ u \in V $  %with parent $ p(u) \in V $ 
+let $ i_p $ be the index of the stub length $\ell_{i_p}(u)$ induced by the intersection point of $s(u)$ and $s(p(u))$.
+%the shorter length between one of $ s(u) $'s endpoints and the intersection point between $ s(u) $ and $ s(p(u)) $. 
+We define the following two values, which allow us to categorize the stub lengths into those not affecting the stubs of the parent and those that do affect the parent:
 %Since we know $ C $ is a tree every vertex $ u \in V $ with degree $ \delta $ has a parent $ p(u) \in V $.
@@
 More formally let $ j $ be the index such that stubs of length $ \ell_j(p(u)) $ do not intersect $ s(u) $, but $ \ell_{j +1}(p(u)) $ would, we set:
-\begin{align*}
-	\solshort(u) = \max\{T_1(u),\dots,T_a(u)\} &&
+%\begin{align*}
+\[	
+	\solshort(u) = \max\{T_1(u),\dots,T_{i_p}(u)\} \qquad
 	\sollong(u) = \max\{T_0(u),\dots,T_{\delta}(u)\}
-\end{align*}
-It remains to define $ T_i(u) $ for a vertex $ u $ and stub length $ \ell_i(u) $:
+\]
+%\end{align*}
+Based on these two values we can recursively define $ T_i(u) $: % for a vertex $ u $ and stub length $ \ell_i(u) $:
 \begin{align}
@@
 	\begin{cases}
-		\solshort(v) & \text{if } s(u) \text{ with length } \ell_i \text{ intersects } s(v) \\
+		\solshort(v) & \text{if } s(u) \text{ with length } \ell_i(u) \text{ intersects } s(v) \\
 		\sollong(v) & \text{else}
@@
 \end{align}
-By solving Recurrence~\ref{rec:tree} we find an optimal solution to the \maxsped problem on $ G $ with drawing $ \Gamma $.	Let $ C = (V,E) $ as above and $ u \in V $ a leaf of $ C $.
-Since $ c(u) $ is empty we simply get $ \solshort(u) = T_0(u) = \ell_1(u) $ and $ \sollong(u) = T_1(u) = \max\{\ell_0(u), \ell_1(u)\} $.
-Now let $ u \in V $ with degree $ \delta $ be an inner vertex.
-The values $ \solshort(v) $ and $ \sollong(v) $ are computed for all $ v \in c(u) $.
-$ T_i(u) $ for one $ 0 \leq i \leq \delta $ is then the sum of the maximum stubs we can draw among the children plus the stub length $ \ell_i(u) $ itself, which are just the values we are interested in.
-Setting then $ \sollong(u) $ and $ \solshort(u) $ as above gives the two necessary values for $ p(u) $.
+The correctness of Recurrence~(\ref{rec:tree}) follows by induction. 
+For a leaf $u$ in $C$ the set $c(u)$ is empty and the correctness of $T_i(u)$ is immediate.
+Further, $\solshort(u) = T_1(u)$ and $\sollong(u) = T_0(u)$ are set correctly for the parent $p(u)$.
+%Let $ C = (V,E) $ as above and $ u \in V $ a leaf of $ C $. Since $ c(u) $ is empty we simply get $ \solshort(u) = T_0(u) = \ell_1(u) $ and $ \sollong(u) = T_1(u) = \max\{\ell_0(u), \ell_1(u)\} $. 
+For an inner vertex $ u $ with degree $\delta$ we can assume by the induction hypothesis that the values $ \solshort(v) $ and $ \sollong(v) $ are computed correctly for all children $ v \in c(u) $. 
+Each value $ T_i(u) $ for $ 0 \leq i \leq \delta $ is then the stub length $ \ell_i(u) $ plus the sum of the maximum ink we can achieve among the children subject to the stubs of $u$ being drawn with length $\ell_i(u)$.
+Setting $ \sollong(u) $ and $ \solshort(u) $ as above yields the two maximum ink values that are relevant for $ p(u) $.
 %\begin{lemma}
@@
 
-Solving Recurrence~\ref{rec:tree} straight forward would yield an algorithm running in $ O(mk^2) $.
-Using the order on the stub lengths though we can improve this to $ O(mk) $ by solving $ T_i(u) $ in linear time for one $ u \in V $.
+Recurrence~(\ref{rec:tree}) can be solved naively in $ O(mk^2) $ time. 
+Using the order on the stub lengths  we can improve this to $ O(mk) $ time by computing all $ T_i(u) $ for one $ u \in V $ in $O(k)$ time.
+Let $u \in V$ be a vertex with degree $\deg(u) = \delta$. 
+The values $T_0(u) = \ell_0(u) + \sum_{v\in c(u)} \solshort(v)$ and $T_1(u) = \ell_1(u) + \sum_{v\in c(u)} \sollong(v)$ for the whole segment $s(u)$ and the shortest stubs can be computed in $O(k)$ time each.
 
@@
 %
+
+By solving Recurrence~(\ref{rec:tree}) we find an optimal solution to the \maxsped problem on $ G $ with drawing $ \Gamma $ as $\max\{T_0(r), \dots, T_{\deg(r)}(r)\}$ for the root $r$ of $C$.
+
 	Retrieving the actual stub length for each edge in $ G $ needed to produce the SPED can be done by backtracking the subsolutions which requires at most $ O(mk) $ time.

### 2018-01-09 eurocg/paper.tex
zwischenstand

@@
 The values $T_0(u) = \ell_0(u) + \sum_{v\in c(u)} \solshort(v)$ and $T_1(u) = \ell_1(u) + \sum_{v\in c(u)} \sollong(v)$ for the whole segment $s(u)$ and the shortest stubs can be computed in $O(k)$ time each.
+Now $T_{j+1}(u)$ can be computed from $T_j(u)$ in $O(1)$ time as follows.
+Let $v_j $ be the neighbor of $u$ that induces stub length $\ell_j(u)$ and assume $v_j \ne p(u)$.
+In $T_j(u)$ we could still count the value $\sollong(v_j)$, but in $T_{j+1}(u)$ the stub length of $u$ implies that $v_j$ can contribute only $\solshort(v_j)$. 
+Then $T_{j+1}(u) = T_j(u) - \sollong(v_j) + \solshort(v_j)$.
+If $v_j=p(u)$ the two values $T_j(u)$ and $T_{j+1}(u)$ are equal as the corresponding change in stub length has no effect on the children of $u$.
+Computing $\solshort(u)$ and $\sollong(u)$ clearly takes $O(k)$ time.
 
-\begin{lemma}
-	\label{lem:compute_vertex}
-	Let $ C = (V,E) $ be a tree with maximum degree $ k $ and vertex weights $ \ell_0,\dots,\ell_{\delta} $ for every $ u \in V $ with degree $ \delta $.
+% \begin{lemma}
+% 	\label{lem:compute_vertex}
+% 	Let $ C = (V,E) $ be a tree with maximum degree $ k $ and vertex weights $ \ell_0,\dots,\ell_{\delta} $ for every $ u \in V $ with degree $ \delta $.
 We can compute $ \solshort(u) $ and $ \sollong(u) $ in $ O(\delta) $ time and space, if the values $ \sollong(v) $ and $ \solshort(v) $ are computed for all children $ v \in c(u) $.
-\end{lemma}
-\begin{proof}
-	 Let $ u \in V $ be a vertex with degree $ \delta < k $ for which we have computed $ \sollong $ and $ \solshort $ for every of its children. 
-%	 The crucial part is that $ T_i(u) $ can be built iteratively by going through the list of stub lengths for $ u $ in ascending order. 
-	First we compute $ T_0(u) $ and $ T_1(u) $, this can be done since they represent the longest (whole edge) and shortest stub lengths. 	 
-	 Now assume we computed $ T_j(u) $ for all stub lengths $ \ell_i(u) $ with $ i \leq j $, hence we want to compute $ T_{j+1}(u) $ next.
+% \end{lemma}
+% \begin{proof}
+% 	 Let $ u \in V $ be a vertex with degree $ \delta < k $ for which we have computed $ \sollong $ and $ \solshort $ for every of its children.
+% %	 The crucial part is that $ T_i(u) $ can be built iteratively by going through the list of stub lengths for $ u $ in ascending order.
+% 	First we compute $ T_0(u) $ and $ T_1(u) $, this can be done since they represent the longest (whole edge) and shortest stub lengths.
+% 	 Now assume we computed $ T_j(u) $ for all stub lengths $ \ell_i(u) $ with $ i \leq j $, hence we want to compute $ T_{j+1}(u) $ next.
 Doing so only requires to consider the next stub-length $ \ell_{j + 1}(u) $, since all other affected by drawing $ s(u) $ with $ \ell_{j+1}(u) $ where already considered.
 Hence we can compute $ T_{j+1}(u) $ in $O(1)$ time by simply subtracting $ \sollong(v) $ and adding $ \solshort(v) $.
-
-	Finally we need to compute $ \solshort(u) $ and $ \sollong(u) $, but this are just two maximizations over at most $ \delta + 1 $ values at most.
+%
+% 	Finally we need to compute $ \solshort(u) $ and $ \sollong(u) $, but this are just two maximizations over at most $ \delta + 1 $ values at most.
 This can be done in $ O(\delta) $ by simply sweeping over the values.
-\end{proof}
+% \end{proof}
 

### 2018-01-09 eurocg/paper.tex
zwischenstand

@@
 
+\martin{$f_s$ mit $\ell_i$ linken}
+
 %\fabian{This part has to be adapted once the intro is fixed}
@@
 	We can assume that $\phi$ comes with a planar drawing of its variable-clause graph $H_\phi$ that has a vertex for each variable $x_i$ placed on a horizontal line and a vertex for each clause $c_j$ either above or below the horizontal line and connected on that side of the line to the three variables appearing in $c_j$.
-	In our gadget-based reduction we mimic the drawing of $H_\phi$ by creating a 3-plane drawing $\Gamma_\phi$ as a set of line segments of uniform length and a value $L$ such that $\Gamma$ has a SPED with ink at least $L$ if and only if $\phi$ is satisfiable.
+	In our gadget-based reduction we mimic the drawing of $H_\phi$ by creating a 3-plane drawing $\Gamma_\phi$ as a set of line segments of uniform length and a value $L$ such that $\Gamma_\phi$ has a SPED with ink at least $L$ if and only if $\phi$ is satisfiable.
 	Figure~\ref{fig:gadget_connection} shows a sketch of the gadgets.
@@
 		\centering
-		\includegraphics[width=.5\linewidth]{variable_clause_connection}
-		\caption{Connection between the clause and variable gadgets, created by chains of an even number of segments.}
+		\includegraphics{Reduction_visual}
+		\caption{Example of three variable gadgets and one clause gadget.}
 		\label{fig:gadget_connection}

### 2018-01-09 eurocg/paper.tex
update gadget figure

@@
 		\includegraphics{Reduction_visual}
-		\caption{Example of three variable gadgets and one clause gadget.}
+		\caption{Example of three variable gadgets and a satisfied clause gadget. Dotted parts do not belong to the SPED.}
 		\label{fig:gadget_connection}
@@
 
-	Each variable gadget is a cycle of segment pairs, with at least one pair for each occurence of the variable in $\phi$, see the three variable gadgets in the bottom of Fig.~\ref{fig:gadget_connection}.
+	Each variable gadget is a cycle of segment pairs, with at least one pair for each occurence of the variable in $\phi$, see Fig.~\ref{fig:gadget_connection}. 
 	Observe that this cycle has exactly two ink-maximal SPEDs: either all red edges are full segments and all green edges are length-1 stubs or vice versa.
@@
 
-	For each clause we construct a triple of mutually intersecting segments, see the black triangle at the top of Fig.~\ref{fig:gadget_connection}.
+	For each clause we construct a triple of mutually intersecting segments, see the gadget on yellow background in the upper part of Fig.~\ref{fig:gadget_connection}.
 	Again, their intersection points are at distance 1 from the endpoints.
@@
 	We draw the clause segment that connects to the selected literal as a full segment and the other two as length-1 stubs.
+	Recall that the literal paths are oriented from the variable gadget to the clause gadget.
 	Since the last segment of the selected literal path must be drawn as length-1 stubs, the only way of having a maximum contribution of that path is by alternating stubs and full segments.

### 2018-01-09 eurocg/paper.tex
smaller illustration for tree
deleted conclusion
trees fertig

@@
 Let $ u \in V $ and $ \deg(u) = \delta $.
-Then for the corresponding segment $ s(u) \in S $ there are $ \delta + 1 $ possible stub pairs including drawing the whole segment.
+Then for the corresponding segment $ s(u) \in S $ there are $ \delta + 1 $ possible stub pairs including drawing the whole segment, see Fig.~\ref{fig:tree}.
 We store the lengths of the possible stub pairs as node weights $ \ell_1(u), \dots, \ell_\delta(u) \in \mathbb{R}_+$ and assume that $ \ell_1(u),\dots,\ell_\delta(u) $ are sorted from shorter to longer stubs.
 We define  $ \ell_0(u) $ as the length of the whole segment $ s(u) $.
+
+\begin{figure}[tbp]
+	\centering
+ 	\includegraphics{tree_illustration}
+	\caption{A segment with  intersecting segments and the different induced stub lengths.} % The shaded areas are the stubs, $ p(u) $ is the parent of $ u $.
+The stubs in the boxed area contribute to $ \solshort(u) $.}
+	\label{fig:tree}
+\end{figure}
 
@@
 \label{sec:tree}
-\begin{figure}[tbp]
-	\centering
- 	\includegraphics{tree_illustration}
-	\caption{One edge $ u $ with intersecting segments. The shaded areas are the stubs, $ p(u) $ is the parent of $ u $. The stubs in the boxed area contribute to $ \solshort(u) $.}
-	\label{fig:tree}
-\end{figure}
+
 Here we assume that $ C =(V,E) $ is a rooted tree of maximum degree $ k $. 
@@
 %\end{align*}
-Based on these two values we can recursively define $ T_i(u) $: % for a vertex $ u $ and stub length $ \ell_i(u) $:
+Figure~\ref{fig:tree} highlights the stub lengths that contribute to $\solshort(u)$. We recursively define % for a vertex $ u $ and stub length $ \ell_i(u) $:
 \begin{align}
@@
 		\solshort(v) & \text{if } s(u) \text{ with length } \ell_i(u) \text{ intersects } s(v) \\
-		\sollong(v) & \text{else}
-	\end{cases}
+		\sollong(v) & \text{otherwise}
+	\end{cases}.
 \end{align}
@@
 
+So by solving Recurrence~(\ref{rec:tree}) in $O(mk)$ time we find an optimal solution to the \maxsped problem on $ G $ with drawing $ \Gamma $ using standard backtracking from $\max\{T_0(r), \dots, T_{\deg(r)}(r)\}$ for the root $r$ of $C$.
+Since the intersection graph $C$ is a tree with $O(m)$ edges it can be computed in $O(m \log m)$ time and we obtain the following theorem.
+
 \begin{theorem}
@@
 	Let $ G $ be a simple graph with $m$ edges and $ \Gamma $ a straight-line drawing of $ G $.
-If the intersection graph $ C = (V,E) $ of $ G $ is a tree with maximum degree $ k \in \mathbb{N} $ the \maxsped problem can be solved in $ O(mk + m \log m)) $ time and space.
+If the intersection graph $ C = (V,E) $ of $ G $ is a tree with maximum degree $ k \in \mathbb{N} $ problem \maxsped can be solved in $ O(mk + m \log m)) $ time and space.
 \end{theorem}
-\begin{proof}
-	Since $ C $ is known to be a tree, and hence $ |E| = O(m) $, we can compute the intersection graph and the stub lengths in $ O(m\log m) $ time and space.
-	With Lemma~\ref{lem:compute_vertex} and Recurrence~\ref{rec:tree} we need one pass through the tree to find the roots maximum in $ O(mk) $ time.
+% \begin{proof}
+% 	Since $ C $ is known to be a tree, and hence $ |E| = O(m) $, we can compute the intersection graph and the stub lengths in $ O(m\log m) $ time and space.
+% 	With Lemma~\ref{lem:compute_vertex} and Recurrence~\ref{rec:tree} we need one pass through the tree to find the roots maximum in $ O(mk) $ time.
 Adding the bounds yields the desired runtime of $ O(mk + m \log m)) $.
-%	To solve the \maxsped problem on $ G $ we need to compute $ \max \{T_i(r) \mid 0 \leq i \leq k \} $ for the root $ r \in V $ of $ C $ as seen above.
+% %	To solve the \maxsped problem on $ G $ we need to compute $ \max \{T_i(r) \mid 0 \leq i \leq k \} $ for the root $ r \in V $ of $ C $ as seen above.
 Doing this bottom up requires one pass through $ C $ hence by applying Lemma~\ref{lem:compute_vertex} we need $ O(mk) $ time to compute all $ T_i(u) $ for all $ u \in V $.
@@
 Adding the bounds yields the desired runtime of $ O(mk + m \log m)) $.
+% %
 %
-
-By solving Recurrence~(\ref{rec:tree}) we find an optimal solution to the \maxsped problem on $ G $ with drawing $ \Gamma $ as $\max\{T_0(r), \dots, T_{\deg(r)}(r)\}$ for the root $r$ of $C$.
-
-	Retrieving the actual stub length for each edge in $ G $ needed to produce the SPED can be done by backtracking the subsolutions which requires at most $ O(mk) $ time.
-\end{proof}
+%
+%
+% 	Retrieving the actual stub length for each edge in $ G $ needed to produce the SPED can be done by backtracking the subsolutions which requires at most $ O(mk) $ time.
+% \end{proof}
 

### 2018-01-09 eurocg/paper.tex
4.1 done

@@
  	\includegraphics{tree_illustration}
-	\caption{A segment with  intersecting segments and the different induced stub lengths.} % The shaded areas are the stubs, $ p(u) $ is the parent of $ u $.
+	\caption{A segment $s(u)$ with five intersecting segments and the different induced stub lengths.
+The boxed stub lengths are considered in $\solshort(u)$.} % The shaded areas are the stubs, $ p(u) $ is the parent of $ u $.
 The stubs in the boxed area contribute to $ \solshort(u) $.}
@@
 \label{sec:tree}
+
+% \begin{figure}[tbp]
+% 	\centering
+%  	\includegraphics{tree_illustration}
+% 	\caption{One edge $ s(u) $ with intersecting segments. The shaded areas are the stubs, $ p(u) $ is the parent of $ u $. The stubs in the boxed area contribute to $ \solshort(u) $.}
+% 	\label{fig:tree}
+% \end{figure}
 
@@
 
-\section{Conclusion}
-
 \bibliography{paper}
-
 \end{document}

### 2018-01-09 eurocg/paper.tex
zwischenstand

@@
 
-\martin{$f_s$ mit $\ell_i$ linken}
+For a vertex $u \in V$ and a neighbor $v$ in $C$ there is a well-defined stub length $\ell$ of $u$
 
@@
 
-\begin{proof}
+%\begin{proof}
 	We reduce from the \NP-hard problem \ppsat~\cite{l-pftu-82} using similar ideas as in Bruckdorfer's general proof of the hardness of \maxsped~\cite{b-sgh-15}.
@@
 		\begin{figure}
+%	\begin{wrapfigure}[15]{r}{0.55\textwidth}
 			\centering
@@
 			\label{fig:gadget_connection}
+%	\end{wrapfigure}		
 		\end{figure}
-
 	All segments in the gadgets are of length 5.
@@
 	But then by construction the selected variable value sets our literal to \emph{true}.
-\end{proof}
+%\end{proof}
 
@@
 Section~\ref{sec:hardness} showed that \maxsped is generally \NP-hard for $k \ge 3$. 
-In this section we consider the special case that the intersection graph of the $k$-plane input drawing is a tree or a cactus.
+Now we consider the special case that the intersection graph of the $k$-plane input drawing is a tree or a cactus.
 In both cases we present polynomial-time dynamic programming algorithms.
@@
 Let $ u \in V $ and $ \deg(u) = \delta $. 
-Then for the corresponding segment $ s(u) \in S $ there are $ \delta + 1 $ possible stub pairs including drawing the whole segment, see Fig.~\ref{fig:tree}.
-We store the lengths of the possible stub pairs as node weights $ \ell_1(u), \dots, \ell_\delta(u) \in \mathbb{R}_+$ and assume that $ \ell_1(u),\dots,\ell_\delta(u) $ are sorted from shorter to longer stubs.
+Then for the corresponding segment $ s(u) \in S $ there are $ \delta + 1 $ relevant stub pairs including the whole segment, see Fig.~\ref{fig:tree}. 
+Let $ \ell_1(u), \dots, \ell_\delta(u) \in \mathbb{R}_+$ be the stub lengths induced by the intersection points of $s(u)$ with the segments of the neighbors of $u$,  sorted from shorter to longer stubs.
 We define  $ \ell_0(u) $ as the length of the whole segment $ s(u) $.
@@
 	\caption{A segment $s(u)$ with five intersecting segments and the different induced stub lengths.
-The boxed stub lengths are considered in $\solshort(u)$.} % The shaded areas are the stubs, $ p(u) $ is the parent of $ u $.
+The boxed stub lengths are considered in $\solshort(u)$ and do not affect $p(u)$.} % The shaded areas are the stubs, $ p(u) $ is the parent of $ u $.
 The stubs in the boxed area contribute to $ \solshort(u) $.}
@@
 %\end{align*}
-Figure~\ref{fig:tree} highlights the stub lengths that contribute to $\solshort(u)$. We recursively define % for a vertex $ u $ and stub length $ \ell_i(u) $:
+Figure~\ref{fig:tree} highlights the stub lengths that are considered in $\solshort(u)$. We recursively define % for a vertex $ u $ and stub length $ \ell_i(u) $:
 \begin{align}
@@
 
-Recurrence~(\ref{rec:tree}) can be solved naively in $ O(mk^2) $ time. 
+Recurrence~(\ref{rec:tree}) can be solved naively in $ O(mk^2) $ time, where $m = |V|$. 
 Using the order on the stub lengths  we can improve this to $ O(mk) $ time by computing all $ T_i(u) $ for one $ u \in V $ in $O(k)$ time.

### 2018-01-09 eurocg/paper.tex
verwaisten satz entfernt
merge

--- 2018-01-09 eurocg/paper.tex
@@
 Let $ u \in V $ and $ \deg(u) = \delta $.
-Then for the corresponding segment $ s(u) \in S $ there are $ \delta + 1 $ possible stub pairs including drawing the whole segment, see Fig.~\ref{fig:tree}.
+Then for the corresponding segment $ s(u) \in S $ there are $ \delta + 1 $ possible stub pairs including drawing the whole segment.
 We store the lengths of the possible stub pairs as node weights $ \ell_1(u), \dots, \ell_\delta(u) \in \mathbb{R}_+$ and assume that $ \ell_1(u),\dots,\ell_\delta(u) $ are sorted from shorter to longer stubs.
 We define  $ \ell_0(u) $ as the length of the whole segment $ s(u) $.
-
-\begin{figure}[tbp]
-	\centering
- 	\includegraphics{tree_illustration}
-	\caption{A segment with  intersecting segments and the different induced stub lengths.} % The shaded areas are the stubs, $ p(u) $ is the parent of $ u $.
-The stubs in the boxed area contribute to $ \solshort(u) $.}
-	\label{fig:tree}
-\end{figure}
 
@@
 \label{sec:tree}
-
+\begin{figure}[tbp]
+	\centering
+ 	\includegraphics{tree_illustration}
+	\caption{One edge $ s(u) $ with intersecting segments. The shaded areas are the stubs, $ p(u) $ is the parent of $ u $. The stubs in the boxed area contribute to $ \solshort(u) $.}
+	\label{fig:tree}
+\end{figure}
 Here we assume that $ C =(V,E) $ is a rooted tree of maximum degree $ k $. 
@@
 %\end{align*}
-Figure~\ref{fig:tree} highlights the stub lengths that contribute to $\solshort(u)$. We recursively define % for a vertex $ u $ and stub length $ \ell_i(u) $:
+Based on these two values we can recursively define $ T_i(u) $: % for a vertex $ u $ and stub length $ \ell_i(u) $:
 \begin{align}
@@
 		\solshort(v) & \text{if } s(u) \text{ with length } \ell_i(u) \text{ intersects } s(v) \\
-		\sollong(v) & \text{otherwise}
-	\end{cases}.
+		\sollong(v) & \text{else}
+	\end{cases}
 \end{align}
@@
 
-So by solving Recurrence~(\ref{rec:tree}) in $O(mk)$ time we find an optimal solution to the \maxsped problem on $ G $ with drawing $ \Gamma $ using standard backtracking from $\max\{T_0(r), \dots, T_{\deg(r)}(r)\}$ for the root $r$ of $C$.
-Since the intersection graph $C$ is a tree with $O(m)$ edges it can be computed in $O(m \log m)$ time and we obtain the following theorem.
-
 \begin{theorem}
@@
 	Let $ G $ be a simple graph with $m$ edges and $ \Gamma $ a straight-line drawing of $ G $.
-If the intersection graph $ C = (V,E) $ of $ G $ is a tree with maximum degree $ k \in \mathbb{N} $ problem \maxsped can be solved in $ O(mk + m \log m)) $ time and space.
+If the intersection graph $ C = (V,E) $ of $ G $ is a tree with maximum degree $ k \in \mathbb{N} $ the \maxsped problem can be solved in $ O(mk + m \log m)) $ time and space.
 \end{theorem}
-% \begin{proof}
-% 	Since $ C $ is known to be a tree, and hence $ |E| = O(m) $, we can compute the intersection graph and the stub lengths in $ O(m\log m) $ time and space.
-% 	With Lemma~\ref{lem:compute_vertex} and Recurrence~\ref{rec:tree} we need one pass through the tree to find the roots maximum in $ O(mk) $ time.
+\begin{proof}
+	Since $ C $ is known to be a tree, and hence $ |E| = O(m) $, we can compute the intersection graph and the stub lengths in $ O(m\log m) $ time and space.
+	With Lemma~\ref{lem:compute_vertex} and Recurrence~\ref{rec:tree} we need one pass through the tree to find the roots maximum in $ O(mk) $ time.
 Adding the bounds yields the desired runtime of $ O(mk + m \log m)) $.
-% %	To solve the \maxsped problem on $ G $ we need to compute $ \max \{T_i(r) \mid 0 \leq i \leq k \} $ for the root $ r \in V $ of $ C $ as seen above.
+%	To solve the \maxsped problem on $ G $ we need to compute $ \max \{T_i(r) \mid 0 \leq i \leq k \} $ for the root $ r \in V $ of $ C $ as seen above.
 Doing this bottom up requires one pass through $ C $ hence by applying Lemma~\ref{lem:compute_vertex} we need $ O(mk) $ time to compute all $ T_i(u) $ for all $ u \in V $.
@@
 Adding the bounds yields the desired runtime of $ O(mk + m \log m)) $.
-% %
 %
-%
-%
-% 	Retrieving the actual stub length for each edge in $ G $ needed to produce the SPED can be done by backtracking the subsolutions which requires at most $ O(mk) $ time.
-% \end{proof}
+
+By solving Recurrence~(\ref{rec:tree}) we find an optimal solution to the \maxsped problem on $ G $ with drawing $ \Gamma $ as $\max\{T_0(r), \dots, T_{\deg(r)}(r)\}$ for the root $r$ of $C$.
+
+	Retrieving the actual stub length for each edge in $ G $ needed to produce the SPED can be done by backtracking the subsolutions which requires at most $ O(mk) $ time.
+\end{proof}
 
@@
 \begin{align*}
-	\solshort(u) &= \max\{T_0(u),\dots,T_\delta(u)\}\\
-	\solmid(u) &= \max\{T_1(u),\dots,T_{a}(u)\}\\
-	\sollong(u) &= \max\{T_1(u),\dots,T_{b}(u)\}
+	\solshort(u) = \max\{T_0(u),\dots,T_\delta(u)\}\\
+	\solmid(u) = \max\{T_1(u),\dots,T_{a}(u)\}\\
+	\sollong(u) = \max\{T_1(u),\dots,T_{b}(u)\}
 \end{align*}

### 2018-01-09 eurocg/paper.tex
dumb cactus

--- 2018-01-09 eurocg/paper.tex
@@
 In the problem \maxsped the task is to find for a given drawing~$\Gamma$ a SPED $D^*$ such that its ink $I(D^*)$ is maximum over all SPEDs.
-
-For a vertex $u \in V$ and a neighbor $v$ in $C$ there is a well-defined stub length $\ell$ of $u$
 
@@
 a simple graph in which no edge $ e \in E $ is part of more than one cycle.
-For a cactus graph $ C $ the \emph{block-cut tree} is the tree $ \mathcal T = (\mathcal V, \mathcal E) $where $ \mathcal V $ contains a vertex for every cycle and \emph{cut vertex} in $ C $.
-We call a vertex $ u \in \mathcal V $ a \emph{block vertex} if it represents a cycle and \emph{articulation vertex} else.
-For a block vertex $ u \in \mathcal V $ representing an induced cycle with vertices $ V_c \subseteq V $ we define $ \textit{cycle}(u) = V_c$.
-For an articulation vertex $ u \in \mathcal V $ representing the cut vertex $ v \in V $ we define $ A(u) = v $.
+For a cactus graph $ C $ the \emph{block-cut tree} is the tree $ \mathcal T = (\mathcal V, \mathcal E) $ where $ \mathcal V $ contains a vertex for every cycle and cut vertex in $ C $.
+We call a vertex $ u \in \mathcal V $ a \emph{block vertex} if it represents an induced cycle with vertices $ V_c \subseteq V $ and \emph{articulation vertex} of the cut vertex $ v \in V $ else.
+For a block vertex $ u \in \mathcal V $ we define $ \textit{cycle}(u) = V_c$ and for an articulation vertex $ u \in \mathcal V $ we define $ A(u) = v $.
 There is an edge $ uv \in \mathcal E $ between a block vertex $ u $ and articulation vertex $ v $ whenever $ v \in \textit{cycle}(u) $ and an edge $ uv \in \mathcal E $ between two articulation vertices whenever $ A(u)A(v) $ was an edge in $ E $.
@@
 Let $ C = (V,E) $ be a cactus graph and moreover the intersection graph of a graph $ G $ and straight-line drawing $ \Gamma $.
-Additionally let $ \mathcal T = (\mathcal V, \mathcal E) $ be the block-cut tree of $ C $ we define for each vertex $ u \in \mathcal V $ two values $ a,b \in \mathbb{R}_+ $, w.l.o.g $ a \leq b $.
-If $ u $ is an articulation vertex this are exactly the positions where the cycle represented by $ p(u) $ intersects $ s(A(u)) $ in $ \Gamma $.
-If $ u $ is a block vertex $ a,b $ are the points where edges $ s(v), v \in \textit{cycle}(u) $ intersect $ s(A(p(u))) $.
+Additionally let $ \mathcal T = (\mathcal V, \mathcal E) $ be the block-cut tree of $ C $ we define for each vertex $ u \in \mathcal V $ two indices $ i_p,i_q $, w.l.o.g $ i_p > i_q $.
+This are exactly the indices of the stub-lengths leading to an intersection between two cycles of segments in $ \Gamma $.
 Similar to Section~\ref{sec:tree} we define three values for every vertex $ u \in \mathcal V $ with degree $ \delta $:
 \begin{align*}
-	\solshort(u) &= \max\{T_0(u),\dots,T_\delta(u)\}\\
-	\solmid(u) &= \max\{T_1(u),\dots,T_{a}(u)\}\\
-	\sollong(u) &= \max\{T_1(u),\dots,T_{b}(u)\}
+	\solshort(u) = \max\{T_0(u),\dots,T_\delta(u)\}\\
+	\solmid(u) = \max\{T_1(u),\dots,T_{i_p}(u)\}\\
+	\sollong(u) = \max\{T_1(u),\dots,T_{i_q}(u)\}
 \end{align*}
-For an articulation vertex $ u \in \mathcal V $:
+For an articulation vertex $ u \in \mathcal V $ we can define a similar recurrence to Recurrence~\ref{rec:tree}:
 \begin{align}
@@
 \end{align}
-	Let $ G $, $ \Gamma $,  $ C = (V,E) $ and $ \mathcal T = (\mathcal V,\mathcal E) $ as above and $ u \in \mathcal V $ a leaf of $ \mathcal T$.
+For a block vertex $ u \in \mathcal V $ we have to cut open the cycle it represents (compare \cite{bcgkmn-pped-17}).
+We do this at $ A(p(u)) $.
+The resulting graph is a path over the vertices in $ \textit{cycle}(u) $ with attached articulation vertices.
+This is especially a tree and we use the aglorithm from Section~\ref{sec:tree}, maximizing over three values of the articulation vertices though,  to compute the $ T_i(u) $.
+
+The correctness follows as in the case of a tree in Section~\ref{sec:tree} if the values computed for the block vertices is correct.
+This is the case since we only extended the cycle by attaching some leafs, which can be dealt with after cutting the cycle open as in \cite{bcgkmn-pped-17}.
+For the runtime we can again do the same as in Section~\ref{sec:tree}, but this time we have to also check cycles of length $ n_c $ which leads to a runtime of $ O(n_c\delta) $ for a block vertex $ u \in \mathcal V $ with degree $ \delta $ and $ n_c = |\textit{cycle}(u)| $.
+In case $ u $ is an articulation vertex we can solve it in $ O(\delta) $.
+As before the optimal solution to the \maxsped problem on $ G $ with drawing $ \Gamma $ can be found by backtracking from the root.
+Additionally the intersection graph can be computed in $ O(m\log m) $ where $ m $ is the number of edges in the original graph $ G $ and we obtain the following theorem.
+%$ G $, $ \Gamma $,  $ C = (V,E) $ and $ \mathcal T = (\mathcal V,\mathcal E) $ as above and $ u \in \mathcal V $ a leaf of $ \mathcal T$.
 Then either $ u $ is an articulation vertex and to be treated as in Lemma~\ref{lem:compute_vertex} or a cycle.
@@
 Using the above described method we derive $ \solshort(u) $, $ \solmid(u) $ and $ \sollong(u) $, this are exactly the optimal values depending on how the cycle interacts with the articulation vertex $ p(u) $.
-
-Now let $ u \in V $ with degree $ \delta $ be an inner vertex of $ \mathcal T $.
+%
+%Now let $ u \in V $ with degree $ \delta $ be an inner vertex of $ \mathcal T $.
 The values $ \solshort(v) $, $ \solmid(v) $ and $ \sollong(v) $ are computed for all $ v \in c(u) $.
@@
 Now let $ u $ be a block vertex.
-We have to cut open the cycle it represents (compare \cite{bcgkmn-pped-17}).
-We do this at $ p = A(p(u)) $.
-The resulting graph is a path over the vertices in $ \textit{cycle}(u) $ with attached articulation vertices.
-This is especially a tree and we use the aglorithm from Section~\ref{sec:tree}, maximizing over three values of the articulation vertices though,  to compute the $ T_i(u) $.
 %\begin{lemma}
@@
 %\end{proof}
-\begin{lemma}
-	\label{lem:compute_block}
-	Let $ \mathcal T = (\mathcal V,\mathcal E) $ be a cactus with maximum degree $ k $ and vertex weights $ \ell_0,\dots,\ell_{\delta} $ for every block vertex $ u \in \mathcal V $ with degree $ \delta $.
+%\begin{lemma}
+%	\label{lem:compute_block}
+%	Let $ \mathcal T = (\mathcal V,\mathcal E) $ be a cactus with maximum degree $ k $ and vertex weights $ \ell_0,\dots,\ell_{\delta} $ for every block vertex $ u \in \mathcal V $ with degree $ \delta $.
 We can compute $ \solshort(u) $, $ \solmid(u) $ and $ \sollong(u) $ in $ O(n_c\delta) $ time and space, if the values $ \sollong(v) $, $ \solmid(v) $ and $ \solshort(v) $ are computed for all children $ v \in c(u) $.
-\end{lemma}
-\begin{proof}
-	Let $ G $, $ \Gamma $,  $ C = (V,E) $ with maximum degree $ k $ and $ \mathcal T = (\mathcal V,\mathcal E) $ as above.
+%\end{lemma}
+%\begin{proof}
+%	Let $ G $, $ \Gamma $,  $ C = (V,E) $ with maximum degree $ k $ and $ \mathcal T = (\mathcal V,\mathcal E) $ as above.
 Let $ u \in \mathcal V $ be a block vertex.
@@
 Following we need $ O(n_ck) $ to compute the three values $ \solshort(u) $, $ \solmid(u) $ and $ \sollong(u) $.
-\end{proof}
+%\end{proof}
 
@@
 \end{theorem}
-\begin{proof}
-	Follows as in Theorem~\ref{thm:tree} by maximizing over the tree using Recurrence~\ref{rec:cactus_articulation}.
+%\begin{proof}
+%	Follows as in Theorem~\ref{thm:tree} by maximizing over the tree using Recurrence~\ref{rec:cactus_articulation}.
 Then the running time follows from constructing the intersection graph and the block-cut tree as well as Lemma~\ref{lem:compute_block} for the block vertices and applying Lemma~\ref{lem:compute_vertex} for the articulation vertices.
-\end{proof}
+%\end{proof}
 

### 2018-01-09 eurocg/paper.tex
minor sec 3 changes

@@
 \usepackage{graphicx}
+%\usepackage{cutwin}
 
@@
 
-%\begin{proof}
+%\renewcommand\windowpagestuff{%
+%	\includegraphics{Reduction_visual}
+%	\captionof{figure}{Example of three variable gadgets and\\ a satisfied clause gadget. Dotted parts do not belong\\ to the SPED.}
+%	\label{fig:gadget_connection}
+%}
+%\opencutleft
+\begin{proof}
+%		\begin{cutout}{2}{0pt}{0.44\textwidth}{15}
 	We reduce from the \NP-hard problem \ppsat~\cite{l-pftu-82} using similar ideas as in Bruckdorfer's general proof of the hardness of \maxsped~\cite{b-sgh-15}.
@@
 	Figure~\ref{fig:gadget_connection} shows a sketch of the gadgets.
+%		\end{cutout}
 
@@
 %	\begin{wrapfigure}[15]{r}{0.55\textwidth}
+
 			\centering
@@
 	Observe that this cycle has exactly two ink-maximal SPEDs: either all red edges are full segments and all green edges are length-1 stubs or vice versa.
-	We associate the configuration with full red segments and green stubs with the value \emph{true} and the configuration with full green segments and red stubs with the value \emph{false}.
+	We associate the configuration with green stubs and full red segments with the value \emph{true} and the configuration with full green segments and red stubs with the value \emph{false}.
 
@@
 	But then by construction the selected variable value sets our literal to \emph{true}.	
-%\end{proof}
+\end{proof}
 
@@
 Let $ C = (V,E) $ be the intersection graph of a given drawing $\Gamma$ of graph $ G $ as defined in Section~\ref{sec:preliminaries}.
-Let $ u \in V $ and $ \deg(u) = \delta $. 
+Let $ u \in V $ and $ \delta  = \deg(u) $. 
 Then for the corresponding segment $ s(u) \in S $ there are $ \delta + 1 $ relevant stub pairs including the whole segment, see Fig.~\ref{fig:tree}. 
@@
 	\solshort(u) = \max\{T_1(u),\dots,T_{i_p}(u)\} \qquad
-	\sollong(u) = \max\{T_0(u),\dots,T_{\delta}(u)\}
+	\sollong(u) = \max\{T_0(u),\dots,T_{\delta}(u)\}.
 \]
@@
 	Let $ G $ be a simple graph with $m$ edges and $ \Gamma $ a straight-line drawing of $ G $.
-If the intersection graph $ C = (V,E) $ of $ G $ is a tree with maximum degree $ k \in \mathbb{N} $ problem \maxsped can be solved in $ O(mk + m \log m)) $ time and space.
+If the intersection graph $ C = (V,E) $ of $ G $ is a tree with maximum degree $ k \in \mathbb{N} $ problem \maxsped can be solved in $ O(mk + m \log m) $ time and space.
 \end{theorem}
@@
 For a cactus graph $ C $ the \emph{block-cut tree} is the tree $ \mathcal T = (\mathcal V, \mathcal E) $ where $ \mathcal V $ contains a vertex for every cycle and cut vertex in $ C $.
-We call a vertex $ u \in \mathcal V $ a \emph{block vertex} if it represents an induced cycle with vertices $ V_c \subseteq V $ and \emph{articulation vertex} of the cut vertex $ v \in V $ else.
+We call a vertex $ u \in \mathcal V $ a \emph{block vertex} if it represents an induced cycle with vertices $ V_c \subseteq V $ and \emph{articulation vertex} representing the cut vertex $ v \in V $ else.
 For a block vertex $ u \in \mathcal V $ we define $ \textit{cycle}(u) = V_c$ and for an articulation vertex $ u \in \mathcal V $ we define $ A(u) = v $.
@@
 
-Let $ C = (V,E) $ be a cactus graph and moreover the intersection graph of a graph $ G $ and straight-line drawing $ \Gamma $.
+Let $ C = (V,E) $ be a cactus graph and moreover the intersection graph of a graph $ G $ with straight-line drawing $ \Gamma $.
 Additionally let $ \mathcal T = (\mathcal V, \mathcal E) $ be the block-cut tree of $ C $ we define for each vertex $ u \in \mathcal V $ two indices $ i_p,i_q $, w.l.o.g $ i_p > i_q $.
@@
 \begin{align*}
-	\solshort(u) = \max\{T_0(u),\dots,T_\delta(u)\}\\
-	\solmid(u) = \max\{T_1(u),\dots,T_{i_p}(u)\}\\
-	\sollong(u) = \max\{T_1(u),\dots,T_{i_q}(u)\}
+	\solshort(u) &= \max\{T_0(u),\dots,T_\delta(u)\}\\
+	\solmid(u) &= \max\{T_1(u),\dots,T_{i_p}(u)\}\\
+	\sollong(u) &= \max\{T_1(u),\dots,T_{i_q}(u)\}
 \end{align*}
+%\begin{eqnarray*}
+%	\solshort(u) = \max\{T_0(u),\dots,T_\delta(u)\}\qquad
+%	\solmid(u) = \max\{T_1(u),\dots,T_{i_p}(u)\}\\
+%	\multicolumn{2}{c}{$\sollong(u) = \max\{T_1(u),\dots,T_{i_q}(u)\}$}
+%\end{eqnarray*}
 For an articulation vertex $ u \in \mathcal V $ we can define a similar recurrence to Recurrence~\ref{rec:tree}:
@@
 \sollong(v) & \text{else}
-\end{cases}
+\end{cases}.
 \end{align}
@@
 We do this at $ A(p(u)) $.
-The resulting graph is a path over the vertices in $ \textit{cycle}(u) $ with attached articulation vertices.
+The resulting graph is a path between the vertices in $ \textit{cycle}(u) $ with attached articulation vertices.
 This is especially a tree and we use the aglorithm from Section~\ref{sec:tree}, maximizing over three values of the articulation vertices though,  to compute the $ T_i(u) $.
 
-The correctness follows as in the case of a tree in Section~\ref{sec:tree} if the values computed for the block vertices is correct.
-This is the case since we only extended the cycle by attaching some leafs, which can be dealt with after cutting the cycle open as in \cite{bcgkmn-pped-17}.
-For the runtime we can again do the same as in Section~\ref{sec:tree}, but this time we have to also check cycles of length $ n_c $ which leads to a runtime of $ O(n_c\delta) $ for a block vertex $ u \in \mathcal V $ with degree $ \delta $ and $ n_c = |\textit{cycle}(u)| $.
-In case $ u $ is an articulation vertex we can solve it in $ O(\delta) $.
+If the values computed for the block vertices is correct, we can follow the correctness of the algorithm as in Section~\ref{sec:tree}.
+Let $ u \in \mathcal V $ be a block vertex then by \cite{bcgkmn-pped-17} we know the optimal solution for \maxsped just on $ C[\textit{cycle}(u)] $ is correctly computed by the above procedure since the algorithm from Section~\ref{sec:tree} correctly handles the created path.
+Now if $ u $ has articulation vertices attached to it the above computes the correct value since this is simply a tree.
+Further there is no $ v \in c(u) $ which is not an articulation vertex, hence the maximum ink value for $ u $ is correctly computed if we know the values for the children in $ c(u) $.
+
+For the runtime we can again argue the same as in Section~\ref{sec:tree}, but this time we have to also check cycles of length $ n_c $ which leads to a runtime of $ O(n_ck^2) $ for a block vertex $ u \in \mathcal V $ with $ n_c = |\textit{cycle}(u)| $.
+In case $ u $ is an articulation vertex we can solve it in $ O(k) $.
 As before the optimal solution to the \maxsped problem on $ G $ with drawing $ \Gamma $ can be found by backtracking from the root.
-Additionally the intersection graph can be computed in $ O(m\log m) $ where $ m $ is the number of edges in the original graph $ G $ and we obtain the following theorem.
+Additionally the intersection graph can be computed in $ O(m\log m) $ where $ m $ is the number of edges in the original graph $ G $ since $ |E| = O(m) $ in a cactus and we obtain the following theorem.
 %$ G $, $ \Gamma $,  $ C = (V,E) $ and $ \mathcal T = (\mathcal V,\mathcal E) $ as above and $ u \in \mathcal V $ a leaf of $ \mathcal T$.
@@
 	Let $ G $ be a simple graph with $m$ edges and $ \Gamma $ a straight-line drawing of $ G $.
-If the intersection graph $ C = (V,E) $ of $ G $ is a cactus with maximum degree $ k \in \mathbb{N} $ the \maxsped problem can be solved in $ O(mk + m \log m)) $ time and space.
+If the intersection graph $ C = (V,E) $ of $ G $ is a cactus with maximum degree $ k \in \mathbb{N} $ the \maxsped problem can be solved in $ O(mk^2 + m \log m) $ time and space.
 \end{theorem}
@@
 
-\bibliography{paper}
+%\bibliography{paper}
+\begin{thebibliography}{1}
+
+	\bibitem{bew-vnd-95}
+	Richard~A. Becker, Stephen~G. Eick, and Allan~R. Wilks.
+	\newblock Visualizing network data.
+	\newblock {\em {IEEE} Trans. Visualization and Computer Graphics}, 1(1):16--28,
+	1995.
+
+	\bibitem{blmt-pedhmitc-16}
+	Carla Binucci, Giuseppe Liotta, Fabrizio Montecchiani, and Alessandra Tappini.
+	\newblock Partial edge drawing: Homogeneity is more important than crossings
+	and ink.
+	\newblock In {\em Information, Intelligence, Systems Applications (IISA'16)},
+	pages 1--6. IEEE, 2016.
+
+	\bibitem{b-sgh-15}
+	Till Bruckdorfer.
+	\newblock {\em Schematics of Graphs and Hypergraphs}.
+	\newblock PhD thesis, Uni Tübingen, 2015.
+
+	\bibitem{bcgkmn-pped-17}
+	Till Bruckdorfer, Sabine Cornelsen, Carsten Gutwenger, Michael Kaufmann,
+	Fabrizio Montecchiani, Martin Nöllenburg, and Alexander Wolff.
+	\newblock Progress on partial edge drawings.
+	\newblock {\em J. Graph Algorithms Appl.}, 21(4):757--786, 2017.
+
+	\bibitem{bk-ecbe-12}
+	Till Bruckdorfer and Michael Kaufmann.
+	\newblock Mad at edge crossings? {Break} the edges!
+	\newblock In {\em Fun with Algorithms (FUN'12)}, volume 7288 of {\em LNCS},
+	pages 40--50. Springer, 2012.
+
+	\bibitem{bkl-us-15}
+	Till Bruckdorfer, Michael Kaufmann, and Simon Leibßle.
+	\newblock Ped user study.
+	\newblock In {\em Graph Drawing (GD'15)}, volume 9411 of {\em LNCS}, pages
+	551--553. Springer International Publishing, 2015.
+
+	\bibitem{bvkw-epdldge-12}
+	Michael Burch, Corinna Vehlow, Natalia Konevtsova, and Daniel Weiskopf.
+	\newblock Evaluating partially drawn links for directed graph edges.
+	\newblock In {\em Graph Drawing (GD'11)}, volume 7034 of {\em LNCS}, pages
+	226--237. Springer, 2012.
+
+	\bibitem{bcko-cgaa-08}
+	Mark de~Berg, Otfried Cheong, Marc van Kreveld, and Mark Overmars.
+	\newblock {\em Computational Geometry: Algorithms and Applications}.
+	\newblock Springer, 3rd edition edition, 2008.
+
+	\bibitem{l-pftu-82}
+	David Lichtenstein.
+	\newblock Planar formulae and their uses.
+	\newblock {\em SIAM J Comput}, 11(2):329--343, 1982.	
+\end{thebibliography}
 \end{document}
@@
 	\caption{Connection between the clause and variable gadgets, created by chains of an even number of segments.}
-	\label{fig:gadget_connection}
+%	\label{fig:gadget_connection}
 \end{figure}

### 2018-01-09 eurocg/paper.tex
cactus update

--- 2018-01-09 eurocg/paper.tex
@@
 	Let $ G $ be a simple graph with $m$ edges and $ \Gamma $ a straight-line drawing of $ G $.
-If the intersection graph $ C = (V,E) $ of $ G $ is a tree with maximum degree $ k \in \mathbb{N} $ problem \maxsped can be solved in $ O(mk + m \log m)) $ time and space.
+If the intersection graph $ C = (V,E) $ of $ G $ is a tree with maximum degree $ k \in \mathbb{N} $, problem \maxsped can be solved in $ O(mk + m \log m)) $ time and space.
 \end{theorem}
@@
 
-\subsection{Cacti}
+\subsection{Cactus graphs}
 \label{sec:cactus}
-In this section we restrict $ C = (V,E) $ to be a cactus graph, i.e.
-a simple graph in which no edge $ e \in E $ is part of more than one cycle.
-For a cactus graph $ C $ the \emph{block-cut tree} is the tree $ \mathcal T = (\mathcal V, \mathcal E) $ where $ \mathcal V $ contains a vertex for every cycle and cut vertex in $ C $.
-We call a vertex $ u \in \mathcal V $ a \emph{block vertex} if it represents an induced cycle with vertices $ V_c \subseteq V $ and \emph{articulation vertex} of the cut vertex $ v \in V $ else.
-For a block vertex $ u \in \mathcal V $ we define $ \textit{cycle}(u) = V_c$ and for an articulation vertex $ u \in \mathcal V $ we define $ A(u) = v $.
-There is an edge $ uv \in \mathcal E $ between a block vertex $ u $ and articulation vertex $ v $ whenever $ v \in \textit{cycle}(u) $ and an edge $ uv \in \mathcal E $ between two articulation vertices whenever $ A(u)A(v) $ was an edge in $ E $.
+We now generalize our previous algorithm to cactus graphs. 
+Due to space constraints we provide only a sketch of the arguments and omit further details.
+Let the intersection graph $ C = (V,E) $ be a \emph{cactus graph}, i.e. a simple graph in which no edge $ e \in E $ is part of more than one cycle. 
+For a cactus graph $ C $ the \emph{block-cut tree} is an arbitrarily rooted tree $ \mathcal T = (\mathcal V, \mathcal E) $, where $ \mathcal V $ contains a node for every cycle and for every cut vertex in $ C $.
+We call a node $ u \in \mathcal V $ a \emph{block node} if it represents an induced cycle on vertices $ V(u) \subseteq V $ and an \emph{articulation node} of the cut vertex $ a(u) = v \in V $ otherwise.
+%For a block node $ u \in \mathcal V $ we define $ \textit{cycle}(u) = V_c$ and f
+%For an articulation node $ u \in \mathcal V $ we define $ a(u) = v $ to be the corresponding cut vertex. 
+We have an edge $ uv \in \mathcal E $ between a block node $ u $ and an articulation node $ v $ whenever $ a(v) \in V(u) $ and an edge $ uv \in \mathcal E $ between two articulation nodes whenever $ a(u)a(v) \in E$.
+% is an edge in $ E $.
+We re-use the notation $p(u)$ and $c(u)$ to refer to the parent and to the children of a node $u \in \mathcal V$.
 
-Let $ C = (V,E) $ be a cactus graph and moreover the intersection graph of a graph $ G $ and straight-line drawing $ \Gamma $.
-Additionally let $ \mathcal T = (\mathcal V, \mathcal E) $ be the block-cut tree of $ C $ we define for each vertex $ u \in \mathcal V $ two indices $ i_p,i_q $, w.l.o.g $ i_p > i_q $.
-This are exactly the indices of the stub-lengths leading to an intersection between two cycles of segments in $ \Gamma $.
-Similar to Section~\ref{sec:tree} we define three values for every vertex $ u \in \mathcal V $ with degree $ \delta $:
+%Let $ C = (V,E) $ be a cactus graph and moreover the intersection graph of a graph $ G $ and straight-line drawing $ \Gamma $.
+Additionally let $ \mathcal T = (\mathcal V, \mathcal E) $ be the block-cut tree of $ C $
+We further define for each node $ u \in \mathcal V $ two indices $ i_p(u) \le i_q(u) $. %, w.l.o.g $ i_p > i_q $. 
+For an articulation node $u$, $i_p$ and $i_q$ are the indices corresponding to the stub lengths induced by the intersection points of $s(a(u))$ and its neighbors in $C$ among the vertex set of $p(u)$.
+For a block node $u$, $i_p$ and $i_q$ are the indices corresponding to the intersection points of the parent's segment $s(a(p(u)))$ with its neighbors in $V(u)$.
+%These are exactly the indices of the stub-lengths leading to an intersection between two cycles of segments in $ \Gamma $. 
+Similar to Section~\ref{sec:tree} we now define three values for each articulation node $ u \in \mathcal V $ with degree $ \delta $:
 \begin{align*}
-	\solshort(u) = \max\{T_0(u),\dots,T_\delta(u)\}\\
-	\solmid(u) = \max\{T_1(u),\dots,T_{i_p}(u)\}\\
-	\sollong(u) = \max\{T_1(u),\dots,T_{i_q}(u)\}
+	\sollong(u) = \max\{T_0(u),\dots,T_\delta(u)\}\\
+	\solmid(u) = \max\{T_1(u),\dots,T_{i_q}(u)\}\\
+	\solshort(u) = \max\{T_1(u),\dots,T_{i_p}(u)\}
 \end{align*}
-For an articulation vertex $ u \in \mathcal V $ we can define a similar recurrence to Recurrence~\ref{rec:tree}:
+Further we define the recurrence $T_i(u)$ for all articulation nodes $u \in \mathcal V$:
 \begin{align}
 \label{rec:cactus_articulation}
-T_i(u) = \ell_i(A(u)) + \sum_{v \in c(u)}
+T_i(u) = \ell_i(a(u)) + \sum_{v \in c(u)}
 \begin{cases}
-\solshort(v) & \text{if } \ell_i(A(u)) \leq a \\
-\solmid(v) & \text{if } a < \ell_i(A(u)) \leq b \\
-\sollong(v) & \text{else}
-\end{cases}
+T_{i_p(v)}(v) & \text{if } i < i_p(v)\\ %\ell_i(a(u)) \leq a \\
+T_{i_q(v)}(v) & \text{if } i_p(v) < i \leq i_q(v) \\
+T_{0}(v) & \text{otherwise}
+% \solshort(v) & \text{if } i < i_p(v)\\ %\ell_i(a(u)) \leq a \\
+% \solmid(v) & \text{if } i_p(v) < i \leq i_q(v) \\
+% \sollong(v) & \text{otherwise}
+\end{cases}.
 \end{align}
-For a block vertex $ u \in \mathcal V $ we have to cut open the cycle it represents (compare \cite{bcgkmn-pped-17}).
-We do this at $ A(p(u)) $.
-The resulting graph is a path over the vertices in $ \textit{cycle}(u) $ with attached articulation vertices.
+For a block node $ u \in \mathcal V $ we cut open the cycle it represents at vertex $a(p(u))$ (refer to Bruckdorfer et al.~\cite{bcgkmn-pped-17} for details) and consider the three relevant stub lengths of the parent's segment $s(a(p(u)))$ induced by $i_p(u)$ and $i_q(u)$ and the whole segment.
+The resulting subgraph is a path on $V(u)$ with attached children that are articulation vertices in $\mathcal T$ so we can apply our algorithm of Section~\ref{sec:tree} with minor modifications.
+Once all values in the subgraph of block node $u$ are computed, we can derive the values $T_i(u)$ for the block node $u$.
+
+% The resulting graph is a path over the vertices in $ \textit{cycle}(u) $ with attached articulation vertices.
 This is especially a tree and we use the aglorithm from Section~\ref{sec:tree}, maximizing over three values of the articulation vertices though,  to compute the $ T_i(u) $.
 
-The correctness follows as in the case of a tree in Section~\ref{sec:tree} if the values computed for the block vertices is correct.
+Correctness can be proved by modifying the inductive arguments of Section~\ref{sec:tree} accordingly. 
+Combining the linear running time for cycles of Bruckdorfer et al.~\cite{bcgkmn-pped-17} with the running time for trees of Theorem~\ref{thm:tree} we can argue that cactus graphs can in fact be handled in the same time and space bounds as trees.
+
+% The correctness follows as in the case of a tree in Section~\ref{sec:tree} if the values computed for the block vertices is correct.
 This is the case since we only extended the cycle by attaching some leafs, which can be dealt with after cutting the cycle open as in \cite{bcgkmn-pped-17}.
@@
 	Let $ G $ be a simple graph with $m$ edges and $ \Gamma $ a straight-line drawing of $ G $.
-If the intersection graph $ C = (V,E) $ of $ G $ is a cactus with maximum degree $ k \in \mathbb{N} $ the \maxsped problem can be solved in $ O(mk + m \log m)) $ time and space.
+If the intersection graph $ C = (V,E) $ of $ G $ is a cactus with maximum degree $ k \in \mathbb{N} $, problem \maxsped can be solved in $ O(mk + m \log m)) $ time and space.
 \end{theorem}

### 2018-01-09 eurocg/paper.tex

@@
 	Let $ G $ be a simple graph with $m$ edges and $ \Gamma $ a straight-line drawing of $ G $.
-If the intersection graph $ C = (V,E) $ of $ G $ is a tree with maximum degree $ k \in \mathbb{N} $, problem \maxsped can be solved in $ O(mk + m \log m)) $ time and space.
+If the intersection graph $ C = (V,E) $ of $ G $ is a tree with maximum degree $ k \in \mathbb{N} $, problem \maxsped can be solved in $ O(mk + m \log m) $ time and space.
 \end{theorem}
@@
 \label{sec:cactus}
+
+%In this section we restrict $ C = (V,E) $ to be a cactus graph, i.e.
+a simple graph in which no edge $ e \in E $ is part of more than one cycle.
+For a cactus graph $ C $ the \emph{block-cut tree} is the tree $ \mathcal T = (\mathcal V, \mathcal E) $ where $ \mathcal V $ contains a vertex for every cycle and cut vertex in $ C $.
+We call a vertex $ u \in \mathcal V $ a \emph{block vertex} if it represents an induced cycle with vertices $ V_c \subseteq V $ and \emph{articulation vertex} representing the cut vertex $ v \in V $ else.
+For a block vertex $ u \in \mathcal V $ we define $ \textit{cycle}(u) = V_c$ and for an articulation vertex $ u \in \mathcal V $ we define $ A(u) = v $.
+There is an edge $ uv \in \mathcal E $ between a block vertex $ u $ and articulation vertex $ v $ whenever $ v \in \textit{cycle}(u) $ and an edge $ uv \in \mathcal E $ between two articulation vertices whenever $ A(u)A(v) $ was an edge in $ E $.
+%
+%Let $ C = (V,E) $ be a cactus graph and moreover the intersection graph of a graph $ G $ with straight-line drawing $ \Gamma $.
+Additionally let $ \mathcal T = (\mathcal V, \mathcal E) $ be the block-cut tree of $ C $ we define for each vertex $ u \in \mathcal V $ two indices $ i_p,i_q $, w.l.o.g $ i_p > i_q $.
+This are exactly the indices of the stub-lengths leading to an intersection between two cycles of segments in $ \Gamma $.
+Similar to Section~\ref{sec:tree} we define three values for every vertex $ u \in \mathcal V $ with degree $ \delta $:
+%\begin{align*}
+%	\solshort(u) &= \max\{T_0(u),\dots,T_\delta(u)\}\\
+%	\solmid(u) &= \max\{T_1(u),\dots,T_{i_p}(u)\}\\
+%	\sollong(u) &= \max\{T_1(u),\dots,T_{i_q}(u)\}
+%\end{align*}
+%%\begin{eqnarray*}
+%%	\solshort(u) = \max\{T_0(u),\dots,T_\delta(u)\}\qquad
+%%	\solmid(u) = \max\{T_1(u),\dots,T_{i_p}(u)\}\\
+%%	\multicolumn{2}{c}{$\sollong(u) = \max\{T_1(u),\dots,T_{i_q}(u)\}$}
+%%\end{eqnarray*}
+%For an articulation vertex $ u \in \mathcal V $ we can define a similar recurrence to Recurrence~\ref{rec:tree}:
+
 We now generalize our previous algorithm to cactus graphs. 
@@
 Further we define the recurrence $T_i(u)$ for all articulation nodes $u \in \mathcal V$:
+
 \begin{align}
@@
 \begin{cases}
+%\solshort(v) & \text{if } \ell_i(A(u)) \leq a \\
+%\solmid(v) & \text{if } a < \ell_i(A(u)) \leq b \\
+%\sollong(v) & \text{else}
+%\end{cases}.
+%\end{align}
+%For a block vertex $ u \in \mathcal V $ we have to cut open the cycle it represents (compare \cite{bcgkmn-pped-17}).
+We do this at $ A(p(u)) $.
+The resulting graph is a path between the vertices in $ \textit{cycle}(u) $ with attached articulation vertices.
+This is especially a tree and we use the aglorithm from Section~\ref{sec:tree}, maximizing over three values of the articulation vertices though,  to compute the $ T_i(u) $.
+%
+%If the values computed for the block vertices is correct, we can follow the correctness of the algorithm as in Section~\ref{sec:tree}.
+Let $ u \in \mathcal V $ be a block vertex then by \cite{bcgkmn-pped-17} we know the optimal solution for \maxsped just on $ C[\textit{cycle}(u)] $ is correctly computed by the above procedure since the algorithm from Section~\ref{sec:tree} correctly handles the created path.
+Now if $ u $ has articulation vertices attached to it the above computes the correct value since this is simply a tree.
+Further there is no $ v \in c(u) $ which is not an articulation vertex, hence the maximum ink value for $ u $ is correctly computed if we know the values for the children in $ c(u) $.
+%
+%For the runtime we can again argue the same as in Section~\ref{sec:tree}, but this time we have to also check cycles of length $ n_c $ which leads to a runtime of $ O(n_ck^2) $ for a block vertex $ u \in \mathcal V $ with $ n_c = |\textit{cycle}(u)| $.
+In case $ u $ is an articulation vertex we can solve it in $ O(k) $.
+As before the optimal solution to the \maxsped problem on $ G $ with drawing $ \Gamma $ can be found by backtracking from the root.
+Additionally the intersection graph can be computed in $ O(m\log m) $ where $ m $ is the number of edges in the original graph $ G $ since $ |E| = O(m) $ in a cactus and we obtain the following theorem.
 T_{i_p(v)}(v) & \text{if } i < i_p(v)\\ %\ell_i(a(u)) \leq a \\
@@
 Additionally the intersection graph can be computed in $ O(m\log m) $ where $ m $ is the number of edges in the original graph $ G $ and we obtain the following theorem.
+
 %$ G $, $ \Gamma $,  $ C = (V,E) $ and $ \mathcal T = (\mathcal V,\mathcal E) $ as above and $ u \in \mathcal V $ a leaf of $ \mathcal T$.
@@
 	Let $ G $ be a simple graph with $m$ edges and $ \Gamma $ a straight-line drawing of $ G $.
-If the intersection graph $ C = (V,E) $ of $ G $ is a cactus with maximum degree $ k \in \mathbb{N} $, problem \maxsped can be solved in $ O(mk + m \log m)) $ time and space.
+If the intersection graph $ C = (V,E) $ of $ G $ is a cactus with maximum degree $ k \in \mathbb{N} $, problem \maxsped can be solved in $ O(mk + m \log m) $ time and space.
 \end{theorem}

### 2018-01-09 eurocg/paper.tex
minor edits

@@
 
-\affil[1]{Algorithms and Complexity Group, TU Wien, Vienna, Austria\\ \texttt{[fklute|noellenburg]@ac.tuwien.ac.at}}
+\affil[1]{Algorithms and Complexity Group, TU Wien, Vienna, Austria\\ \texttt{[fklute|noellenburg]@ac.tuwien.ac.at, soeren.nickel@tuwien.ac.at}}
 %\affil[2]{Algorithms and Complexity Group, TU Wien, Vienna, Austria} %\texttt{fklute@ac.tuwien.ac.at}}
@@
 	\solmid(u) = \max\{T_1(u),\dots,T_{i_q}(u)\}\\
-	\solshort(u) = \max\{T_1(u),\dots,T_{i_p}(u)\}
+	\solshort(u) = \max\{T_1(u),\dots,T_{i_p}(u)\}.
 \end{align*}
 Further we define the recurrence $T_i(u)$ for all articulation nodes $u \in \mathcal V$:
-
 \begin{align}
@@
 Additionally the intersection graph can be computed in $ O(m\log m) $ where $ m $ is the number of edges in the original graph $ G $ since $ |E| = O(m) $ in a cactus and we obtain the following theorem.
-T_{i_p(v)}(v) & \text{if } i < i_p(v)\\ %\ell_i(a(u)) \leq a \\
+T_{i_p(v)}(v) & \text{if } 1 \le i < i_p(v)\\ %\ell_i(a(u)) \leq a \\
 T_{i_q(v)}(v) & \text{if } i_p(v) < i \leq i_q(v) \\

### 2018-01-09 eurocg/paper.tex
formatting
Korrekturen von Michael

@@
 We also denote the segment in $S$ corresponding to a vertex $v \in V$ by $s(v)$.
-Observe that the intersection graph $ C $ of a $ k $-plane drawing $ \Gamma $ has maximum degree $ k $.  
+Observe that the intersection graph $ C $ of a $ k $-plane drawing $ \Gamma $ has maximum degree~$ k $.  
 %
@@
 	Let $\phi$ be a planar 3-Sat formula with $n$ variables $\{x_1, \dots, x_n\}$ and $m$ clauses $\{c_1, \dots, c_m\}$, each consisting of three literals.
-	We can assume that $\phi$ comes with a planar drawing of its variable-clause graph $H_\phi$ that has a vertex for each variable $x_i$ placed on a horizontal line and a vertex for each clause $c_j$ either above or below the horizontal line and connected on that side of the line to the three variables appearing in $c_j$.
+	We can assume that $\phi$ comes with a planar drawing of its variable-clause graph $H_\phi$ that has a vertex for each variable $x_i$ and a vertex for each clause $c_j$.
+	Each clause vertex is connected to the three variables appearing in the clause.
+	In the drawing of $H_\phi$ all variable vertices are placed on a horizontal line and the clause vertices connect to the adjacent variable vertices either from above or from below the horizontal line.
 	In our gadget-based reduction we mimic the drawing of $H_\phi$ by creating a 3-plane drawing $\Gamma_\phi$ as a set of line segments of uniform length and a value $L$ such that $\Gamma_\phi$ has a SPED with ink at least $L$ if and only if $\phi$ is satisfiable.
@@
 
-	Each variable gadget is a cycle of segment pairs, with at least one pair for each occurence of the variable in $\phi$, see Fig.~\ref{fig:gadget_connection}. 
+	Each variable gadget is a cycle of segment pairs, with at least one pair for each occurrence of the variable in $\phi$, see Fig.~\ref{fig:gadget_connection}. 
 	Observe that this cycle has exactly two ink-maximal SPEDs: either all red edges are full segments and all green edges are length-1 stubs or vice versa.
@@
 More formally let $ j $ be the index such that stubs of length $ \ell_j(p(u)) $ do not intersect $ s(u) $, but $ \ell_{j +1}(p(u)) $ would, we set:
-%\begin{align*}
-\[	
-	\solshort(u) = \max\{T_1(u),\dots,T_{i_p}(u)\} \qquad
+\begin{align*}
+%\[	
+	\solshort(u) = \max\{T_1(u),\dots,T_{i_p}(u)\}\\ % \qquad
 	\sollong(u) = \max\{T_0(u),\dots,T_{\delta}(u)\}.
-\]
-%\end{align*}
+%\]
+\end{align*}
 Figure~\ref{fig:tree} highlights the stub lengths that are considered in $\solshort(u)$. We recursively define % for a vertex $ u $ and stub length $ \ell_i(u) $:
@@
 	Till Bruckdorfer, Michael Kaufmann, and Simon Leibßle.
-	\newblock Ped user study.
+	\newblock PED user study.
 	\newblock In {\em Graph Drawing (GD'15)}, volume 9411 of {\em LNCS}, pages
-	551--553. Springer International Publishing, 2015.
+	551--553. Springer, 2015.
 

### 2018-01-09 eurocg/paper.tex
kleinigkeit

@@
 	Let $\phi$ be a planar 3-Sat formula with $n$ variables $\{x_1, \dots, x_n\}$ and $m$ clauses $\{c_1, \dots, c_m\}$, each consisting of three literals.
-	We can assume that $\phi$ comes with a planar drawing of its variable-clause graph $H_\phi$ that has a vertex for each variable $x_i$ and a vertex for each clause $c_j$.
+	We can assume that $\phi$ comes with a planar drawing of its variable-clause graph $H_\phi$, which has a vertex for each variable $x_i$ and a vertex for each clause $c_j$.
 	Each clause vertex is connected to the three variables appearing in the clause.

### 2018-01-11 eurocg/paper.tex

@@
 These edge stubs are obtained by erasing one contiguous central piece of the straight-line segment connecting the two endpoints of each edge.
-In other words each straight-line edge is divided into three parts, of which only the two outer ones are drawn, see Fig.~\ref{fig:examples}.
+In other words each straight-line edge is divided into three parts, of which only the two outer ones are drawn (see Fig.~\ref{fig:examples}).
 More restricted and better readable~\cite{blmt-pedhmitc-16} variations of PEDs are \emph{symmetric} PEDs, in which both stubs of an edge must have the same length (see Fig.~\ref{fig:examples}(b)), and \emph{homogeneous} PEDs, in which the ratio of the stub length to the total edge length is the same for all edges.
@@
 The \emph{ink} or \emph{ink value}  $I(D)$ of a SPED $D$ is the total stub length $I(D) = \sum_{s \in S} f_s |s|$.
-In the problem \maxsped the task is to find for a given drawing~$\Gamma$ a SPED $D^*$ such that its ink $I(D^*)$ is maximum over all SPEDs.
+In the problem \maxsped, the task is to find for a given drawing~$\Gamma$ a SPED $D^*$ such that its ink $I(D^*)$ is maximum over all SPEDs.
 
@@
 	The intersection point of each red-green segment pair is at distance 1 from an endpoint.
-	Thus the maximum amount of ink contributed by such a pair is 7 (one full segment of length 5 and the other one with two stubs of length 1 each).
+	Thus, the maximum amount of ink contributed by such a pair is 7 (one full segment of length 5 and the other one with two stubs of length 1 each).
 
@@
 	For each clause we construct a triple of mutually intersecting segments, see the gadget on yellow background in the upper part of Fig.~\ref{fig:gadget_connection}.
-	Again, their intersection points are at distance 1 from the endpoints.
-	It is clear that in such a clause triangle at most one of the three segments can be fully drawn, while the stubs of the other two can have length at most 1.
-	Hence the maximum amount of ink in a SPED contributed by a clause gadget is 9.
+	Again, their intersection points are at distance~1 from the endpoints.
+	It is clear that in such a clause triangle at most one of the three segments can be fully drawn, while the stubs of the other two can have length at most~1.
+	Hence, the maximum amount of ink in a SPED contributed by a clause gadget is~9.
 
@@
 	Since the last segment of the selected literal path must be drawn as length-1 stubs, the only way of having a maximum contribution of that path is by alternating stubs and full segments.
-	Hence the first segment of the path must be a full segment.
+	Hence, the first segment of the path must be a full segment.
 	But because the variable is in the state that sets the literal to \emph{true}, the intersecting variable segment is drawn as two stubs and the path configuration is valid.
@@
 	For each clause, one segment must have full length and as argued before the path connecting that segment to a variable gadget must have all segments at even positions as length-1 stubs and all segments at odd positions as full segments.
-	In particular the first segment of the literal path is fully drawn and therefore in a SPED with ink $L$ the intersecting segment of the variable cannot be fully drawn.
+	In particular, the first segment of the literal path is fully drawn and therefore, in a SPED with ink $L$, the intersecting segment of the variable cannot be fully drawn.
 	In fact, it must be drawn as length-1 stubs for otherwise the variable gadget would not contribute its maximum ink value.
-	But then by construction the selected variable value sets our literal to \emph{true}.	
+	But then by construction, the selected variable value sets our literal to \emph{true}.	
 \end{proof}
@@
 \end{align*}
-Further we define the recurrence $T_i(u)$ for all articulation nodes $u \in \mathcal V$:
+Further, we define the recurrence $T_i(u)$ for all articulation nodes $u \in \mathcal V$:
 \begin{align}
@@
 For a block node $ u \in \mathcal V $ we cut open the cycle it represents at vertex $a(p(u))$ (refer to Bruckdorfer et al.~\cite{bcgkmn-pped-17} for details) and consider the three relevant stub lengths of the parent's segment $s(a(p(u)))$ induced by $i_p(u)$ and $i_q(u)$ and the whole segment.
-The resulting subgraph is a path on $V(u)$ with attached children that are articulation vertices in $\mathcal T$ so we can apply our algorithm of Section~\ref{sec:tree} with minor modifications.
+The resulting subgraph is a path on $V(u)$ with attached children that are articulation vertices in $\mathcal T$, so we can apply our algorithm of Section~\ref{sec:tree} with minor modifications.
 Once all values in the subgraph of block node $u$ are computed, we can derive the values $T_i(u)$ for the block node $u$.

### 2018-02-14 eurocg/paper.tex
minor remarks addressed

@@
 More restricted and better readable~\cite{blmt-pedhmitc-16} variations of PEDs are \emph{symmetric} PEDs, in which both stubs of an edge must have the same length (see Fig.~\ref{fig:examples}(b)), and \emph{homogeneous} PEDs, in which the ratio of the stub length to the total edge length is the same for all edges.
+\textcolor{blue}{Symmetric stubs facilitate finding adjacent vertices due to the identical stub lengths at both vertices, and symmetric homogeneous stubs additionally indicate the distance at which to find a neighboring vertex.}
 The natural optimization problem in this formal setting is \emph{ink maximization}, i.e., maximizing the total stub length, so that as much information as possible is given in the drawing while all crossings disappear in the negative background space.
@@
 
-Throughout the paper let $ G $ be a \emph{simple graph} with edge set $ S = \{s_1,\dots,s_m\}$ and $ \Gamma $ a straight line drawing of $ G $ in the plane.
+Throughout the paper let $ G $ be a \emph{simple graph} with edge set $ S = \{s_1,\dots,s_m\}$ and $ \Gamma $ a straight-line drawing of $ G $ in the plane.
 We call $ \Gamma $ \emph{$ k $-plane} if every edge $ s_i \in S $ is crossed by at most $ k $ other edges from $ S $ in $ \Gamma $.
@@
 %		\begin{cutout}{2}{0pt}{0.44\textwidth}{15}
-	We reduce from the \NP-hard problem \ppsat~\cite{l-pftu-82} using similar ideas as in Bruckdorfer's general proof of the hardness of \maxsped~\cite{b-sgh-15}.
-%, which reduces from the more restricted problem \textsc{Planar Positive 1-in-3 Sat}.
+	We reduce from the \NP-hard problem \ppsat~\cite{l-pftu-82} using similar ideas as in Bruckdorfer's sketch of the hardness proof for general \maxsped~\cite{b-sgh-15}. 
+%	Bruckdorfer reduces from the more restricted problem \textsc{Planar Positive 1-in-3 Sat}.
 	Let $\phi$ be a planar 3-Sat formula with $n$ variables $\{x_1, \dots, x_n\}$ and $m$ clauses $\{c_1, \dots, c_m\}$, each consisting of three literals.
@@
 
-	Each variable gadget is a cycle of segment pairs, with at least one pair for each occurrence of the variable in $\phi$, see Fig.~\ref{fig:gadget_connection}. 
+	Each variable gadget is a cycle of segment pairs, with (at least) one pair for each occurrence of the variable in $\phi$, see Fig.~\ref{fig:gadget_connection}. 
 	Observe that this cycle has exactly two ink-maximal SPEDs: either all red edges are full segments and all green edges are length-1 stubs or vice versa.
@@
 In both cases we present polynomial-time dynamic programming algorithms.
-Let $ C = (V,E) $ be the intersection graph of a given drawing $\Gamma$ of graph $ G $ as defined in Section~\ref{sec:preliminaries}.
+Let $ C = (V,E) $ be the intersection graph of a given drawing $\Gamma$ of a graph $ G $ as defined in Section~\ref{sec:preliminaries}.
 Let $ u \in V $ and $ \delta  = \deg(u) $. 
@@
 Let $v_j $ be the neighbor of $u$ that induces stub length $\ell_j(u)$ and assume $v_j \ne p(u)$.
-In $T_j(u)$ we could still count the value $\sollong(v_j)$, but in $T_{j+1}(u)$ the stub length of $u$ implies that $v_j$ can contribute only $\solshort(v_j)$. 
+In $T_j(u)$ we could still count the value $\sollong(v_j)$, but in $T_{j+1}(u)$ the stub length of $u$ implies that $v_j$ can contribute only to $\solshort(v_j)$. 
 Then $T_{j+1}(u) = T_j(u) - \sollong(v_j) + \solshort(v_j)$.
-If $v_j=p(u)$ the two values $T_j(u)$ and $T_{j+1}(u)$ are equal as the corresponding change in stub length has no effect on the children of $u$.
+If $v_j=p(u)$, then the two values $T_j(u)$ and $T_{j+1}(u)$ are equal as the corresponding change in stub length has no effect on the children of $u$.
 Computing $\solshort(u)$ and $\sollong(u)$ clearly takes $O(k)$ time.
@@
 	Let $ G $ be a simple graph with $m$ edges and $ \Gamma $ a straight-line drawing of $ G $.
-If the intersection graph $ C = (V,E) $ of $ G $ is a tree with maximum degree $ k \in \mathbb{N} $, problem \maxsped can be solved in $ O(mk + m \log m) $ time and space.
+If the intersection graph $ C = (V,E) $ of $ G $ is a tree with maximum degree $ k \in \mathbb{N} $, then problem \maxsped can be solved in $ O(mk + m \log m) $ time and space.
 \end{theorem}
@@
 T_{i_q(v)}(v) & \text{if } i_p(v) < i \leq i_q(v) \\
-T_{0}(v) & \text{otherwise}
+T_{0}(v) & \text{otherwise.}
 % \solshort(v) & \text{if } i < i_p(v)\\ %\ell_i(a(u)) \leq a \\
@@
 % \sollong(v) & \text{otherwise}
-\end{cases}.
+\end{cases}
 \end{align}

### 2018-02-14 eurocg/paper.tex
Added small section at the end of Hardness proof, explaining the extension to Bruckdorfers proof

--- 2018-01-11 eurocg/paper.tex
@@
 %		\begin{cutout}{2}{0pt}{0.44\textwidth}{15}
-	We reduce from the \NP-hard problem \ppsat~\cite{l-pftu-82} using similar ideas as in Bruckdorfer's general proof of the hardness of \maxsped~\cite{b-sgh-15}.
+	We reduce from the \NP-hard problem \ppsat~\cite{l-pftu-82} using similar ideas as in Bruckdorfer's general proof\textcolor{blue}{-sketch} of the hardness of \maxsped~\cite{b-sgh-15}.
 %, which reduces from the more restricted problem \textsc{Planar Positive 1-in-3 Sat}.
@@
 \end{proof}
+
+\textcolor{blue}{We extended Bruckdorfers proof-sketch by giving the ink contributions of each part of $\Gamma_\phi$, in order to give an ink value $L$ that can be achieved if and only if $\phi$ is satisfiable.
+Further we changed the variable gadgets to cycles of edge pairs, partly to guarantee, that the gadgets have the same ink value regardless of their corresponding truth values and partly for aesthetic reasons.
+And finally we changed the clause gadgets from three edges intersecting in the same point to the described triangles in order to achieve general position.}
 

### 2018-02-15 eurocg/paper.tex
formatting
finalize

--- 2018-02-14 eurocg/paper.tex
@@
 }
-\linenumbers
+%\linenumbers
 
@@
 \authorrunning{M. Höller, F. Klute, S. Nickel, M. N\"ollenburg, and B. Schreiber}
+
+\ArticleNo{50}
 
@@
 More restricted and better readable~\cite{blmt-pedhmitc-16} variations of PEDs are \emph{symmetric} PEDs, in which both stubs of an edge must have the same length (see Fig.~\ref{fig:examples}(b)), and \emph{homogeneous} PEDs, in which the ratio of the stub length to the total edge length is the same for all edges.
-\textcolor{blue}{Symmetric stubs facilitate finding adjacent vertices due to the identical stub lengths at both vertices, and symmetric homogeneous stubs additionally indicate the distance at which to find a neighboring vertex.}
+Symmetric stubs facilitate finding adjacent vertices due to the identical stub lengths at both vertices, and symmetric homogeneous stubs additionally indicate the distance at which to find a neighboring vertex.
 The natural optimization problem in this formal setting is \emph{ink maximization}, i.e., maximizing the total stub length, so that as much information as possible is given in the drawing while all crossings disappear in the negative background space.
@@
 %	Bruckdorfer reduces from the more restricted problem \textsc{Planar Positive 1-in-3 Sat}.
+Here we specify precisely the maximum ink contributions of all gadgets needed for a satisfying variable assignment. 
+Our variable gadgets are cycles of edge pairs that admit exactly two maximum-ink states.
+%Further we changed the variable gadgets to cycles of edge pairs, partly to guarantee, that the gadgets have the same ink value regardless of their corresponding truth values and partly for aesthetic reasons.
+Finally we construct clause gadgets consisting of three pairwise intersecting edges so that all crossings are between two edges only, while Bruckdorfer's gadgets have multiple edges intersecting in the same point.
 	Let $\phi$ be a planar 3-Sat formula with $n$ variables $\{x_1, \dots, x_n\}$ and $m$ clauses $\{c_1, \dots, c_m\}$, each consisting of three literals.
@@
 	In the drawing of $H_\phi$ all variable vertices are placed on a horizontal line and the clause vertices connect to the adjacent variable vertices either from above or from below the horizontal line.
-	In our gadget-based reduction we mimic the drawing of $H_\phi$ by creating a 3-plane drawing $\Gamma_\phi$ as a set of line segments of uniform length and a value $L$ such that $\Gamma_\phi$ has a SPED with ink at least $L$ if and only if $\phi$ is satisfiable.
-	Figure~\ref{fig:gadget_connection} shows a sketch of the gadgets.
+	In our reduction (see Fig.~\ref{fig:gadget_connection}) we mimic the drawing of $H_\phi$ by creating a 3-plane drawing $\Gamma_\phi$ as a set of line segments of uniform length and a value $L$ such that $\Gamma_\phi$ has a SPED with ink at least $L$ if and only if $\phi$ is satisfiable.
+%	Figure~\ref{fig:gadget_connection} shows a sketch of the gadgets.
 %		\end{cutout}
@@
 	In particular, each variable gadget is either in state \emph{true} or \emph{false}.
-	We claim that this truth assignment satisfies $\phi$.
-	For each clause, one segment must have full length and as argued before the path connecting that segment to a variable gadget must have all segments at even positions as length-1 stubs and all segments at odd positions as full segments.
-	In particular, the first segment of the literal path is fully drawn and therefore, in a SPED with ink $L$, the intersecting segment of the variable cannot be fully drawn.
-	In fact, it must be drawn as length-1 stubs for otherwise the variable gadget would not contribute its maximum ink value.
-	But then by construction, the selected variable value sets our literal to \emph{true}.	
+	By design of the gadgets it is straight-forward to verify that the corresponding truth assignment satisfies~$\phi$.
+%
+	% We claim that this truth assignment satisfies $\phi$.
+	% For each clause, one segment must have full length and as argued before the path connecting that segment to a variable gadget must have all segments at even positions as length-1 stubs and all segments at odd positions as full segments.
+	% In particular, the first segment of the literal path is fully drawn and therefore, in a SPED with ink $L$, the intersecting segment of the variable cannot be fully drawn.
+	% In fact, it must be drawn as length-1 stubs for otherwise the variable gadget would not contribute its maximum ink value.
+	% But then by construction, the selected variable value sets our literal to \emph{true}.
 \end{proof}
@@
 %\[	
-	\solshort(u) = \max\{T_1(u),\dots,T_{i_p}(u)\}\\ % \qquad
+	\solshort(u) = \max\{T_1(u),\dots,T_{i_p}(u)\}\\
 	\sollong(u) = \max\{T_0(u),\dots,T_{\delta}(u)\}.
@@
 		\solshort(v) & \text{if } s(u) \text{ with length } \ell_i(u) \text{ intersects } s(v) \\
-		\sollong(v) & \text{otherwise}
-	\end{cases}.
+		\sollong(v) & \text{otherwise.}
+	\end{cases}
 \end{align}
@@
 \begin{align*}
-	\sollong(u) = \max\{T_0(u),\dots,T_\delta(u)\}\\
+	\sollong(u) = \max\{T_0(u),\dots,T_\delta(u)\} \\
 	\solmid(u) = \max\{T_1(u),\dots,T_{i_q}(u)\}\\

### 2018-10-22 wads/paper.tex
added folder for wads version, copied eurocg paper into llncs format

--- 2018-02-14 eurocg/paper.tex
@@
-% for section-numbered lemmas etc., use "numberwithinsect"
-\documentclass[a4paper,english,numberwithinsect]{eurocg18}
+\documentclass[a4paper, USenglish]{llncs}
 
-% the recommended bibstyle
 \bibliographystyle{plainurl}
 
-%------------------------------------------------------------------- 
-%if unwanted, comment out or use option "draft"
 \usepackage{todonotes}
-\usepackage{microtype}
-\usepackage[noend, linesnumbered]{algorithm2e}
-\usepackage{amsmath}
-\usepackage{color}
-\usepackage{cite}
-\usepackage{pifont}
-\usepackage{enumerate}
+\usepackage{graphicx}
 \usepackage[basic]{complexity}
-\usepackage{subcaption}
 \usepackage{wrapfig}
-\usepackage{graphicx}
-%\usepackage{cutwin}
+\usepackage{subfig}
+\usepackage{xspace}
+\usepackage{hyperref}
+\usepackage{amsmath, amssymb}
 
-% Line numbers are helpful for refereeing
-\usepackage[mathlines]{lineno}
-\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
-\expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
-\expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
-\renewenvironment{#1}%
-     {\linenomath\csname old#1\endcsname}%
-     {\csname oldend#1\endcsname\endlinenomath}}%
-\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
-  \patchAmsMathEnvironmentForLineno{#1}%
-  \patchAmsMathEnvironmentForLineno{#1*}}%
-\AtBeginDocument{%
-\patchBothAmsMathEnvironmentsForLineno{equation}%
-\patchBothAmsMathEnvironmentsForLineno{align}%
-\patchBothAmsMathEnvironmentsForLineno{flalign}%
-\patchBothAmsMathEnvironmentsForLineno{alignat}%
-\patchBothAmsMathEnvironmentsForLineno{gather}%
-\patchBothAmsMathEnvironmentsForLineno{multline}%
-}
-\linenumbers
-
-%helpful if your graphic files are in another directory
-%\graphicspath{{./graphics/}}
-
-% Author macros::begin %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\newcommand\polylog{{\rm polylog}}
-
-% Author metadata::begin %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \title{Maximizing Ink in Symmetric Partial Edge Drawings of $k$-plane Graphs}
-%{Maximal Symmetric Partial Edge Drawings With Degree $ > 2$}
-%optional, in case that the title is too long; 
-%the running title should fit into the top page column
-%\titlerunning{MaxSPEDs with degree $ > 2$}
-
-%% Please provide for each author the \author and \affil macro, 
-%even when authors have the same affiliation, i.e. for each 
-%author there needs to be the  \author and \affil macros
-\author[1]{Michael H\"oller}
-\author[1]{Fabian Klute}
-\author[1]{Soeren Nickel}
-\author[1]{Martin~N\"ollenburg}
-\author[1]{Birgit Schreiber}
-
-\affil[1]{Algorithms and Complexity Group, TU Wien, Vienna, Austria\\ \texttt{[fklute|noellenburg]@ac.tuwien.ac.at, soeren.nickel@tuwien.ac.at}}
-%\affil[2]{Algorithms and Complexity Group, TU Wien, Vienna, Austria} %\texttt{fklute@ac.tuwien.ac.at}}
-%\affil[3]{Algorithms and Complexity Group, TU Wien, Vienna, Austria \texttt{}}
-%\affil[4]{Algorithms and Complexity Group, TU Wien, Vienna, Austria} %\texttt{noellenburg@ac.tuwien.ac.at}}
-%\affil[5]{Algorithms and Complexity Group, TU Wien, Vienna, Austria \texttt{}}
-%mandatory. First: Use abbreviated first/middle names. 
-%Second (only in severe cases): Use first author plus 'et. al.'
-\authorrunning{M. Höller, F. Klute, S. Nickel, M. N\"ollenburg, and B. Schreiber}
-
-% Author macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\author{Fabian Klute \and Soeren Nickel \and Martin~N\"ollenburg}
+\institute{Algorithms and Complexity Group, TU Wien, Vienna, Austria \email{[fklute|noellenburg]@ac.tuwien.ac.at, soeren.nickel@tuwien.ac.at}}
 
@@
 \begin{document}
-
 	\maketitle
@@
 
-\begin{wrapfigure}[11]{r}{0.45\textwidth}
+	\begin{figure}
 		\centering
-	\includegraphics{export_full}
-	\hfill
-	\includegraphics{export_alpha_nontouching}
+		\subfloat{
+			\includegraphics{figures/export_full}
+		}
+		\subfloat {
+			\includegraphics{figures/export_alpha_nontouching}
+		}	
 		\caption{A straight-line graph drawing (a) and a maximum-ink symmetric partial edge drawing (b) of the same graph.}
 		\label{fig:examples}
-\end{wrapfigure}
+	\end{figure}
 
@@
 	\section{Preliminaries}\label{sec:preliminaries}
-%\begin{itemize}
-%	\item basic definitions and notation
-%	\item segment intersection graph
-%	\item alle planaren Graphen können auftreten (insbes. Bäume und Kakteen)
-%\end{itemize}
-%Input: $k$-plane graph drawing $\Gamma$ of a graph $G$ with edge set $S={s_1, \dots, s_m}$ (set of segments)
-%Intersection graph $C=(V,E)$ with $V=S$ and edges if two segments intersect 
-%Consider maximum degree of $C$ -- here mostly 3.
 
@@
 
-%\fabian{This part has to be adapted once the intro is fixed}
-%
-%We define a \emph{partial edge drawing} (\ped) $ \Gamma' $ of $ G $ as a drawing using the same embedding as $ \Gamma $, but edges are drawn as \emph{stubs}, i.e., we do not depict an edge $ s_i \in S $ by a line segment connecting the two vertices, but remove the middle part of the edges such that no two stubs intersect in $ \Gamma' $.
-We call a \ped $ \Gamma' $ \emph{symmetric} if for every edge $ s_i s_j \in S $ its two stubs have equal lengths in $ \Gamma' $.
-Finally we define the problem \maxsped as finding a symmetric \ped $ \Gamma' $ of $ G $ where the sum of the lengths of the stubs is maximized.
-This is also known as ink maximization~\cite{bcgkmn-pped-17}.
-%
-%\martin{Maybe we don't need the above paragraph as PED, SPED, maxSPED are all somewhat defined in the introduction}
-
 	\section{Hardness of \maxsped for $k\ge3$}\label{sec:hardness}
@@
 	\end{theorem}
-
-%\renewcommand\windowpagestuff{%
-%	\includegraphics{Reduction_visual}
-%	\captionof{figure}{Example of three variable gadgets and\\ a satisfied clause gadget. Dotted parts do not belong\\ to the SPED.}
-%	\label{fig:gadget_connection}
-%}
-%\opencutleft
 	\begin{proof}
-%		\begin{cutout}{2}{0pt}{0.44\textwidth}{15}
-	We reduce from the \NP-hard problem \ppsat~\cite{l-pftu-82} using similar ideas as in Bruckdorfer's general proof\textcolor{blue}{-sketch} of the hardness of \maxsped~\cite{b-sgh-15}.
-%, which reduces from the more restricted problem \textsc{Planar Positive 1-in-3 Sat}.
+		We reduce from the \NP-hard problem \ppsat~\cite{l-pftu-82} using similar ideas as in Bruckdorfer's general proof of the hardness of \maxsped~\cite{b-sgh-15}. 
 		Let $\phi$ be a planar 3-Sat formula with $n$ variables $\{x_1, \dots, x_n\}$ and $m$ clauses $\{c_1, \dots, c_m\}$, each consisting of three literals.
@@
 		Figure~\ref{fig:gadget_connection} shows a sketch of the gadgets.
-%		\end{cutout}
-
 		\begin{figure}
-%	\begin{wrapfigure}[15]{r}{0.55\textwidth}
-
 			\centering
-\includegraphics{Reduction_visual}
+			\includegraphics{figures/Reduction_visual}
 			\caption{Example of three variable gadgets and a satisfied clause gadget. Dotted parts do not belong to the SPED.}
 			\label{fig:gadget_connection}
-%	\end{wrapfigure}		
 		\end{figure}
@@
 
-\textcolor{blue}{We extended Bruckdorfers proof-sketch by giving the ink contributions of each part of $\Gamma_\phi$, in order to give an ink value $L$ that can be achieved if and only if $\phi$ is satisfiable.
-Further we changed the variable gadgets to cycles of edge pairs, partly to guarantee, that the gadgets have the same ink value regardless of their corresponding truth values and partly for aesthetic reasons.
-And finally we changed the clause gadgets from three edges intersecting in the same point to the described triangles in order to achieve general position.}
-
-%\newpage %delete later
 	\section{Two polynomial special cases}
-
-%\begin{itemize}
-%	\item dynamic programming
-%	\item table entries $T_i(s)$ for $i = 1, \dots, \deg(s)$
-%	\item $i$ corresponds to increasing stub lengths
-%	\item stub lengths $l_i(s)$ for $i = 1, \dots, \deg(s)$ (not the pair of stubs)
-%\end{itemize}
 
@@
 		\centering
- 	\includegraphics{tree_illustration}
+		\includegraphics{figures/tree_illustration}
 		\caption{A segment $s(u)$ with five intersecting segments and the different induced stub lengths.
@@
 
-%Computing the intersection graph requires to identify all the intersections in $ \Gamma $.
-Using a standard sweep line algorithm \cite{}\fabian{add citation} this requires $ O(|S|\log |S| + I) $ where $ I $ is the number of crossings.
-Building the graph then takes $ O(|S|) $ time.
-When computing the intersection we can compute the stub lengths and their order as well.
-%
-%\martin{time to build graph should be $O(S+I)$; but maybe this preparation work should go to preliminaries so that we can assume the intersection graph as input}
-%\fabian{add to prelims}
-
 	\subsection{Trees}
 	\label{sec:tree}
-
-% \begin{figure}[tbp]
-% 	\centering
-%  	\includegraphics{tree_illustration}
-% 	\caption{One edge $ s(u) $ with intersecting segments. The shaded areas are the stubs, $ p(u) $ is the parent of $ u $. The stubs in the boxed area contribute to $ \solshort(u) $.}
-% 	\label{fig:tree}
-% \end{figure}
 
@@
 	let $ i_p $ be the index of the stub length $\ell_{i_p}(u)$ induced by the intersection point of $s(u)$ and $s(p(u))$.
-%the shorter length between one of $ s(u) $'s endpoints and the intersection point between $ s(u) $ and $ s(p(u)) $. 
+
 	We define the following two values, which allow us to categorize the stub lengths into those not affecting the stubs of the parent and those that do affect the parent:
-%Since we know $ C $ is a tree every vertex $ u \in V $ with degree $ \delta $ has a parent $ p(u) \in V $.
-For this parent there exists one stub length $ \ell_j(p(u)) $ in the sequence of stub lengths $ \ell_0(p(u)),\dots,\ell_{\delta}(p(u)) $ such that drawing $ p(u) $ with these stubs does not intersect a completely drawn $ u $, while drawing $ p(u) $ with $ \ell_{j + 1}(p(u)) $, i.e.
-the next longer stub length, intersects $ \ell_0(u) $.
-We call the solution which would intersect the parent $ \sollong(u) $ and the one not intersecting the parent $ \solshort(u) $.
-See Figure~\ref{fig:tree} for an illustration.
-More formally let $ j $ be the index such that stubs of length $ \ell_j(p(u)) $ do not intersect $ s(u) $, but $ \ell_{j +1}(p(u)) $ would, we set:
 	\begin{align*}
-%\[	
 		\solshort(u) = \max\{T_1(u),\dots,T_{i_p}(u)\}\\ % \qquad
 		\sollong(u) = \max\{T_0(u),\dots,T_{\delta}(u)\}.
-%\]
 	\end{align*}
-Figure~\ref{fig:tree} highlights the stub lengths that are considered in $\solshort(u)$. We recursively define % for a vertex $ u $ and stub length $ \ell_i(u) $:
+	Figure~\ref{fig:tree} highlights the stub lengths that are considered in $\solshort(u)$. We recursively define 
 	\begin{align}
@@
 	Setting $ \sollong(u) $ and $ \solshort(u) $ as above yields the two maximum ink values that are relevant for $ p(u) $.
-%\begin{lemma}
-%	\label{lem:tree_correctness}	
-%	Let $ G $ be a simple graph and $ \Gamma $ a straight-line drawing of $ G $.
-If the intersection graph $ C = (V,E) $ of $ G $ is a tree with root $ r \in V $ and maximum degree $ k $, then $ \sollong(r) = \max \{T_i(r) \mid 0 \leq i \leq k \} $ is an optimal solution to the \maxsped problem on $ G $ with drawing $ \Gamma $.
-%\end{lemma}
-%\begin{proof}
-%	Let $ G $, $ \Gamma $ and $ C = (V,E) $ as above and $ u \in V $ a leaf of $ C $.
-Since $ c(u) $ is empty we simply get $ \solshort(u) = T_0(u) = \ell_1(u) $ and $ \sollong(u) = T_1(u) = \max\{\ell_0(u), \ell_1(u)\} $.
-Now let $ u \in V $ with degree $ \delta $ be an inner vertex.
-The values $ \solshort(v) $ and $ \sollong(v) $ are computed for all $ v \in c(u) $.
-$ T_i(u) $ for one $ 0 \leq i \leq \delta $ is then the sum of the maximum stubs we can draw among the children plus the stub length $ \ell_i(u) $ itself, which are just the values we are interested in.
-Setting then $ \sollong(u) $ and $ \solshort(u) $ as above gives the two necessary values for $ p(u) $.
-%	and $ t = \sollong(r) $ is the maximum value for some $ T_i(r) $'s and $ r $ the root.
-Now let $ t' $ be the value of a solution to the \maxsped problem on $ G $ with drawing $ \Gamma $ and $ t' > t $.
-This solution has to be the sum of a set of stub lengths for each edge in $ G $ such that no two stubs intersect by definition.
-Further we know the stubs can be categorized into two sets relative to their parent.
-Recurrence~\ref{rec:tree} builds just the maximum over the $ \solshort(v) $ and $ \sollong(v) $ values for all $ v \in c(u) $, such that no two stubs intersect.
-Hence if $ t $ was not optimal we would need to find at least a third stub length for at least one child $ v \in c(u) $ which was neither equal to $ \sollong(v) $ nor $ \solshort(v) $.
-%\end{proof}
 
@@
 
-% \begin{lemma}
-% 	\label{lem:compute_vertex}
-% 	Let $ C = (V,E) $ be a tree with maximum degree $ k $ and vertex weights $ \ell_0,\dots,\ell_{\delta} $ for every $ u \in V $ with degree $ \delta $.
-We can compute $ \solshort(u) $ and $ \sollong(u) $ in $ O(\delta) $ time and space, if the values $ \sollong(v) $ and $ \solshort(v) $ are computed for all children $ v \in c(u) $.
-% \end{lemma}
-% \begin{proof}
-% 	 Let $ u \in V $ be a vertex with degree $ \delta < k $ for which we have computed $ \sollong $ and $ \solshort $ for every of its children.
-% %	 The crucial part is that $ T_i(u) $ can be built iteratively by going through the list of stub lengths for $ u $ in ascending order.
-% 	First we compute $ T_0(u) $ and $ T_1(u) $, this can be done since they represent the longest (whole edge) and shortest stub lengths.
-% 	 Now assume we computed $ T_j(u) $ for all stub lengths $ \ell_i(u) $ with $ i \leq j $, hence we want to compute $ T_{j+1}(u) $ next.
-Doing so only requires to consider the next stub-length $ \ell_{j + 1}(u) $, since all other affected by drawing $ s(u) $ with $ \ell_{j+1}(u) $ where already considered.
-Hence we can compute $ T_{j+1}(u) $ in $O(1)$ time by simply subtracting $ \sollong(v) $ and adding $ \solshort(v) $.
-%
-% 	Finally we need to compute $ \solshort(u) $ and $ \sollong(u) $, but this are just two maximizations over at most $ \delta + 1 $ values at most.
-This can be done in $ O(\delta) $ by simply sweeping over the values.
-% \end{proof}
-
 	So by solving Recurrence~(\ref{rec:tree}) in $O(mk)$ time we find an optimal solution to the \maxsped problem on $ G $ with drawing $ \Gamma $ using standard backtracking from $\max\{T_0(r), \dots, T_{\deg(r)}(r)\}$ for the root $r$ of $C$.
@@
 	\end{theorem}
-% \begin{proof}
-% 	Since $ C $ is known to be a tree, and hence $ |E| = O(m) $, we can compute the intersection graph and the stub lengths in $ O(m\log m) $ time and space.
-% 	With Lemma~\ref{lem:compute_vertex} and Recurrence~\ref{rec:tree} we need one pass through the tree to find the roots maximum in $ O(mk) $ time.
-Adding the bounds yields the desired runtime of $ O(mk + m \log m)) $.
-% %	To solve the \maxsped problem on $ G $ we need to compute $ \max \{T_i(r) \mid 0 \leq i \leq k \} $ for the root $ r \in V $ of $ C $ as seen above.
-Doing this bottom up requires one pass through $ C $ hence by applying Lemma~\ref{lem:compute_vertex} we need $ O(mk) $ time to compute all $ T_i(u) $ for all $ u \in V $.
-For every vertex we only need to compute $ O(k) $ values and store two of them, namely $ \sollong(u) $ and $ \solshort(u) $, so we do not need more than $ O(mk) $ memory.
-Adding the bounds yields the desired runtime of $ O(mk + m \log m)) $.
-% %
-%
-%
-%
-% 	Retrieving the actual stub length for each edge in $ G $ needed to produce the SPED can be done by backtracking the subsolutions which requires at most $ O(mk) $ time.
-% \end{proof}
-
-%\begin{itemize}
-%	\item segment $s$ with $d$ children $u_1, \dots, u_{d}$ and parent $p(s) = p$.
-%	\item Assume children are sorted by their intersection points on $s$, i.e., $u_1$ produces stub length $l_1$ etc.
-%	\item for stub length relating to parent use offset from next longer stub length
-%\end{itemize}
 
@@
 	\label{sec:cactus}
-
-%In this section we restrict $ C = (V,E) $ to be a cactus graph, i.e.
-a simple graph in which no edge $ e \in E $ is part of more than one cycle.
-For a cactus graph $ C $ the \emph{block-cut tree} is the tree $ \mathcal T = (\mathcal V, \mathcal E) $ where $ \mathcal V $ contains a vertex for every cycle and cut vertex in $ C $.
-We call a vertex $ u \in \mathcal V $ a \emph{block vertex} if it represents an induced cycle with vertices $ V_c \subseteq V $ and \emph{articulation vertex} representing the cut vertex $ v \in V $ else.
-For a block vertex $ u \in \mathcal V $ we define $ \textit{cycle}(u) = V_c$ and for an articulation vertex $ u \in \mathcal V $ we define $ A(u) = v $.
-There is an edge $ uv \in \mathcal E $ between a block vertex $ u $ and articulation vertex $ v $ whenever $ v \in \textit{cycle}(u) $ and an edge $ uv \in \mathcal E $ between two articulation vertices whenever $ A(u)A(v) $ was an edge in $ E $.
-%
-%Let $ C = (V,E) $ be a cactus graph and moreover the intersection graph of a graph $ G $ with straight-line drawing $ \Gamma $.
-Additionally let $ \mathcal T = (\mathcal V, \mathcal E) $ be the block-cut tree of $ C $ we define for each vertex $ u \in \mathcal V $ two indices $ i_p,i_q $, w.l.o.g $ i_p > i_q $.
-This are exactly the indices of the stub-lengths leading to an intersection between two cycles of segments in $ \Gamma $.
-Similar to Section~\ref{sec:tree} we define three values for every vertex $ u \in \mathcal V $ with degree $ \delta $:
-%\begin{align*}
-%	\solshort(u) &= \max\{T_0(u),\dots,T_\delta(u)\}\\
-%	\solmid(u) &= \max\{T_1(u),\dots,T_{i_p}(u)\}\\
-%	\sollong(u) &= \max\{T_1(u),\dots,T_{i_q}(u)\}
-%\end{align*}
-%%\begin{eqnarray*}
-%%	\solshort(u) = \max\{T_0(u),\dots,T_\delta(u)\}\qquad
-%%	\solmid(u) = \max\{T_1(u),\dots,T_{i_p}(u)\}\\
-%%	\multicolumn{2}{c}{$\sollong(u) = \max\{T_1(u),\dots,T_{i_q}(u)\}$}
-%%\end{eqnarray*}
-%For an articulation vertex $ u \in \mathcal V $ we can define a similar recurrence to Recurrence~\ref{rec:tree}:
 
@@
 	We call a node $ u \in \mathcal V $ a \emph{block node} if it represents an induced cycle on vertices $ V(u) \subseteq V $ and an \emph{articulation node} of the cut vertex $ a(u) = v \in V $ otherwise.
-%For a block node $ u \in \mathcal V $ we define $ \textit{cycle}(u) = V_c$ and f
-%For an articulation node $ u \in \mathcal V $ we define $ a(u) = v $ to be the corresponding cut vertex. 
 	We have an edge $ uv \in \mathcal E $ between a block node $ u $ and an articulation node $ v $ whenever $ a(v) \in V(u) $ and an edge $ uv \in \mathcal E $ between two articulation nodes whenever $ a(u)a(v) \in E$.
@@
 
-%Let $ C = (V,E) $ be a cactus graph and moreover the intersection graph of a graph $ G $ and straight-line drawing $ \Gamma $.
-Additionally let $ \mathcal T = (\mathcal V, \mathcal E) $ be the block-cut tree of $ C $
-We further define for each node $ u \in \mathcal V $ two indices $ i_p(u) \le i_q(u) $. %, w.l.o.g $ i_p > i_q $. 
+	We further define for each node $ u \in \mathcal V $ two indices $ i_p(u) \le i_q(u) $.  
 	For an articulation node $u$, $i_p$ and $i_q$ are the indices corresponding to the stub lengths induced by the intersection points of $s(a(u))$ and its neighbors in $C$ among the vertex set of $p(u)$.
 	For a block node $u$, $i_p$ and $i_q$ are the indices corresponding to the intersection points of the parent's segment $s(a(p(u)))$ with its neighbors in $V(u)$.
-%These are exactly the indices of the stub-lengths leading to an intersection between two cycles of segments in $ \Gamma $. 
 	Similar to Section~\ref{sec:tree} we now define three values for each articulation node $ u \in \mathcal V $ with degree $ \delta $:
@@
 	\begin{cases}
-%\solshort(v) & \text{if } \ell_i(A(u)) \leq a \\
-%\solmid(v) & \text{if } a < \ell_i(A(u)) \leq b \\
-%\sollong(v) & \text{else}
-%\end{cases}.
-%\end{align}
-%For a block vertex $ u \in \mathcal V $ we have to cut open the cycle it represents (compare \cite{bcgkmn-pped-17}).
-We do this at $ A(p(u)) $.
-The resulting graph is a path between the vertices in $ \textit{cycle}(u) $ with attached articulation vertices.
-This is especially a tree and we use the aglorithm from Section~\ref{sec:tree}, maximizing over three values of the articulation vertices though,  to compute the $ T_i(u) $.
-%
-%If the values computed for the block vertices is correct, we can follow the correctness of the algorithm as in Section~\ref{sec:tree}.
-Let $ u \in \mathcal V $ be a block vertex then by \cite{bcgkmn-pped-17} we know the optimal solution for \maxsped just on $ C[\textit{cycle}(u)] $ is correctly computed by the above procedure since the algorithm from Section~\ref{sec:tree} correctly handles the created path.
-Now if $ u $ has articulation vertices attached to it the above computes the correct value since this is simply a tree.
-Further there is no $ v \in c(u) $ which is not an articulation vertex, hence the maximum ink value for $ u $ is correctly computed if we know the values for the children in $ c(u) $.
-%
-%For the runtime we can again argue the same as in Section~\ref{sec:tree}, but this time we have to also check cycles of length $ n_c $ which leads to a runtime of $ O(n_ck^2) $ for a block vertex $ u \in \mathcal V $ with $ n_c = |\textit{cycle}(u)| $.
-In case $ u $ is an articulation vertex we can solve it in $ O(k) $.
-As before the optimal solution to the \maxsped problem on $ G $ with drawing $ \Gamma $ can be found by backtracking from the root.
-Additionally the intersection graph can be computed in $ O(m\log m) $ where $ m $ is the number of edges in the original graph $ G $ since $ |E| = O(m) $ in a cactus and we obtain the following theorem.
 	T_{i_p(v)}(v) & \text{if } 1 \le i < i_p(v)\\ %\ell_i(a(u)) \leq a \\
@@
 	T_{0}(v) & \text{otherwise}
-% \solshort(v) & \text{if } i < i_p(v)\\ %\ell_i(a(u)) \leq a \\
-% \solmid(v) & \text{if } i_p(v) < i \leq i_q(v) \\
-% \sollong(v) & \text{otherwise}
 	\end{cases}.
@@
 
-% The resulting graph is a path over the vertices in $ \textit{cycle}(u) $ with attached articulation vertices.
-This is especially a tree and we use the aglorithm from Section~\ref{sec:tree}, maximizing over three values of the articulation vertices though,  to compute the $ T_i(u) $.
-
 	Correctness can be proved by modifying the inductive arguments of Section~\ref{sec:tree} accordingly. 
 	Combining the linear running time for cycles of Bruckdorfer et al.~\cite{bcgkmn-pped-17} with the running time for trees of Theorem~\ref{thm:tree} we can argue that cactus graphs can in fact be handled in the same time and space bounds as trees.
-
-% The correctness follows as in the case of a tree in Section~\ref{sec:tree} if the values computed for the block vertices is correct.
-This is the case since we only extended the cycle by attaching some leafs, which can be dealt with after cutting the cycle open as in \cite{bcgkmn-pped-17}.
-For the runtime we can again do the same as in Section~\ref{sec:tree}, but this time we have to also check cycles of length $ n_c $ which leads to a runtime of $ O(n_c\delta) $ for a block vertex $ u \in \mathcal V $ with degree $ \delta $ and $ n_c = |\textit{cycle}(u)| $.
-In case $ u $ is an articulation vertex we can solve it in $ O(\delta) $.
-As before the optimal solution to the \maxsped problem on $ G $ with drawing $ \Gamma $ can be found by backtracking from the root.
-Additionally the intersection graph can be computed in $ O(m\log m) $ where $ m $ is the number of edges in the original graph $ G $ and we obtain the following theorem.
-
-%$ G $, $ \Gamma $,  $ C = (V,E) $ and $ \mathcal T = (\mathcal V,\mathcal E) $ as above and $ u \in \mathcal V $ a leaf of $ \mathcal T$.
-Then either $ u $ is an articulation vertex and to be treated as in Lemma~\ref{lem:compute_vertex} or a cycle.
-In the latter case we find the optimal solution to the \maxsped problem on $ C[\textit{cycle}(u)] $ with the algorithm from Bruckdorfer et al.
-\cite{bcgkmn-pped-17}.
-Using the above described method we derive $ \solshort(u) $, $ \solmid(u) $ and $ \sollong(u) $, this are exactly the optimal values depending on how the cycle interacts with the articulation vertex $ p(u) $.
-%
-%Now let $ u \in V $ with degree $ \delta $ be an inner vertex of $ \mathcal T $.
-The values $ \solshort(v) $, $ \solmid(v) $ and $ \sollong(v) $ are computed for all $ v \in c(u) $.
-Again first let $ u $ be an articulation vertex, then the maximum for stub $ \ell_i(A(u)) $ is said value plus the sum of the compatible children, which is just Recurrence~\ref{rec:cactus_articulation}.
-Now let $ u $ be a block vertex.
-%\begin{lemma}
-%	\label{lem:cactus_correctness}	
-%	Let $ G $ be a simple graph and $ \Gamma $ a straight-line drawing of $ G $.
-If the intersection graph $ C = (V,E) $ of $ G $ is a cactus with root $ r \in V $ and maximum degree $ k $, then $ \sollong(r) = \max \{T_i(r) \mid 0 \leq i \leq k \} $ is an optimal solution to the \maxsped problem on $ G $ with drawing $ \Gamma $.
-%\end{lemma}
-%\begin{proof}	
-%	Let $ G $, $ \Gamma $,  $ C = (V,E) $ and $ \mathcal T = (\mathcal V,\mathcal E) $ as above and $ u \in \mathcal V $ a leaf of $ \mathcal T$.
-Then either $ u $ is an articulation vertex and to be treated as in Lemma~\ref{lem:compute_vertex} or a cycle.
-In the latter case we find the optimal solution to the \maxsped problem on $ C[\textit{cycle}(u)] $ with the algorithm from Bruckdorfer et al.
-\cite{bcgkmn-pped-17}.
-Using the above described method we derive $ \solshort(u) $, $ \solmid(u) $ and $ \sollong(u) $, this are exactly the optimal values depending on how the cycle interacts with the articulation vertex $ p(u) $.
-%
-%	Now let $ u \in V $ with degree $ \delta $ be an inner vertex of $ \mathcal T $.
-The values $ \solshort(v) $, $ \solmid(v) $ and $ \sollong(v) $ are computed for all $ v \in c(u) $.
-Again first let $ u $ be an articulation vertex, then the maximum for stub $ \ell_i(A(u)) $ is said value plus the sum of the compatible children, which is just Recurrence~\ref{rec:cactus_articulation}.
-Now let $ u $ be a block vertex.
-We have to cut open the cycle it represents (compare \cite{bcgkmn-pped-17}).
-We do this at $ p = A(p(u)) $.
-The resulting graph is a path over the vertices in $ \textit{cycle}(u) $ with attached articulation vertices.
-This is especially a tree and we use the aglorithm from Section~\ref{sec:tree}, maximizing over three values of the articulation vertices though,  to compute the $ T_i(u) $.
-%\end{proof}
-%\begin{lemma}
-%	\label{lem:compute_block}
-%	Let $ \mathcal T = (\mathcal V,\mathcal E) $ be a cactus with maximum degree $ k $ and vertex weights $ \ell_0,\dots,\ell_{\delta} $ for every block vertex $ u \in \mathcal V $ with degree $ \delta $.
-We can compute $ \solshort(u) $, $ \solmid(u) $ and $ \sollong(u) $ in $ O(n_c\delta) $ time and space, if the values $ \sollong(v) $, $ \solmid(v) $ and $ \solshort(v) $ are computed for all children $ v \in c(u) $.
-%\end{lemma}
-%\begin{proof}
-%	Let $ G $, $ \Gamma $,  $ C = (V,E) $ with maximum degree $ k $ and $ \mathcal T = (\mathcal V,\mathcal E) $ as above.
-Let $ u \in \mathcal V $ be a block vertex.
-Computing the value for the cycle itself $ \textit{cycle}(u) $ takes $ O(n_c) $ time by Bruckdorfer et al.
-with $ n_c $ being the number of vertices in the cycle.
-Now at every vertex $ v \in \textit{cycle}(u) $ there might be a set of up to $ k $ many children, which are all articulation points.
-Their values are already known and maximization over them can hence be done in $ O(k) $ time by the same argument as in Lemma~\ref{lem:compute_vertex}.
-Following we need $ O(n_ck) $ to compute the three values $ \solshort(u) $, $ \solmid(u) $ and $ \sollong(u) $.
-%\end{proof}
 
@@
 	\end{theorem}
-%\begin{proof}
-%	Follows as in Theorem~\ref{thm:tree} by maximizing over the tree using Recurrence~\ref{rec:cactus_articulation}.
-Then the running time follows from constructing the intersection graph and the block-cut tree as well as Lemma~\ref{lem:compute_block} for the block vertices and applying Lemma~\ref{lem:compute_vertex} for the articulation vertices.
-%\end{proof}
 
-%\bibliography{paper}
-\begin{thebibliography}{1}
-
-	\bibitem{bew-vnd-95}
-	Richard~A. Becker, Stephen~G. Eick, and Allan~R. Wilks.
-	\newblock Visualizing network data.
-	\newblock {\em {IEEE} Trans. Visualization and Computer Graphics}, 1(1):16--28,
-	1995.
-
-	\bibitem{blmt-pedhmitc-16}
-	Carla Binucci, Giuseppe Liotta, Fabrizio Montecchiani, and Alessandra Tappini.
-	\newblock Partial edge drawing: Homogeneity is more important than crossings
-	and ink.
-	\newblock In {\em Information, Intelligence, Systems Applications (IISA'16)},
-	pages 1--6. IEEE, 2016.
-
-	\bibitem{b-sgh-15}
-	Till Bruckdorfer.
-	\newblock {\em Schematics of Graphs and Hypergraphs}.
-	\newblock PhD thesis, Uni Tübingen, 2015.
-
-	\bibitem{bcgkmn-pped-17}
-	Till Bruckdorfer, Sabine Cornelsen, Carsten Gutwenger, Michael Kaufmann,
-	Fabrizio Montecchiani, Martin Nöllenburg, and Alexander Wolff.
-	\newblock Progress on partial edge drawings.
-	\newblock {\em J. Graph Algorithms Appl.}, 21(4):757--786, 2017.
-
-	\bibitem{bk-ecbe-12}
-	Till Bruckdorfer and Michael Kaufmann.
-	\newblock Mad at edge crossings? {Break} the edges!
-	\newblock In {\em Fun with Algorithms (FUN'12)}, volume 7288 of {\em LNCS},
-	pages 40--50. Springer, 2012.
-
-	\bibitem{bkl-us-15}
-	Till Bruckdorfer, Michael Kaufmann, and Simon Leibßle.
-	\newblock PED user study.
-	\newblock In {\em Graph Drawing (GD'15)}, volume 9411 of {\em LNCS}, pages
-	551--553. Springer, 2015.
-
-	\bibitem{bvkw-epdldge-12}
-	Michael Burch, Corinna Vehlow, Natalia Konevtsova, and Daniel Weiskopf.
-	\newblock Evaluating partially drawn links for directed graph edges.
-	\newblock In {\em Graph Drawing (GD'11)}, volume 7034 of {\em LNCS}, pages
-	226--237. Springer, 2012.
-
-	\bibitem{bcko-cgaa-08}
-	Mark de~Berg, Otfried Cheong, Marc van Kreveld, and Mark Overmars.
-	\newblock {\em Computational Geometry: Algorithms and Applications}.
-	\newblock Springer, 3rd edition edition, 2008.
-
-	\bibitem{l-pftu-82}
-	David Lichtenstein.
-	\newblock Planar formulae and their uses.
-	\newblock {\em SIAM J Comput}, 11(2):329--343, 1982.	
-\end{thebibliography}
+	\bibliography{paper}
 \end{document}
-
-%%%% OLD TEXT PARTS %%%%%
-
-\begin{proof}
-
-% To show that \maxsped is \NP-hard even for 3-plane input drawings, we will provide a reduction from the \NP-hard \ppsat to \maxsped, similar to Till Bruckdorfers Dissertation \cite{b-sgh-15}.
-An incremental explanation of \ppsat would be: Given a formula $\phi$ in 3-CNF, in which every single occurring literal is not negated (positive).
-Consider the graph $P(\phi)$ that contains a vertex for each clause (clause vertex) and a vertex for each variable (variable vertex) and every variable vertex is connected via an edge with a clause vertex, if and only if the variable occurs in the corresponding clause.
-If $P(\phi)$ is planar, $\phi$ and a planar embedding $\Pi$ of $P(\phi)$ is an instance of \ppsat.
-The problem is to decide whether $\phi$ has a satisfying truth assignment, such that only one (and since it is a satisfying assignment, exactly one) variable in each clause is set to true.
-
-\begin{figure}
-	\centering
-	\hfill
-	\includegraphics[width=.2\linewidth]{clause_to_triangle}
-	\hfill
-	\includegraphics[width=.5\linewidth]{variable_to_circle}
-	\caption{A triangle representing a clause, with unit lengths (a) and circles of segments representing variables, coloured in red and green (b).}
-	\label{fig:gadgets}
-\end{figure}
-
-% We create a graph G which is an instance of \maxsped out of an instance of \ppsat by creating the following components.
-
-%\begin{itemize}
-%	\item a triangle of overlapping lines for every clause vertex (Fig.~\ref{fig:gadgets} (a))
-%	\item a circle of overlapping lines of even length for every variable vertex (Fig.~\ref{fig:gadgets} (b))
-%	\item a path of overlapping lines of even length for every connection between a variable vertex and a clause vertex in $\Pi$ (Fig.~\ref{fig:gadget_connection})
-%\end{itemize}
-
-A triangle (gadget clause) of overlapping lines $l_{x_a}$, $l_{x_b}$, and $l_{x_c}$ is constructed for each clause $(x_a \lor x_b \lor x_c)$ such that each edge of the trinagle is split into two stubs of length 1 and a middle part of length 2 (see Fig~\ref{fig:gadgets}(a)).
-Each edge of this clause gadget will be connected to exactly one variable gadget since every clause in $\phi$ consists of exactly three literals.
-Observe that only one edge at a time can be drawn completely since it intersects both other edges, that the optimal solution in terms of maximized ink is drawing one edge completely (4 units) and the stubs of the other two edges (2 units respectively), so 8 units in total and that all three solutions are equivalent.
-
-\begin{wrapfigure}[11]{r}{0.45\textwidth}
-	\centering
-	\includegraphics[width=.45\linewidth]{circle_true}
-	\hfill
-	\includegraphics[width=.45\linewidth]{circle_false}
-	\caption{Variable gadget representing a true variable (a) and a false variable (b)}
-	\label{fig:optimal_circles}
-\end{wrapfigure}
-
-Circles of segments (variable gadgets) are constructed in a very similar fashion.
-Let $m$ be the maximum number of clauses in which a single variable appears.
-Every variable $x_n$ will be represented by $m$ pairs of segments one red and one green which are connected as seen in Fig.~\ref{fig:gadgets} (b) and form a circle.
-It can be observed that every segment $s_n$ intersects exactly 2 other segments $s_{n-1}$ and $s_{n+1}$ ($s_{2m}$ intersects with $s_1$ to create the circle).
-These intersections split every segmet in three parts, which again have length 1 for the stubs and 2 for the middle part.
-If $x$ is true, the green edges will be drawn completely, if it is false the red edges will be drawn completely.
-Observe that fixing a green edge to be drawn completely fixes all green edges to be drawn completely, like in Fig.~\ref{fig:optimal_circles} (a) (in order to have an optimal solution for this circle) and conversely for the red edges, like in Fig.~\ref{fig:optimal_circles} (b).
-We will associate the first version, with subdivided green edges, with the variable being true and the second one with the variable being false.
-Since all edges are cut in the same fashion, the two solutions are equivalent in terms of maximized ink and amount to $4m + 2m = 6m$.
-
-The variable clauses are arranged on a spine and the clause gadgets are placed above or below the spine, according to the planar embedding of the \ppsat.
-Then the clause gadgets are connected to the variable gadgets corresponding to the occuring variables via segment chains (see Fig.\ref{fig:gadget_connection}, segment chains are for the sake of visualization not all equally long).
-These chains are composed of $t$ (where $t$ is an sufficiently large even integer) equal length segments $s_1$, \dots, $s_t$, which are again coloured alternatingly red and green.
-Every segment $s_i$ crosses with $s_{i-1}$ (except for $s_1$) and $s_{i+1}$ (except for $s_t$).
-All segments are divided again into 3 parts with the stubs having length 1, and the middle part having length 2.
-$s_1$ is also split in this ratio by one of the green stubs of a variable gadget, however the stub is intersected such that the remaining stub has length $1-\varepsilon$.
-The same holds for $s_t$ and a stub of an edge in the clause gadget.
-Here again the optimal solution in terms of maximizing ink is by drawing either all green edges completely and all red ones as stubs, or vice versa.
-Both solutions have a value of $4\frac{t}{2} + 2\frac{t}{2} = 3t$.
-It should be noted that completely drawing the red edges cuts of a small piece (of length $\varepsilon$) of the stubs of both the green edge of the variable gadget and the edge of the clause gadget, to which the chain is connected.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.8\linewidth]{variable_clause_connection}
-	\caption{Connection between the clause and variable gadgets, created by chains of an even number of segments.}
-%	\label{fig:gadget_connection}
-\end{figure}
-
-% Let $q$ be the number of clauses and $p$ be the number of distinct variables in $\phi$.
-%
-% \begin{lemma}
-% 	\label{lem:optimal value}
-% 	The construcetd Graph G has an optimal  value in terms of maximizing ink of $8q + 6mp + 9qt - (4q + 2p)\varepsilon$ if $\varepsilon$ is chosen sufficiently small.
-% \end{lemma}
-%
-% \begin{proof}
-% 	\soeren{i have an idea for this but i don't know if it would be out of scope, or if i might not even need this lemma.
-Also is our \maxsped defined as "Does this graph have a drawing with at least x amount of ink?" since thats what this bound is aiming at?"}
-% \end{proof}
-\soeren{it also still needs to be shown that this bound ensures that the two criterias from~\cite{b-sgh-15}, namely that at least one edge in the triangle is drawn and that the segment chains are alternatingly drawn and not drawn.}
-
-	Let $\phi$ be a boolean formula and $\Pi$ be a planar embedding corresponding to that formula sucht that ($\phi$, $\Pi$) is an instance of \ppsat.
-Let $\Gamma$ be a drawing constructed from that instance in the previously described way.
-We will show that there is a valid truth assignement for $\phi$ if and only if $\Gamma$ admits a \maxsped with a value of at least $8q + 6mp + 3q (3t - 4\varepsilon)$ in terms of maximizing ink.
-	~
-	\begin{enumerate}
-		\item[$\Rightarrow$:] Let $x$ be a variable and $c$ be a clause containing $x$.
-If $x$ is true draw all red edges in the associated variable gadget completely, which in turn means all green edges are drawn partially and vice versa for when $x$ is false.
-By drawing the green edges partially we enable the red edges in the segment chains to be drawn completely (if we shorten the stubs for one green edge in the variable gadget by $\varepsilon$).
-this also means we can draw the edges of the clause gadgets to which this segment chain is connected completely.
-Since $\phi$ has a satisfying variable assignement where exactly one variable per clause is true we have exactly one of these segment chains which allows exactly one edge per clause graph to be drawn completely.
-This results in every clause graph having the value $8 - 4\varepsilon$, every variable gadget having the value of $6m - 2\varepsilon$ and every segment chain having the value of $3t$.
-Let $q$ be the number of clauses and $p$ be the number of distinct variables in $\phi$.
-This results in
-
-		\soeren{is there a nice way to format this?}
-		$(8 - 4\varepsilon)q + (6m - 2\varepsilon)p + 3t*3q = $
-		$8q + 6mp + 9qt - (4q + 2p)\varepsilon$
-
-		\item[$\Leftarrow$:]  Assume $\phi$ does not have a stisfying variable assignement.
-Then there must always exist a clause in which all three contained variables $x_1$, $x_2$ and $x_3$ are false.
-That however means that, by construction, there is a clause gadget where all the corresponding variable gadgets have all their green edges completely drawn.
-This adds at most the value of $2\varepsilon$.
-In order to get the maximal drawing, this propagates through the segment chains up to the clause gadget where it causes all three edges of that gadget to be not drawn.
-This in turn means that this specific gadget has only the value $6 - 6\varepsilon$.
-The total value of this drawing in terms of maximizing ink is therefore at least $2$ smaller than normal.
-If we would try to supplement the missing drawn edge of the clause gadget, we would need to switch up the alternation of drawing segments in the chains, which causes 2 consecutive segments to not be drawn.
-Since the last edge in the chain was dran before, it now can't be drawn and it therefore again results in a difference of at least $2$ in the complete drawing.
-	\end{enumerate}
-
-\soeren{TODO: argue equivalence (both direction)}
-
-\end{proof}

### 2018-10-22 eurocg/short_abstract/abstract.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%% article.tex %%%%%%%%%%%%%%%%%%%%%%%
%
% edit this file for the abstract of your own talk
% and save it under the original name "abstract.tex".
% Then run "pdflatex eurocg18-short-abstract.tex" to see how
% your abstract will look like.
%

\authors{Michael H\"oller, Fabian Klute, Soeren Nickel, Martin~N\"ollenburg, \speaker{Birgit Schreiber}}
% The author who will give the presentation should be marked with the
% "\speaker{...}" command

\title{Maximizing Ink in Symmetric Partial Edge Drawings of $k$-plane Graphs}

% Exactly one of the following three lines should be present
%

%\nofigures
\onefigure{MISPE-fig} % In this case, we have one thumbnail figure
%\twofigures{SIRAR-fig1}{SIRAR-fig2}

% To distinguish your figures from those of other particupants,
% the file names should start with the first five initials of the
% title. (SIRAR in the example)
%
% The figures must be PDF files with the extension .pdf
% (SIRAR-fig1.pdf in the example), and they must reside in the same
% directory as the main file.
%
% The figures will be scaled to a height of 25 millimeters and
% enclosed by boxes with a margin of 3pt. It is best to design your
% figures in this size. Otherwise, you should make sure that the
% thickness of your lines survives the rescaling.

% By default, the figures will be enclosed in a box.
% If your figures already have an encloding box, you can disable the
% addition of an enclosing box by uncommenting the following line:
%
% \FigureBorderfalse

\begin{abstract}
  % Here comes the text of your abstract.
  % A good length is 6-15 lines in the output. The limit is 20 lines.
  % The presence of thumbnail pictures reduces the available amount
  % for text.
%
	We study symmetric partial edge drawings (SPEDs) for non-planar graphs, in which edges are drawn only partially as pairs of opposing stubs of equal length on the respective end-vertices. It is known that maximizing the ink (or the total stub length) when transforming a straight-line drawing with crossings into a SPED is tractable for 2-plane input drawings, but generally NP-hard. We show that the problem remains NP-hard even for 3-plane input drawings. We also present efficient algorithms for ink maximization of $k$-plane input drawings whose edge intersection graph forms a collection of trees or cacti. %\verb|\FigureBorderfalse|.

\end{abstract}

%%%%%%%%%% =========================================
